### 2022-08-11 Scripts/ASR_stocks.Rmd

---
title: "ICM testing"
output:
  word_document: 
    reference_docx: template.docx
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

The stocks chosen for testing were those with the most complete data.

Some catch information was missing for certain years. These years are marked with "?" in the plots that follow. The median annual catch was assumed for missing years.

Catch data for _Sprattus sprattus_ (European sprat) was manually converted to be on the 1000s scale, in order to align with abundance data.

```{r, echo=F, message=F, warning=F, fig.width=6, fig.height=4}
# OK, so using the ICES assessments here's what we get for North Sea cod.
library(readxl)
library(tidyverse)
library(rio)
library(ggthemes)

# # OK, lets give this a spin...
# funs <- c("https://raw.githubusercontent.com/Dave-Keith/ICM/main/Scripts/functions/backwards_sim.r")
# # Now run through a quick loop to load each one, just be sure that your working directory is read/write!
# for(fun in funs) 
# {
#   download.file(fun,destfile = basename(fun))
#   source(paste0(getwd(),"/",basename(fun)))
#   file.remove(paste0(getwd(),"/",basename(fun)))
# }

source("./functions/backwards_sim.R")

# Choose 5 ICES stocks that we have the necessary data for
ASR <- read_xlsx("../Data/ASR_2018.xlsx", sheet = "ICES")
datatypes <- unique(gsub(x = names(ASR), pattern = "[^a-zA-Z]", replacement=""))
# we want:
# Year, Num, WA, Catch, AM, NM, StockID, Management, Area, Order, Family, Genus, Species
ASRdat <- ASR[,c(grep(x=names(ASR), "Num"),
                 grep(x=names(ASR), "WA"),
                 grep(x=names(ASR), "Catch"),
                 grep(x=names(ASR), "AM"),
                 grep(x=names(ASR), "NM"))]
ASRdat <- apply(X = ASRdat, 2, as.numeric)
ASRsp <- ASR[, which(!1:length(names(ASR)) %in% grep(x=names(ASR), ".", fixed=T))]
ASR_trim <- cbind(ASRsp, ASRdat)

# need a unique ID for stock
#table(ASR_trim$Management, ASR_trim$Species)
ASR_trim$Stock <- paste0(ASR_trim$Management, "_", ASR_trim$Area, "_", ASR_trim$Genus, "_", ASR_trim$Species)

ASR_long <- ASR_trim %>%
  pivot_longer(!c("Management", "Area", "Order", "Family", "Genus", "Species", "Stock", "Year")) %>%
  separate(col=name, into=c("type", "age"), sep = "\\.")

# summary plots for ALL stocks
for (i in unique(ASR_long$Stock)) {
  ASR_sub <- ASR_long %>%
    select(Year, Stock, type, age, value) %>%
    filter(Stock==i)
  maxage <- max(as.numeric(ASR_sub$age[!is.na(ASR_sub$value)]))
  maxYear <- max(ASR_sub$Year[!is.na(ASR_sub$value)])
  minYear <- min(ASR_sub$Year[!is.na(ASR_sub$value)])
  forplot <- ASR_sub[as.numeric(ASR_sub$age) < (maxage+1) & ASR_sub$Year %in% minYear:maxYear,]
  p1 <- ggplot() + geom_line(data=forplot, aes(Year, value)) + facet_grid(type~as.numeric(age), scales="free_y") +
        geom_vline(data=forplot[is.na(forplot$value),], aes(xintercept=Year), colour="red") + facet_grid(type~as.numeric(age), scales="free_y") +
        theme_bw() + 
        ggtitle(i) +
    guides(x =  guide_axis(angle = 90)) 
  print(p1)

  if(!dir.exists(paste0("../Figures/", i))) dir.create(paste0("../Figures/", i))

  png(filename=paste0("../Figures/", i, "/data_summary_", i, ".png"), height=4, width=12, res=400, units="in")
  print(p1)
  dev.off()
}

Stocks <- ASR_long %>%
  filter(!is.na(value)) %>%
  group_by(Stock, type) %>%
  summarize(count=length(unique(value))) %>%
  group_by(Stock) %>%
  summarize(types=length(unique(type))) %>%
  filter(types==5) %>%
  select(Stock)

ASR_stocks <- ASR_long %>%
  filter(Stock %in% Stocks$Stock) %>%
  filter(!is.na(value)) %>%
  arrange(Stock, Year, type, as.numeric(age))

Stocks <- ASR_stocks %>%
  group_by(Stock, Species, type) %>%
  summarize(ages=length(unique(age)),
            years=length(unique(Year))) %>%
  arrange(-years, -ages) %>%
  filter(!Species=="morhua") %>%
  distinct(Stock) %>%
  pull(Stock)

print(Stocks)

# Bring in the data
# Handy import function from the rio package... tricky part was getting the right filename here...
for(i in Stocks){
  print(i)
  ASR_sub <- ASR_long %>%
    select(Year, Stock, type, age, value) %>%
    filter(Stock==i)

  # maxage <- max(as.numeric(ASR_sub$age[!is.na(ASR_sub$value)]))
  # maxYear <- max(ASR_sub$Year[!is.na(ASR_sub$value)])
  # minYear <- min(ASR_sub$Year[!is.na(ASR_sub$value)])
  # forplot <- ASR_sub[as.numeric(ASR_sub$age) < (maxage+1) & ASR_sub$Year %in% minYear:maxYear,]
  # 
  # # summary plots
  # print(ggplot() + geom_line(data=forplot, aes(Year, value)) + facet_grid(type~as.numeric(age), scales="free_y") +
  #         geom_vline(data=forplot[is.na(forplot$value),], aes(xintercept=Year), colour="red") + facet_grid(type~as.numeric(age), scales="free_y") +
  #         theme_bw() + 
  #         ggtitle(i)
  # )

  age.mat <- ASR_sub %>% filter(type=="AM") %>% rename(AM=value) %>% select(-type)
  nat.mort <-  ASR_sub %>% filter(type=="NM") %>% rename(NM=value) %>% select(-type)
  abund <- ASR_sub %>% filter(type=="Num") %>% rename(Num=value) %>% select(-type)
  weight.age <- ASR_sub %>% filter(type=="WA") %>% rename(WA=value) %>% select(-type)
  removals <- ASR_sub %>% filter(type=="Catch") %>% rename(Catch=value) %>% select(-type)

  if(i=="ICES-WGBFAS_BS 22-32_Sprattus_sprattus") removals$Catch <- removals$Catch/1000

  data <- age.mat %>%
    full_join(nat.mort) %>%
    full_join(abund) %>%
    full_join(weight.age) %>%
    full_join(removals)

  data$available <- apply(is.na(data[, c("AM", "Num", "WA")]), 1, function(x) all(!x==T))
  data <- data[data$available==T,]

  # Tidy up the data for input...
  data$prop.nat.mort <- 1-exp(-data$NM)
  #prop.nat.mort[,-which(names(prop.nat.mort) %in% c("Year", "Stock", "type"))] <- 1-exp(-prop.nat.mort[,-which(names(prop.nat.mort) %in% c("Year", "Stock", "type"))])

  rem <- data %>% group_by(Year) %>% summarize(rem=sum(Catch)) %>% pull(rem)

  missing_rem<- NULL
  if(any(is.na(rem))) {
    missing_rem <- unique(data$Year)[which(is.na(rem))]
    rem[is.na(rem)] <- median(rem, na.rm=T)
  }

  #rowSums(removals[,-which(names(removals) %in% c("Year", "Stock", "type"))], na.rm=T)
  years <- data %>% pull(Year) %>% unique() %>% sort()
  N.end <- sum(data[data$Year==max(years),]$Num)
  #N.end <- rowSums(abund[nrow(abund),-which(names(abund) %in% c("Year", "Stock", "type"))], na.rm=T)
  #vpa.abund <- rowSums(abund[,-which(names(abund) %in% c("Year", "Stock", "type"))], na.rm=T)
  vpa.abund <- data %>% group_by(Year) %>% summarize(vpa=sum(Num)) %>% pull(vpa)

  # The real mx matrix, recruits produced per individual in each age class... Not perfect as I need to offset recruits/ssb, but close enough for the moment..
  #minage
  minage <- min(as.numeric(data$age))
  #maxage
  maxage <- max(as.numeric(data$age))
  #recruits
  annual <- data.frame(Year=data$Year[data$age==minage], recruits=data$Num[data$age==minage])
  #ssn
  data$ssn <- data$Num * data$AM
  #ssb
  data$ssb <- data$ssn * data$WA
  #tot.ssb
  annual <- data %>% group_by(Year) %>% summarize(tot.ssb = sum(ssb)) %>% left_join(annual)
  #r.p.ssb
  annual$r.p.ssb <- annual$recruits/annual$tot.ssb
  # recs.per.age
  data <- left_join(data, annual)
  data$recs.per.age <- data$ssb*data$r.p.ssb
  # mx
  data$mx <- data$recs.per.age/data$ssn/2 # Moms only! Dividing by around 8 really nails it for NS cod for whatever reason :-)
  # Something is wrong with the decline rate method, but the exponential and logistic are working pretty... pretty... pretty good...

  age.mat <- data %>% select("Year", "age", "AM") %>% pivot_wider(names_from=age, values_from = AM) %>% select(-Year)
  prop.nat.mort <- data %>% select("Year", "age", "prop.nat.mort") %>% pivot_wider(names_from=age, values_from = prop.nat.mort) %>% select(-Year)
  weight.age <- data %>% select("Year", "age", "WA") %>% pivot_wider(names_from=age, values_from = WA) %>% select(-Year)
  mx <- data %>% select("Year", "age", "mx") %>% pivot_wider(names_from=age, values_from = mx) %>% select(-Year) %>% as.data.frame()

  tst <- icm.sim(years = years,
                 mat.age = age.mat,
                 nm = prop.nat.mort,
                 w.age = weight.age,
                 ages = minage:maxage,
                 rems = rem,
                 fecund = mx,
                 N.end = N.end,
                 pop.model = 'exponential',
                 n.sims = 5,
                 sd.mat = 0.5,
                 sd.nm = 0.5,
                 sd.wt = 0.5,
                 sd.fecund = 0.5)

  #Combine the data
  did.it.work <- data.frame(abund = c(vpa.abund,tst$Pop$abund),years = c(years,tst$Pop$years),sim = c(rep('VPA',length(years)),tst$Pop$sim))
  # Here get the Upper and lower 50% quantiles to make a functional boxplot
  quants <- did.it.work %>% dplyr::group_by(years) %>% dplyr::summarise(L.50 = quantile(abund,probs=c(0.25)),
                                                                        med = median(abund),
                                                                        U.50 = quantile(abund,probs=c(0.75)))

  did.it.work$missing[did.it.work$years %in% missing_rem] <- "missing"

  # All the results in a plot
  p2 <- ggplot() +
          geom_line(data=did.it.work, aes(x=years,y=abund,color=sim)) +
          xlab("") +
          ylab("Abundance (1000s)") +
          geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + scale_color_viridis_d(end = 0.75) +
          ggtitle(i) +
          geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  # Same thing but functional boxplots
  p3 <- ggplot() +
          geom_line(data=quants, aes(x=years,y= med)) +
          geom_ribbon(data=quants, aes(x=years,ymax=U.50,ymin = L.50),alpha=0.5,fill='blue',color='blue') +
          geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          xlab("") + ylab("Abundance (1000s)") +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') +
          ggtitle(i)+
          geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  # Plotting how the estimate of r changes over time (which you can do when we have time varying inputs (e.g. Wgt, M, or maturity at age)
  p4 <- ggplot(tst$r) +
          geom_line(aes(x=years,y=r,group=n.sims,color=n.sims)) +
          xlab("") +
          ylab("Estimated growth rate (r)") +
          #scale_y_continuous(breaks = seq(0,1.5,by=0.1)) +
          #scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') +
          scale_color_viridis_c(end = 0.75) +
          ggtitle(i)

  print(p2)
  print(p3)
  print(p4)

  png(filename=paste0("../Figures/", i, "/runs_vs_vpa_", i, ".png"), height=4, width=6, res=400, units="in")
  print(p2)
  dev.off()
  png(filename=paste0("../Figures/", i, "/runs_vs_vpa_ribbon_", i, ".png"), height=4, width=6, res=400, units="in")
  print(p3)
  dev.off()
  png(filename=paste0("../Figures/", i, "/growth_", i, ".png"), height=4, width=6, res=400, units="in")
  print(p4)
  dev.off()
}
```