### 2024-04-16 Drafts/Paper_1/07-tables.Rmd

---
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---
\clearpage
\newpage

# Tables {-}