### 2024-04-16 Drafts/Paper_1/08-figures.Rmd

---
output:
  pdf_document:
    fig_caption: yes
    extra_dependencies: float
linkcolor: blue
header-includes: \usepackage{float}
editor_options:
  chunk_output_type: console
---

```{r, include=F, echo=F}
knitr::opts_knit$set(eval.after = "fig.cap")
knitr::opts_chunk$set(fig.pos = 'H')
options(knitr.kable.NA = '')
require(tidyverse)
options(tinytex.verbose = TRUE)

```