### 2024-04-25 Drafts/Paper_1/02-intro.Rmd

# INTRODUCTION  {#intro}

Intro – life history paragraph, environmental variation paragraph, management ties paragraph, goals.

theoretical expectations related to productivity and life history for various species (e.g. herring vs. Cod) (ties to regional variation in lifetime reproductive rates and doubling time).

Evidence for specific species to make use of environmental conditions – boom in recruitment.

How our understanding of productivity is used in a management context – projections, reference points, risk, (ties directly to density dependence figure).  
Species-environment relationship – is it different from what we expect and how does that influence our capacity for management of exploited fishes.

The underappreciated influence of regional variation in determining recovery potential of exploited teleosts.  - can’t look at a single species and expect it will behave the same way.

Objectives
  - Quantify the life history variability of exploited marine teleosts
  - Explore regional and taxonomic correlates of life history variability
  - Based on the observed life history, quantify the recovery potential of marine teleosts in the absence of fishing