### 2024-04-25 Drafts/Paper_1/05-discussion.Rmd

# DISCUSSION {#disin}

I think we disagree with @kindsvaterTenPrinciplesEvolutionary2016 because we can't generalize based on life-history very easily (cod haddock herring).

Looks like density dependence is real. Dave's pet theroy is there is solid evidence of density dependence for many stocks above around 0.4, below 0.4 the evidence hints more at density independence for many stocks, BUT, some stocks have an ability to very occasionally very spectacularly take advantage of the environment.  But not all stocks and species have this ability and the strength and frequency seems to be influenced by where you life and what species you are. Little evidence for spectacular reproduction events in NW Atlantic, while the majority of species assessed in the north sea have this ability. Cod show this dicomaty perfectly....

Looks like the median lambda is around 0.2 at low abundance (mean is much higher because outliers) for many stocks. So potential growth at low abundance is around 20%, tends to trend towards 0 as you approach the maximum observed abundance. This is very interesting b/c on average is suggests that the maximum observed abundance may be a reasonable proxy for the carrying capacity of a stock. Of course there is a great deal of variability around this, which may be related to the timing of the start of the time series in relation to historical fishing activity, fishing pressure, environmental variability, etc... but still it is kinda interesting.

The recovery potential for some species is tied to the variability in their growth rate... Based on historically observed lambdas, all the Haddock stock in this analysis were likely to double in size within a decade of fishing cessation. The worst situation is low lambda + low variability... you can have a low lamda 'most of the time', but if highly variable the stock can take off (this in many ways is what fisheries science is based on)... Also of note these large events do tend to be far more likely when abundance is < 40% of maximum (of course this would be in part because it's easier to go from having \$10 to \$20 than it is to go from \$1,000 to \$2,000).

Idea- realized productivity of cod/haddock has been highly variable. But from life history theory you would have expected that to be herring/smelt. Episodic recruitment events. (boom and bust?) not population boom but more like a baby boom.

Likelihood of stock collapse related to life history characteristics, maybe reframe relative to large productivity events of cohorts – link to idea of cod/haddock above. No expectation that intersection between life history traits and environmental conditions to result in such marked differences among populations within a species in realized productivity.

Center around the idea of equilibrium theory? That management might be fairly good for stocks that aren’t very variable, but when you make a decision based on averages you can make really bad ones.