### 2024-05-09 Drafts/Paper_1/05-discussion.Rmd

# DISCUSSION {#disin}

I think we disagree with @kindsvaterTenPrinciplesEvolutionary2016 because we can't generalize based on life-history very easily (cod haddock herring).

Looks like density dependence is real. Dave's pet theory is there is solid evidence of density dependence for many stocks above around 0.4, below 0.4 the evidence suggest that density dependence weakens for some stocks, BUT, some stocks have an ability to very occasionally very spectacularly take advantage of the environment.  But not all stocks and species have this ability and the strength and frequency seems to be influenced by where you live and what species you are. Little evidence for high lambda for most stocks in the NW Atlantic, while many stocks assessed in the North Sea had anomalously high lambdas in 1 or more years. Cod show this  perfectly....

Looks like the median lambda is around 0.2 at low abundance (mean is much higher because outliers) for many stocks. So potential growth at low abundance is around 20%, tends to trend towards 0 as you approach the maximum observed abundance. This is very interesting b/c on average is suggests that the maximum observed abundance may be a reasonable proxy for the carrying capacity of a stock. Of course there is a great deal of variability around this, which may be related to the timing of the start of the time series in relation to historical fishing activity, fishing pressure, environmental variability, etc... but still it is kinda interesting.

The recovery potential for some species is tied to the variability in their growth rate... Based on historically observed lambdas, all the Haddock stock in this analysis were likely to double in size within a decade of fishing cessation. The worst situation is low lambda + low variability... you can have a low lamda 'most of the time', but if highly variable the stock can take off (this in many ways is what fisheries science is based on)... Also of note these large events do tend to be far more likely when abundance is < 40% of maximum (of course this would be in part because it's easier to go from having \$10 to \$20 than it is to go from \$1,000 to \$2,000).

Idea- realized productivity of cod/haddock has been highly variable. But from life history theory you would have expected that to be herring/smelt. Episodic recruitment events. (boom and bust?) not population boom but more like a baby boom.

Likelihood of stock collapse related to life history characteristics, maybe reframe relative to large productivity events of cohorts – link to idea of cod/haddock above. No expectation that intersection between life history traits and environmental conditions to result in such marked differences among populations within a species in realized productivity.

Center around the idea of equilibrium theory? That management might be fairly good for stocks that aren’t very variable, but when you make a decision based on averages you can make really bad ones.

By ignoring the density dependence patterns observed in this analysis, the doubling time analysis would overestimate the doubling time for stocks that were at low abundance with classic negative density dependence dynamics, while it would underestimate it for stocks with positive density dependence. Here perhaps talk about infrequent big events and how using a theoretical rmax within a logistic model isn't a good idea because this will way over-estimate recovery potential at low abundance.

The optimization procedure assumes that both the estimates of abundance and the fishery removals from the stock assessment are correct. *A couple sentences about why fisheries removals may be underestimated*. This would result in the estimates of the $\lambda$ excluding fishing mortality in this analysis being lower than reality, as this 'missing' mortality was accounted for by reductions in either fecundity or an increase in natural mortality. The optimization routine resulted in a median decrease in the no-fishing $\lambda$ estimates of approximately `r med.lam.per.diff`%, the overall directionality of this change is consistent with fishing mortality being generally underestimated.

Fisheries appear to slightly weaken the effects of density dependence overall. Maybe due to fisheries having positive density dependence? Seemingly below around 0.5 fishing leads to a positive lambda, below 0.5 it is a negative lambda... that's kinda kooky cool!

The stocks which reached very low abundance (< 10% of maximum) tended to have high values of lambda, suggesting strong compensation in stocks when they are at extremely low abundance. Oddly, the path to this isn't linear, and positive density dependence (or at least density independence was evident between 10 and 40% of maximum abundance).

Probably something worth citing in here @thorsonPredictingRecruitmentDensity2020