### 2024-05-21 Drafts/Paper_1/paper_1.Rmd
mid-edit update

---
title: |
   "Life history variability influence on the recovery potential of globally exploited teleosts "
french_title: "Titre ici (*Nom latin de l'espèce*)"
author: |
  David. M. Keith^1^ 
  Freya M. Keyser^1^
  Julie M. Charbonneau^2^
  Heather 'The Boss' Bowlby^1^
author_list: "Keith, D.M., Keyser, F.M., Charbonneau, J.M., Bowlby, H.T.B."
address: |
  ^1^Bedford Institute of Oceanography\
     Fisheries and Oceans Canada, 1 Challenger Dr.\
     Dartmouth, Nova Scotia, B2Y 4A2 , Canada\
  \smallskip
  ^2^Simon Fraser University\
     Department of Biology\
french_address: |

month: ""
french_month: ""
year: 2024
report_number: nnn
region: "Maritimes Region"
french_region: "Maritimes Region"
isbn: ""
cat_no: ""
citation_other_language: ""

abstract: |
  Here is the abstract text. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

  Start new paragraphs after a blank line and with 2 spaces indent. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

french_abstract: |

header: "Draft working paper --- Do not cite or circulate" # or "" to omit
output:
 csasdown::resdoc_word:
   french: false
   copy_sty: true
   line_nums: true
   line_nums_mod: 1
   lot_lof: false
   draft_watermark: false
   include_section_nums: true
   highlight: tango
# ------------
# End of options to set
knit: (function(input, ...) {
       csasdown::render('_bookdown.yml')
      })
link-citations: true
bibliography: Y:/Zotero/MAR_SABHU.bib
# Any extra LaTeX code for the header:
# header-includes:
# - \usepackage{tikz}
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, results='hide', warning=FALSE}
library(knitr)
if (is_latex_output()) {
  knitr_figs_dir <- "knitr-figs-pdf/"
  knitr_cache_dir <- "knitr-cache-pdf/"
  fig_out_type <- "png"
} else {
  knitr_figs_dir <- "knitr-figs-docx/"
  knitr_cache_dir <- "knitr-cache-docx/"
  fig_out_type <- "png"
}
fig_asp <- 0.618
fig_width <- 9
fig_out_width <- "6in"
fig_dpi <- 180
fig_align <- "center"
fig_pos <- "htb"
opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.path = knitr_figs_dir,
  cache.path = knitr_cache_dir,
  fig.asp = fig_asp,
  fig.width = fig_width,
  out.width = fig_out_width,
  echo = FALSE,
  #  autodep = TRUE,
  #  cache = TRUE,
  cache.comments = FALSE,
  dev = fig_out_type,
  dpi = fig_dpi,
  fig.align = fig_align,
  fig.pos = fig_pos
)
```

```{r load-libraries-n-data, echo = FALSE, cache = FALSE, message = FALSE, results = 'hide', warning = FALSE}
# add other packages here:
library(dplyr)
library(readr)
library(tibble)
library(csasdown)
# OK, so using the ICES assessments here's what we get for North Sea cod.
library(readxl)
library(tidyverse)
library(rio)
library(ggthemes)
library(cowplot)
library(gtable)
library(viridis)
library(arm)
library(mgcv)
library(lme4)
library(scales)
library(ggh4x)
library(gratia)
theme_set(theme_few(base_size = 14))

# The location of your data
loc <- 'D:/GitHub/ICM'
# The parameters for the doubling time simulation
n.dt.sims <- 1000 # Increase to 1000 for final run, which takes like a full day or something 
n.sim.years <- 100
# If we already have the results don't re-run the models or load data we don't need.
load.results <- F
# What tunning model are you running or did you run!
what.2.tune <- 'fec_nm'

#source(paste0(loc,"/Scripts/functions/tuning_sim_fast.R"))

# Download the functions we'll need from github
funs <- c("https://raw.githubusercontent.com/dave-keith/ICM/main/Scripts/functions/simple_Lotka_r.r",
          "https://raw.githubusercontent.com/dave-keith/ICM/main/Scripts/functions/backwards_project.r",
          "https://raw.githubusercontent.com/dave-keith/ICM/main/Scripts/functions/forward_project.r",
          "https://raw.githubusercontent.com/dave-keith/ICM/main/Scripts/functions/tuning_sim_fast.r"
)
  # Now run through a quick loop to load each one, just be sure that your working directory is read/write!
for(fun in funs) 
{
  download.file(fun,destfile = basename(fun))
  source(paste0(getwd(),"/",basename(fun)))
  file.remove(paste0(getwd(),"/",basename(fun)))
}

if(load.results==F)
{
  # Choose 5 ICES stocks that we have the necessary data for
  #ASR <- read_xlsx("../Data/ASR_2018.xlsx" sheet = "ICES")
  ASR1 <- read.csv(paste0(loc,"/Data/ICM_data_NE_atlantic_stocks.csv"))
  ASR <- read.csv(paste0(loc,"/Data/ICM_data_NW_atlantic_stocks.csv"))

  ASR <- rbind(ASR1,ASR)
  datatypes <- unique(gsub(x = names(ASR), pattern = "[^a-zA-Z]", replacement=""))

  # Replace 0's in Num.tot with NA so the rest of this works as Freya designed it to...
  ASR$Num.tot[ASR$Num.tot == 0] <- NA
  # we want:
  # Year, Num, WA, Catch, AM, NM, StockID, Management, Area, Order, Family, Genus, Species
  ASRdat <- ASR[,c(grep(x=names(ASR), "Num"),
                   grep(x=names(ASR), "WA"),
                   grep(x=names(ASR), "Catch"),
                   grep(x=names(ASR), "AM"),
                   grep(x=names(ASR), "NM"))]
  # Making all the data numeric that should be numeric
  ASRdat <- apply(X = ASRdat, 2, as.numeric)
  # Getting the species info back
  ASRsp <- ASR[, which(!1:length(names(ASR)) %in% grep(x=names(ASR), ".", fixed=T))]
  # And binding it all back together
  ASR.trim <- cbind(ASRsp, ASRdat)
  # need a unique ID for stock
  #table(ASR.trim$Management, ASR.trim$Species)
  ASR.trim$Stock <- paste0(ASR.trim$Management, "_", ASR.trim$Area, "_", ASR.trim$Genus, "_", ASR.trim$Species)

  ASR.long <- ASR.trim %>%
    pivot_longer(!c("Management", "Area", "Order", "Family", "Genus", "Species", "Stock", "Year","Meeting_or_reference","Model","Case","Notes")) %>%
    separate(col=name, into=c("type", "age"), sep = "\\.")

  # Something is weird with the mentella stock, I'm going to drop it...
  ASR.long <- ASR.long %>% dplyr::filter(Stock != "ICES-AFWG_DEEP1-2_Sebastes_mentella")
}# end if load results
```

```{r bilingual-code, cache=FALSE}
options(
  # Prevent xtable from adding a timestamp comment to the table code it produces
  xtable.comment = FALSE,
  # Do not allow kableExtra to load packages, we add them manually in csasdown
  kableExtra.latex.load_packages = FALSE,
  # Stop chunk output (echo) running into the margins
  width = 80,
  # Do not use scientific notation (stops tables from showing 1.2e3, etc.)
  scipen = 999)

meta <- rmarkdown::metadata
meta_out <- rmarkdown::metadata$output
csl <- "csl/csas.csl"
options(OutDec = ".")
if (is.null(getOption("french"))) {
  stop("`french` was not set up correctly in YAML header in index.Rmd. ",
       "It must be true or false",
       call. = FALSE)
}
if (getOption("french")) {
  csl <- "csl/csas-french.csl"
  options(OutDec = ",")
}

french <- isTRUE(getOption("french")) # for backwards compatibility

# This hook simplifies document translation for the author.
# When building in French, it draws a box around paragraphs contained in chunks
#  which have the option `needs_trans = TRUE`. It also labels
#  the box with a "Needs translation" tag in red and the chunk label in blue.
# You need to change the `needs_trans` option to `FALSE` for a chunk once you
#  have inserted the translated text into it. You will get a utf-8 error if
#  you leave it as `TRUE` and there is French in the chunk.
#  This function assumes you are translating from English to French.
#  If you wrote your document in French and want to translate to English,
#  put a ! before `getOption("french")` below and add the `need_trans`
#  chunk options to your English paragraph chunks instead of the French ones.
#  
#  IMPORTANT NOTES
#  - Use `csasdown::render()` to render the document. This runs a
#    pre-processing step which ensures any inline R chunks
#    `r print("Like this one")` are taken care of correctly and that all
#    backslash variables (eg. \pi, \alpha, \@ref, \cite) are all processed
#    correctly.
#  - French latex places a space before the colon by default so if you need a colon
#    with no space before it, use \\hc .
knit_hooks$set(needs_trans = function(before, options){
  if(getOption("french") && options$needs_trans){
    if (before){
      paste0("\\
   \\begin{lrbox}{\\userinput}
   \\begin{minipage}{\\dimexpr\\linewidth-2\\fboxsep-2\\fboxrule}
   \\textcolor{red}{\\textbf{Needs translation - \\textcolor{blue}{knitr chunk: ", options$label, "}}}
   \\begin{lstlisting}
  ")
    } else {
      "
   \\end{lstlisting}
   \\end{minipage}
   \\end{lrbox}
   \\noindent
   \\fbox{\\usebox{\\userinput}}
  "
    }
  }
})
```

---
title: `r ifelse(fr(), meta$french_title, meta$title)`
month: `r ifelse(fr(), meta$french_month, meta$month)`
region: `r ifelse(fr(), meta$french_region, meta$region)`
csl: `r csl`
---

<!--chapter:end:tmp-index.Rmd-->

---
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r, prep-data,echo=F, message=F, warning=F, fig.width=6, fig.height=4,include=F,message=F}

if(load.results==F)
{
  Stocks <- ASR.long |> 
                      collapse::fsubset(!is.na(value)) |>
                      collapse::fgroup_by(Stock, type) |>
                      collapse::fsummarize(count=length(unique(value))) |>
                      collapse::fgroup_by(Stock) |>
                      collapse::fsummarize(types=length(unique(type))) |>
                      collapse::fsubset(types==5) |>
                      collapse::fselect(Stock)

  ASR.tmp <- ASR.long 
  ASR.tmp$age <- as.numeric(ASR.tmp$age)
  ASR.stocks <- ASR.tmp |> 
                          collapse::fsubset(Stock %in% Stocks$Stock) |>
                          collapse::fsubset(!is.na(value)) |>
                          collapse::roworder(Stock, Year, type, age)

  # print(Stocks)
  Stocks <- Stocks$Stock

  # Going to remove Sebastes norvegicus because we can't get fecundity as the only data we have is total numbers.
  #Stocks <- Stocks[Stocks != "ICES-AFWG_DEEP1-2_Sebastes_norvegicus"]
  # Esmarki doesn't work great as it is assessed using a quarterly model.
  #Stocks <- Stocks[Stocks != "ICES-WGNSSK_NS 4-3aN_Trisopterus_esmarkii"]
  #i = 'ICES-AFWG_NEA1-2_Melanogrammus_aeglefinus'
  years.tmp <- NULL
  pnm.tmp <- NULL
  waa.tmp <- NULL
  ages.tmp <- NULL
  rem.tmp <- NULL
  mx.tmp <- NULL
  NE.tmp <- NULL
  vpa.tmp <- NULL
  am.tmp <- NULL
  #mr.tmp <- NULL
  abund.tmp <- NULL
  full.abund.tmp <- NULL
  full.rem.tmp <- NULL
  fs.tmp <- NULL
  for(i in Stocks)
  {
  print(i)
  ASR.sub <- ASR.long |>
    collapse::fselect(Year, Stock, type, age, value) |>
    collapse::fsubset(Stock==i)

    # Dropping the first 7 years as the 8 year olds are all missing
  if(i == "ICES-WGNSSK_NS  4-6a-20_Melanogrammus_aeglefinus") ASR.sub <- ASR.sub |> collapse::fsubset(Year >= 1972)
  # maxage <- max(as.numeric(ASR.sub$age[!is.na(ASR.sub$value)]))
  # maxYear <- max(ASR.sub$Year[!is.na(ASR.sub$value)])
  # minYear <- min(ASR.sub$Year[!is.na(ASR.sub$value)])
  # forplot <- ASR.sub[as.numeric(ASR.sub$age) < (maxage+1) & ASR.sub$Year %in% minYear:maxYear,]
  # 
  # # summary plots
  # print(ggplot() + geom_line(data=forplot, aes(Year, value)) + facet_grid(type~as.numeric(age), scales="free_y") +
  #         geom_vline(data=forplot[is.na(forplot$value),], aes(xintercept=Year), colour="red") + facet_grid(type~as.numeric(age), scales="free_y") +
  #         theme_bw() + 
  #         ggtitle(i)
  # )

  age.mat <- ASR.sub |> collapse::fsubset(type=="AM") |> dplyr::rename(AM=value) |> collapse::fselect(-type)
  # All of the age at maturities are NA in 0 and 1 year olds, so make them 0's, do that carefully just
  # in case that changes later...
  am.0s <- length(which((!is.na(age.mat$AM[age.mat$age ==0]))))
  am.1s <- length(which((!is.na(age.mat$AM[age.mat$age ==1]))))
  am.2s <- length(which((!is.na(age.mat$AM[age.mat$age ==2]))))
  if(am.0s == 0) age.mat$AM[age.mat$age ==0] <- 0
  if(am.1s == 0) age.mat$AM[age.mat$age ==1] <- 0
  if(am.2s == 0) age.mat$AM[age.mat$age ==2] <- 0

  #if(am.1s | am.0s | am.2s > 0) print("Stop, you need to check the age at maturity for either age 0 or 1 as there is data in there.")
  nat.mort <-  ASR.sub |> collapse::fsubset(type=="NM") |> dplyr::rename(NM=value) |> collapse::fselect(-type)
  abund <- ASR.sub |> collapse::fsubset(type=="Num") |> dplyr::rename(Num=value) |> collapse::fselect(-type)
  weight.age <- ASR.sub |> collapse::fsubset(type=="WA") |> dplyr::rename(WA=value) |> collapse::fselect(-type)
  if(i == Stocks[30]) weight.age$WA <- weight.age$WA/1000
  removals <- ASR.sub |> collapse::fsubset(type=="Catch") |> dplyr::rename(Catch=value) |> collapse::fselect(-type)

  data <- age.mat |>
    full_join(nat.mort) |>
    full_join(abund) |>
    full_join(weight.age) |>
    full_join(removals)

  # Aahh??  Is this NA dropping useful data, needed to quick fix the age at maturity data...
  data$available <- apply(is.na(data[, c("AM", "Num", "WA")]), 1, function(x) all(!x==T))
  data <- data[data$available==T,]

  # Tidy up the data for input...
  data$prop.nat.mort <- 1-exp(-data$NM)
  #prop.nat.mort[,-which(names(prop.nat.mort) %in% c("Year", "Stock", "type"))] <- 1-exp(-prop.nat.mort[,-which(names(prop.nat.mort) %in% c("Year", "Stock", "type"))])

  rem <- data |> collapse::fgroup_by(Year) |> collapse::fsummarize(rem=sum(Catch,na.rm=T)) #|> dplyr::pull(rem)
  #rem <- data |> dplyr::group_by(Year,.drop=F) |> collapse::fsummarize(rem=sum(Catch,na.rm=T)) #|> dplyr::pull(rem)

  # Never run, so not doing anything.
  # missing_rem<- NULL
  # if(any(is.na(rem$rem))) 
  # {
  #   missing_rem <- unique(data$Year)[which(is.na(rem$rem))]
  #   rem$rem[is.na(rem$rem)] <- median(rem$rem, na.rm=T)
  # }

  #rowSums(removals[,-which(names(removals) %in% c("Year", "Stock", "type"))], na.rm=T)
  years <- data |> dplyr::pull(Year) |> unique() |> sort()

  N.end <- sum(data[data$Year==max(years),]$Num)
  vpa.abund <- data |> dplyr::group_by(Year) |> dplyr::summarize(vpa=sum(Num,na.rm=T)) |> pull(vpa)

  # The real mx matrix, recruits produced per individual in each age class... Not perfect as I need to offset recruits/ssb, but close enough for the moment..
  #minage
  minage <- min(as.numeric(data$age))
  #maxage
  maxage <- max(as.numeric(data$age))
  #recruits
  annual <- data.frame(Year=data$Year[data$age==minage], recruits=data$Num[data$age==minage])
  #ssn
  data$ssn <- data$Num * data$AM
  #ssb
  data$ssb <- data$ssn * data$WA #I need to figure out how to line this up with the number of recruits for the right years in the data object.... got it right for the overall r.p.ssb below.
  #tot.ssb
  annual <- data |> group_by(Year) |> summarize(tot.ssb = sum(ssb)) |> left_join(annual)

  #tst <- data |> group_by(Year) |> summarize(tot.ssn = sum(ssn)) |> left_join(annual)

  #r.p.ssb This needs to be offset by the age of recruits. So here's my lining this up logic.  For age 0, we get
  # the recruit and SSB estimates from the same survey, thus these age 0 recruits will have been produced by the SSB in the previous year...
  # So the R in 1970 were produced by SSB in 1969 when min age =0
  # If min.age = 1 then R in 1970 is produced by SSB in 1968... and so on.
  # FIX 1 "Recruit year" = SSB year + min.age + 1 (or min.age +2 as SSB.year is the first column)
  # Now for RPS there's another wrinkle because of this, because our model is N(1970) = lambda * N(1969) and lambda is made up of lx and mx
  # so if the recruits show up in 1970, the fecundity term that gets us there (irrespective of the age offset) has to be for 1969, becuase
  # that gets us to the babies that show up in 1970
  # FIX 2 So "RPS year, recs.per.age year, and mx year" = SSB year + min.age
  # I believe this to be correct, but we all need to discuss this to make sure the logic holds.
  annual$r.p.ssb <- c(rep(NA,minage),annual$recruits[(minage+2):nrow(annual)]/annual$tot.ssb[1:(nrow(annual)-minage-1)],NA)
  # recs.per.age
  data <- left_join(data, annual)
  # So I need to make an offset SSB in the 'data' object to line up with the correct r.p.ssb field.  This is probably gonna suck...
  tmp <- NULL
  for(j in 1:length(years))
  {
    tst <- data|> collapse::fsubset(Year == years[j]) |> collapse::fselect(ssb,ssn,Year,age)
    tst$Year <- tst$Year + minage # We want this to line up the the RPS year, so just adding minage here.
    names(tst) <- c("ssb.offset","ssn.offset","Year","age")
    tmp[[j]] <- tst
  }
  # Unpack the list
  ssb.off <- do.call('rbind',tmp)
  # And merge it with the data object
  data <- left_join(data,ssb.off,by=c("Year","age"))
  data$recs.per.age <- data$ssb.offset*data$r.p.ssb
  # mx
  # Some of the stocks are just the males and females, so these are already half the population.  Probably can just not include the male stocks in the end.
  # No longer dividing by 2 because we should really divide recruits and ssn by 2 to be female only, but that's pointless so we roll along with this.
  data$mx <- data$recs.per.age/data$ssn.offset # if(grepl("males",i)) 
  #if(!grepl("males",i))data$mx <- data$recs.per.age/data$ssn.offset/2 # Moms only! 
  data$mx[is.nan(data$mx)] <- 0 # if we don't have any spawners in an age class in a year their fecundity is 0
  # Easier to remove the years where we don't have recs.per.age...
  #data <- data |> collapse::fsubset(Year %in% years[(minage+1):length(years)])

  age.mat <- data |> collapse::fselect("Year", "age", "AM") |> 
                     pivot_wider(names_from=age, values_from = AM) |> collapse::fselect(-Year)
  prop.nat.mort <- data |> collapse::fselect("Year", "age", "prop.nat.mort") |> 
                            pivot_wider(names_from=age, values_from = prop.nat.mort) |> 
                            collapse::fselect(-Year) |> as.data.frame()

  weight.age <- data |> collapse::fselect("Year", "age", "WA") |> pivot_wider(names_from=age, values_from = WA) |> collapse::fselect(-Year)
  mx <- data |> collapse::fselect("Year", "age", "mx") |> 
                 pivot_wider(names_from=age, values_from = mx) |> 
                 collapse::fselect(-Year) |> as.data.frame()
  # Get the abundance matrix
  abund.age <- data |> collapse::fselect("Year", "age", "Num") |> 
                        pivot_wider(names_from=age, values_from = Num) |> 
                        collapse::fselect(-Year) |> as.data.frame()

  # Now we can save out the catch and get an instantaneous F
    data$Catch[is.na(data$Catch)] <- 0
    data$Exp <- data$Catch/data$Num 
    # If it is exactly 1 that leads to an Inf, which is annoying, so make it 1.1 and fix it with the rest below
    data$Exp[data$Exp == 1] <- 1.1
    # And make this an F
    data$Fs <- -log(1-data$Exp)
    if(any(is.na(data$Fs)))
    {
      fixers <- which(is.na(data$Fs))
      print(paste0("Had to replace ",length(fixers), ' exploitation rates for stock ',i))
      for(f in fixers)
      {
        yr <- data$Year[f]
        fs <- max(data$Fs[data$Year == yr],na.rm=T)
        data$Fs[f] <- fs
        data$Exp[f] <- 1-exp(-fs)
      }
    }

  full.rems <- data |> collapse::fselect("Year", "age", "Catch") |> pivot_wider(names_from=age, values_from = Catch) |> collapse::fselect(-Year)
  full.Fs <-  data |> collapse::fselect("Year", "age", "Fs") |> pivot_wider(names_from=age, values_from = Fs) |> collapse::fselect(-Year)

    if(i == "DFO_2J3KL_Gadus_morhua")
    {
      age.mat[,8:10] <- 1
      prop.nat.mort[,8:10] <- prop.nat.mort[,7]
      weight.age[12:14,9:10] <- weight.age[12:14,8]
      weight.age[15,10] <- weight.age[15,9]
      mx[12:16,9:10] <- mx[12:16,8]
      mx[17,10] <- mx[17,9]
      full.Fs[12:14,9:10] <- full.Fs[12:14,8]
      full.Fs[15,10] <- full.Fs[15,9]
      abund.age[12:14,9:10] <- abund.age[12:14,8]
      abund.age[15,10] <- abund.age[15,9]
      # I could calculate, but this will get us to the same place...
      full.rems[12:14,9:10] <- full.rems[12:14,8]
      full.rems[15,10] <- full.rems[15,9]
    }

    if(i == "DFO_4T-Spring_Clupea_harengus")
    {
      age.mat[6:7,10] <- 1
      prop.nat.mort[6:7,10] <- prop.nat.mort[6:7,9]
      weight.age[6:7,10] <- weight.age[6:7,9]
      mx[6:9,10] <- mx[6:9,9]
      full.Fs[6:7,10] <- full.Fs[6:7,9]
      abund.age[6:7,10] <- abund.age[6:7,9]
      full.rems[6:7,10] <- full.rems[6:7,9]
    }

    if(i == "ICES-HAWG_CS 6a- 7b-7c_Clupea_harengus") 
    {
      age.mat[30,8] <- 1
      prop.nat.mort[30,8] <- prop.nat.mort[30,7]
      weight.age[30,8] <- weight.age[29,8]
      mx[30:31,8] <- mx[30:31,9]
      full.Fs[30,8] <- full.Fs[30,7]
      abund.age[30,8] <- abund.age[30,7]
      full.rems[30,8] <- full.rems[30,7]
    }

    if(i == "ICES-WGCSE_IS6a-7b-7j_Dicentrarchus _labrax")
    {
      age.mat[c(14,27,40),5] <- age.mat[1,5]
      prop.nat.mort[c(14,27,40),5] <- prop.nat.mort[1,5]
      weight.age[c(14,27,40),5] <- weight.age[c(13,26,39),5]
      mx[c(14,16,27,29,40,42),5] <- mx[c(14,16,27,29,40,42),6]
      full.Fs[c(14,27,40),5] <- full.Fs[c(14,27,40),4]
      abund.age[c(14,27,40),5] <- abund.age[c(14,27,40),4]
      full.rems[c(14,27,40),5] <- full.rems[c(14,27,40),4]
    }
    # Clear out some 0's...
    if( i == "AFSC_EEBSAI_Limanda_aspera")
    {
      mx[5,18] <- mean(unlist(mx[5,c(17,19)]))
      mx[c(6,8),19] <- mx[c(6,8),18]
      mx[2:7,20] <- mx[2:7,19] 
    }

  # I had filled the years with no SSB data to inform them.  instead I'm going to remove those years...

   if(minage >0) 
  {
    rm <- -(1:minage)
    years <- years[rm]
    prop.nat.mort <- prop.nat.mort[rm,]
    weight.age <- weight.age[rm,]
    rem <- rem[rm,]
    mx <- mx[rm,]
    vpa.abund <- vpa.abund[rm]
    age.mat <- age.mat[rm,]
    abund.age <- abund.age[rm,]
    full.Fs <- full.Fs[rm,]
    full.rems <- full.rems[rm,]
   }
  #   mx.fill <- as.data.frame(matrix(rep(colMeans(mx,na.rm=T),minage),nrow=minage,byrow=T),colnames = names(mx))
  #   names(mx) <- names(mx.fill)
  #  # mx <- rbind(mx[(minage+1):nrow(mx),],mx.fill) # or is it
  #   mx <- rbind(mx.fill,mx[(minage+1):nrow(mx),]) # I think this is right!
  # }

    years.tmp[[i]] <- years
    pnm.tmp[[i]] <- prop.nat.mort
    waa.tmp[[i]] <- weight.age
    ages.tmp[[i]] <- minage:maxage
    rem.tmp[[i]] <- rem
    mx.tmp[[i]] <- mx
    NE.tmp[[i]] <- N.end
    vpa.tmp[[i]] <- vpa.abund
    am.tmp[[i]] <- age.mat
    #mr.tmp[[i]] <- missing_rem
    abund.tmp[[i]] <- abund.age
    fs.tmp[[i]] <- full.Fs
    full.rem.tmp[[i]] <- full.rems
  } #end input data loop

  save(years.tmp,pnm.tmp,waa.tmp,ages.tmp,rem.tmp,mx.tmp,NE.tmp,vpa.tmp,#mr.tmp,
       am.tmp,Stocks,ASR.stocks,ASR.long,abund.tmp,fs.tmp,full.rem.tmp,
       file =  paste0(loc,"/Results/model_inputs_no_age_correction.Rdata"))

  } # end load results
```

  So now I can try to do some tuning, both forwards and backwards, lets see what we get from that...

```{r tunes,include=F,echo=F,warning=F,message=F}

  ########################## Now run the tuning sims
  #

  if(load.results==F)
  {
  load(file = paste0(loc,"/Results/model_inputs_no_age_correction.Rdata"))

  for.tune.all.res <- NULL
  for.tune.summary <- NULL
  for(i in Stocks)
  {
    years <- years.tmp[[i]]
    prop.nat.mort <- pnm.tmp[[i]] 
    #rem <- rem.tmp[[i]] 
    mx <- mx.tmp[[i]] 
    N.end <- NE.tmp[[i]] 
    ages <- ages.tmp[[i]]
    vpa.abund <- vpa.tmp[[i]] 
    abund <- abund.tmp[[i]]
    fs <-fs.tmp[[i]]
    full.rem <- full.rem.tmp[[i]]
    #full.abund <- full.abund.tmp[[i]]
    #all.abund <- full.abund.tmp[[i]]
    #N.end <- vpa.abund[length(vpa.abund)]
    #Pick your step size
    ss.size <- 0.005
    sensistive.stocks <- c("AFSC_GOA_Theragra_chalcogramma",
                            "DFO_2J3KL_Gadus_morhua",
                            "DFO_3Pn-4RS_Gadus_morhua",
                            "DFO_4T-4VN_Gadus_morhua",
                            "DFO_4T_Scomber_scombrus",
                            "ICES-AFWG_NEA1-2_Melanogrammus_aeglefinus",
                            "ICES-AFWG_NEA1-2_Gadus_morhua",
                            "ICES-HAWG_NS_Ammodytes_dubius",
                            "ICES-NWWG_FA5b_Pollachius_virens",
                            "ICES-NWWG_FAPL5b1_Gadus_morhua",
                            "ICES-WGBFAS_BS 22-32_Sprattus_sprattus",
                            "ICES-WGCSE_IS7a_Gadus_morhua",
                            "ICES-WGCSE_CS7a_Merlangius_merlangus",
                            "ICES-WGCSE_CS7e-k_Gadus_morhua",
                            "ICES-WGCSE_CS27.6a_Merlangius_merlangus",
                            "ICES-WGHANSA_SP8abd_Sardina _pilchardus", 
                            "ICES-WGNSSK_NS 4-7d,20_Gadus_morhua",
                            "ICES-WGNSSK_NS  4-6a-20_Melanogrammus_aeglefinus",
                            "ICES-WGNSSK_NS 4-7d_Merlangius_merlangus",
                            "ICES-WGNSSK_NS 4-3aN_Trisopterus_esmarkii",
                            "NEFSC-GARMIII_SNE- MA_Limanda_ferruginea",
                            "NEFSC-GARMIII_MA_Paralichthys_dentatus",
                            "NEFSC-GARMIII_5Y_Melanogrammus_Aeglefinus",
                            "NEFSC_NWA_Scomber _scombrus")

    if(i %in% sensistive.stocks) ss.size = ss.size/10

     N.start <- vpa.abund[1]
     print(i)
     # The fecundities are 
     fast.for.tune <- fast.tunes(years,
                             tuner=what.2.tune,
                             step.size = ss.size, 
                             abund.ts = vpa.abund,
                             ages=ages,
                             nm = -(log(1-prop.nat.mort)),
                             fecund = mx,
                             fm = fs,
                             abund.age = abund, 
                             catch.age = full.rem, 
                             N.init = N.start,
                             direction= 'forwards'
                             )

     for.tune.all.res[[i]] <- fast.for.tune
     for.tune.tmp <- data.frame(fast.for.tune$res,Stock = i)
     for.tune.summary[[i]] <- for.tune.tmp
  }

  saveRDS(for.tune.summary,paste0(loc,"/Results/forwards_fully_tuned_summary_no_age_correction_",what.2.tune,".Rds"))
  saveRDS(for.tune.all.res,paste0(loc,"/Results/forwards_fully_tuned_all_no_age_corection_res_",what.2.tune,".Rds"))
  # 
}# end if load results

```

```{r paper-other-metrics-and-metadata,echo=F,include=F,warning=F,message=F}

if(load.results==F)
{
  # We are now looking at everything, lifetime reproductive success

  for.tune.summary <- readRDS(paste0(loc,"/Results/forwards_fully_tuned_summary_no_age_correction_",what.2.tune,".Rds"))
  for.tune.all <- readRDS(paste0(loc,"/Results/forwards_fully_tuned_all_no_age_corection_res_",what.2.tune,".Rds"))

  load(file =  paste0(loc,"/Results/model_inputs_no_age_correction.Rdata"))
  # Get the stocks...
  Stocks <- names(for.tune.summary)
  n.stocks <- length(Stocks)

  ############ Meta data and clean up, could be it's own chunk really, but it's really fast, so meh...

  #  A quick way to get the other potentially interesting info for the stock...
  drop <- which(names(ASR.long) %in% c("Year","type","age","value"))
  meta <- NULL
  for(i in Stocks) 
  {
    meta[[i]] <- ASR.long[ASR.long$Stock == i,-drop][1,]
    meta[[i]]$age.min <- min(ages.tmp[[i]])
    meta[[i]]$age.max <- max(ages.tmp[[i]])
  }
  meta <- do.call('rbind',meta)

  # First making a meta data object and adding a bunch of columns to it... very messy!!
  meta$Species[meta$Species == ' harengus']  <- "harengus"
  meta$Species[meta$Species == "Aeglefinus"] <- "aeglefinus"
  meta$Species[meta$Species == "Solea"] <- "solea"
  meta$Family[meta$Family == "Gadide"] <- "Gadidae"
  meta$Genus[meta$Genus == " Hippoglossoides"] <- "Hippoglossoides"
  meta$G.short <- toupper(paste0(substr(meta$Genus,1,1),"."))
  meta$G.Species <- paste0(meta$G.short," ",meta$Species)
  # Simplified Management
  meta$Manage.short <- meta$Management
  meta$Manage.short[grep("TRAC",meta$Management)] <- "DFO-TRAC"
  meta$Manage.short[grep("DFO",meta$Management)] <- "DFO-TRAC"
  meta$Manage.short[grep("NEFSC",meta$Management)] <- "NEFSC"
  meta$Manage.short[grep("ICES",meta$Management)] <- "ICES"

  # The 3 pieces of the ocean
  meta$LME <- "NW Atlantic"
  meta$LME[meta$Management == "AFSC"] <- "Pacific (Alaska)"
  meta$LME[grepl("ICES",meta$Management)] <- "NE Atlantic"
  # An area typo
  meta$Area[grep("BSAI-females",meta$Area)] <- "BSAI-Females"
  meta$Area[grep("BASI-Females",meta$Area)] <- "BSAI-Females"
  # Some other Area summaries
  meta$Area.broad <- meta$Area
  meta$Area.long <- meta$Area
  meta$Area.narrow <- "" # Leave this blank if not needed

  meta$Area.broad[grep("NS",meta$Area)] <- "NS"
  meta$Area.long[grep("NS",meta$Area)] <- "North Sea"
  meta$Area.broad[grep("CS",meta$Area)] <- "CS"
  meta$Area.long[grep("CS",meta$Area)] <- "Celtic Sea"
  meta$Area.broad[grep("ROCK6b",meta$Area)] <- "CS"
  meta$Area.long[grep("ROCK6b",meta$Area)] <- "Celtic Sea"
  meta$Area.broad[grep("IS",meta$Area)] <- "IS"
  meta$Area.long[grep("IS",meta$Area)] <- "Irish Sea"
  meta$Area.broad[grep("BS",meta$Area)] <- "BS"
  meta$Area.long[grep("BS",meta$Area)] <- "Baltic Sea"
  meta$Area.broad[grep("NEA",meta$Area)] <- "NEA"
  meta$Area.long[grep("NEA",meta$Area)] <- "NE Arctic"
  meta$Area.broad[grep("B28.1",meta$Area)] <- "BS"
  meta$Area.long[grep("B28.1",meta$Area)] <- "Baltic Sea"
  meta$Area.broad[grep("SP8abd",meta$Area)] <- "BoB"
  meta$Area.long[grep("SP8abd",meta$Area)] <- "Bay of Biscay"
  meta$Area.broad[grep("FA",meta$Area)] <- "Ice-Faroe"
  meta$Area.broad[grep("ICE",meta$Area)] <- "Ice-Faroe"
  meta$Area.long[grep("FA",meta$Area)] <- "Faroes"
  meta$Area.long[grep("ICE",meta$Area)] <- "Iceland"

  meta$Area.broad[grep("GOM",meta$Area)] <- "NAFO 5-6"
  meta$Area.long[grep("GOM",meta$Area)] <- "NAFO 5-6"
  meta$Area.broad[grep("GB",meta$Area)] <- "NAFO 5-6"
  meta$Area.long[grep("GB",meta$Area)] <- "NAFO 5-6"
  meta$Area.broad[grep("5Y",meta$Area)] <- "NAFO 5-6"
  meta$Area.long[grep("5Y",meta$Area)] <- "NAFO 5-6"
  meta$Area.broad[grep("5Z",meta$Area)] <- "NAFO 5-6"
  meta$Area.long[grep("5Z",meta$Area)] <- "NAFO 5-6"
  meta$Area.broad[grep("USATL",meta$Area)] <- "NAFO 5-6" # The fishery is broader than just GOM, but it does also include it.
  meta$Area.long[grep("USATL",meta$Area)] <- "NAFO 5-6" # The fishery is broader than just GOM, but it does also include it.
  meta$Area.broad[grep("NWA",meta$Area)] <- "NAFO 5-6" # Mackerel... so that definition is a bit narrow....
  meta$Area.long[grep("NWA",meta$Area)] <- "NAFO 5-6" # Mackerel... so that definition is a bit narrow....
  #meta$Area1[grep("SNE- MA",meta$Area)] <- "GOM-GB" # YT in US, so no GB so that definition is a bit narrow....
  meta$Area.broad[grep("MA",meta$Area)] <- "NAFO 5-6" # YT  and summer flounder in US, so no GB so that definition is a bit narrow....
  meta$Area.long[grep("MA",meta$Area)] <- "NAFO 5-6" # YT  and summer flounder in US, so no GB so that definition is a bit narrow....
  meta$Area.broad[grep("BSAI",meta$Area)] <- "BSAI"
  meta$Area.broad[grep("4T",meta$Area)] <- "NAFO 2-4"
  meta$Area.long[grep("4T",meta$Area)] <- "NAFO 2-4"
  meta$Area.broad[grep("2J3",meta$Area)] <- "NAFO 2-4"
  meta$Area.long[grep("2J3",meta$Area)] <- "NAFO 2-4"
  meta$Area.broad[grep("3Pn-4RS",meta$Area)] <- "NAFO 2-4"
  meta$Area.long[grep("3Pn-4RS",meta$Area)] <- "NAFO 2-4"
  meta$Area.long[grep("GOA",meta$Area)] <- "Gulf of Alaska"
  # Lumping everything in Berring Sea and AI into both.
  meta$Area.long[grep("AI",meta$Area)] <- "Aleutian Islands"
  meta$Area.long[grep("ESB",meta$Area)] <- "Bering Sea & AI" 
  meta$Area.long[grep("BSAI",meta$Area)] <- "Bering Sea & AI"
  meta$Area[grep("BASI-Females",meta$Area)] <- "BSAI"
  meta$Area[grep("BSAI-Females",meta$Area)] <- "BSAI"
  meta$Area[grep("BSAI-females",meta$Area)] <- "BSAI"
  meta$Area.long[grep("BSAI-Females",meta$Area)] <- "Bering Sea & AI"
  meta$Area.narrow[meta$Stock == "ICES-NWWG_FA5b_Pollachius_virens"] <- "5b"
  meta$Area.narrow[meta$Stock == "ICES-NWWG_FA5a_Pollachius_virens"] <- "5a"
  meta$Area.narrow[meta$Stock == "ICES-WGNSSK_NS4 _Solea_solea"] <- "4"
  meta$Area.narrow[meta$Stock == "ICES-WGNSSK_NS 7d._Solea_solea"] <- "7d"
  meta$Area.narrow[meta$Stock == "ICES-WGBFAS_NS IIIa 22-24_Solea_Solea"] <- "22-24"
  meta$Area.narrow[meta$Stock == "ICES-WGNSSK_NS 7d_Pleuronectes_platessa"] <- "7d"
  meta$Area.narrow[meta$Stock == "ICES-WGNSSK_NS 4,20_Pleuronectes_platessa"] <- "4 20"
  meta$Area.narrow[meta$Stock == "ICES-WGCSE_IS7e_Pleuronectes_platessa"] <- "7e"
  meta$Area.narrow[meta$Stock == "ICES-WGCSE_IS27.7a_Pleuronectes_platessa"] <- "7a"
  meta$Area.narrow[meta$Stock == "ICES-NWWG_FA5b_Melanogrammus_aeglefinus"] <- "5b"
  meta$Area.narrow[meta$Stock == "ICES-NWWG_ICE5a_Melanogrammus_aeglefinus"] <- "5a"
  meta$Area.narrow[meta$Stock == "ICES-WGBFAS_B28.1_Clupea_harengus"] <- "28"
  meta$Area.narrow[meta$Stock == "ICES-WGBFAS_BS 25-29-32_Clupea_harengus" ] <- "25-32"
  meta$Area.narrow[meta$Stock == "ICES-HAWG_WBS 22-24_Clupea_harengus"] <- "22-24"
  meta$Area.narrow[meta$Stock == "DFO_4T-Fall_Clupea_harengus"] <- "Fa-SGSL"
  meta$Area.narrow[meta$Stock == "DFO_4T-Spring_Clupea_harengus"] <- "Sp-SGSL"
  meta$Area.narrow[meta$Stock == "ICES-WGCSE_CS27.6a_Merlangius_merlangus"] <- "27.6a"
  meta$Area.narrow[meta$Stock == "ICES-WGCSE_CS7a_Merlangius_merlangus"] <- "7a"
  meta$Area.narrow[meta$Stock == "ICES-NWWG_ICE5a_Gadus_morhua"] <- "5a"
  meta$Area.narrow[meta$Stock == "ICES-NWWG_FAPL5b1_Gadus_morhua" ] <- "5b"
  meta$Area.narrow[meta$Stock == "ICES-WGNSSK_NS 4-7d,20_Gadus_morhua"] <- "4-7 20"
  meta$Area.narrow[meta$Stock == "ICES-WGCSE_NS6a_Gadus_morhua"] <- "6a"
  meta$Area.narrow[meta$Stock == "DFO_2J3KL_Gadus_morhua"] <- "NFLD"
  meta$Area.narrow[meta$Stock == "DFO_3Pn-4RS_Gadus_morhua"] <- "NGSL"
  meta$Area.narrow[meta$Stock == "NEFSC-CERT-TRAC_GB _Gadus_morhua"] <- "EGB"
  meta$Area.narrow[meta$Stock == "ICES-WGCSE_ROCK6b_Melanogrammus_aeglefinus"] <- "RA"
  meta$Area.narrow[meta$Stock == "DFO_4T-4VN_Gadus_morhua"] <- "SGSL"
  meta$Area.narrow[meta$Stock == "DFO_4T_Hippoglossoides _platessoides"] <- "SGSL"
  meta$Area.narrow[meta$Stock == "DFO_4T_Scomber_scombrus"] <- "SGSL"
  meta$Area.narrow[meta$Stock == "NEFSC-SAW_GOM_Gadus_morhua"] <- "GOM"
  meta$Area.narrow[meta$Stock == "NEFSC-GARMIII_GB_Gadus_morhua"] <- "GB"
  meta$Area.narrow[meta$Stock == "NEFSC-GARMIII_5Y_Melanogrammus_Aeglefinus"] <- "GOM"
  meta$Area.narrow[meta$Stock == "CERT-TRAC_GB_Melanogrammus_Aeglefinus"] <- "GB"

  meta$Stock.short <- paste0(meta$Area.broad," ",meta$Area.narrow," ",meta$G.short," ",meta$Species)
  meta$Stock.no.species <- paste0(meta$Area.long," ",meta$Area.narrow)

  for.tune.stocks <- NULL
  fta <- NULL

  # Going to rename these female stocks.
  fem.names <- c("AFSC_BASI-Females_Pleuronectes_quadrituberculatus",
                    "AFSC_EEBSAI-Females_Lepidopsetta_polyxystra",
                    "AFSC_BSAI-females_Atheresthes_stomais")
  fem.rename <- c("AFSC_BSAI_Pleuronectes_quadrituberculatus",
                    "AFSC_EEBSAI_Lepidopsetta_polyxystra",
                    "AFSC_BSAI_Atheresthes_stomais")
  # Going to drop these stocks
  dont <- paste(c("Mles","Males","COASTNOR"),collapse = "|")

  meta.sub <- NULL
  for(i in Stocks) 
  {
    org.nam <- i
    tmp <- for.tune.summary[[org.nam]]
    new.stock.nam <- org.nam
    # Don't bother for these stocks...
    if(!grepl(dont,tmp$Stock[1]))
    {
      # Grab the metadata and add in some new useful things...
      meta.sub[[i]] <- meta[meta$Stock == org.nam,]

      # Rename the ladies
      if(tmp$Stock[1] == fem.names[1]) new.stock.nam <- tmp$Stock <- fem.rename[1]
      if(tmp$Stock[1] == fem.names[2]) new.stock.nam <- tmp$Stock <- fem.rename[2]
      if(tmp$Stock[1] == fem.names[3]) new.stock.nam <- tmp$Stock <- fem.rename[3]

      meta.sub[[i]]$Stock <- new.stock.nam
      tmp <- data.frame(tmp,meta.sub[[i]])
      # For tunes all
      fta[[new.stock.nam]] <- for.tune.all[[org.nam]]
      fta[[new.stock.nam]]$meta <- meta.sub[[i]]
      for.tune.stocks[[new.stock.nam]] <- tmp
    }# End skip stocks...
  }
  # Now unpack them...
  for.tunes <- do.call("rbind",for.tune.stocks)
  for.tune.all <- fta
  meta.sub <- do.call("rbind",meta.sub)

  ######################### End Meta data and clean up #####################

  # Now some summaries...

  # Reset on the stocks...
  Stocks <- sort(names(for.tune.all))
  n.stocks <- length(Stocks)

  # Doubling time, no fishing
  dt.res <- NULL
  # These are both set in the index file now, go there!!
  n.dt.sims <- n.dt.sims
  n.sim.years <- n.sim.years
  for(i in Stocks)
  {
    str.time <- Sys.time()
    try <- for.tune.all[[i]]
    meta.tmp <- try$meta
    age.4.sim <- meta.tmp$age.min:meta.tmp$age.max
    double.time <- NA
    pop <- 1000
    tmp.res <- NULL
  for(s in 1:n.dt.sims)
  {
    tmp <- rep(0,n.sim.years)
    print(paste0("Sim # ", s))

    for(y in 2:n.sim.years)
    {

      # Get our sample for this run
      f.pick <- sample(nrow(try$fecund.opt),size=1)
      nm.pick <- sample(nrow(try$nm.opt),size=1)
      fec.samp <- try$fecund.opt[f.pick,]
      # This is instantaneous rate, which is what the Lotka function is expected (it is set up expecting instantaneous)
      nm.samp <- try$nm.opt[nm.pick,] 
      # now put these into Euler-lotka to get an r estimate
      res <- simple.lotka.r(mort = nm.samp,fecund=fec.samp,ages=age.4.sim)
      r.val <- res$res$r
      # If r is > 5 (lambda = exp(5)) it generally means the simple.lotka didn't converge (bad combo of fec and nm)
      # So we'll need to resample them until it is < 5
      while(r.val >= 5)
      {
        print("Resampling due to bat shit r value")
        f.pick <- sample(nrow(try$fecund.opt),size=1)
        nm.pick <- sample(nrow(try$nm.opt),size=1)
        fec.samp <- try$fecund.opt[f.pick,]
        # This is instantaneous rate, which is what the Lotka function is expected (it is set up expecting instantaneous)
        nm.samp <- try$nm.opt[nm.pick,] 
        # now put these into Euler-lotka to get an r estimate
        res <- simple.lotka.r(mort = nm.samp,fecund=fec.samp,ages=age.4.sim)
        r.val <- res$res$r
      } # end while loop
      pop[y] <- exp(r.val)*pop[y-1]
      if(pop[y] >= 2*pop[1] & tmp[y] == 0) tmp[y:n.sim.years] <- 1
    } # end the years loop
    tmp.res[[s]] <- data.frame(year =1:n.sim.years,doubled = tmp,sim=s)

    } # end sim loop
    res.tmp <- do.call("rbind",tmp.res)

  dt.res[[i]] <- data.frame(res.tmp,meta.tmp)
  count <- 0
  timer <- round(difftime(Sys.time(),str.time,units='mins'),digits=1)
  print(paste0("Stock ",i," time =", timer," minutes: Stock # ",count+1))
  } # end Stocks loop

  dub.time <- do.call("rbind",dt.res)

  dub.time.summary <- dub.time |> collapse::fgroup_by(Stock,year)  |> collapse::fsummarise(prop = sum(doubled)/n.dt.sims)
  dub.time.summary <- left_join(dub.time.summary,meta.sub,by="Stock")

  # Now to get generation lengths + reproductive rates....
  # First I need to get the fecundity and natural mortalities by cohort instead of year.
  res.fin <- NULL
  res.year.fin <- NULL
  res.lambda.fin <- NULL
  for(i in Stocks) # 
  {
    tst <- for.tune.all[[i]]
    meta.tmp <- tst$meta
    # Fishing mortality
    fm.org.tmp <- tst$fm.org
    fm.tmp <- tst$fm.opt
    # fecundity
    fec.tmp <- tst$fecund.opt
    fec.org.tmp <- tst$fecund.org
    # Natural mortality
    m.org.tmp <- exp(-tst$nm.org) # This is survivorship
    m.tmp <- exp(-tst$nm.opt) # This is survivorship
    # Total mortality
    z.tmp <- exp(-tst$z.opt) # This is survivorship
    z.org.tmp <- exp(-tst$nm.org -fm.org.tmp) # This is survivorship

    ages <- meta.tmp$age.min:meta.tmp$age.max
    n.ages <- length(ages)
    years <- tst$res$year
    n.years <- length(years)
    # Now I need to grab the fecundities and nat morts by cohort
    if(n.years > n.ages)
    {
    res.cohort <- NULL
    res.year <- NULL
    res.lambda <- NULL
    for(y in 1:(n.years-n.ages+1))
    {
      count = 0
      fs <- NA
      # The lx is no fishing survivorship, zx I'm calling the survivorship including fishing
      lx <- lx.org <- lx.year <- zx <- zx.org <- zx.year <- 1
      fs.org <- NA

      for(a in 1:n.ages)
      {
      fs[a] <- fec.tmp[y+count,a]  
      fs.org[a] <- fec.org.tmp[y+count,a]  
      if(a > 1) 
      {
        # No fishing
        lx[a] <- lx[a-1] * m.tmp[y+count,a]  
        lx.org[a] <- lx.org[a-1]* m.org.tmp[y+count,a]  
        lx.year[a] <- lx.year[a-1] * m.tmp[y,a]
        # Including fishing
        zx[a] <- zx[a-1] * z.tmp[y+count,a]  
        zx.org[a] <- zx.org[a-1]* z.org.tmp[y+count,a]  
        zx.year[a] <- zx.year[a-1] * z.tmp[y,a]
      }  # end the a > 1 if
          count <- count + 1
      }# end the ages loop
          res.cohort[[y]] <- data.frame(mx.opt = fs,lx.opt = lx,mx.org = fs.org,
                                 mx.lx.opt = fs*lx,mx.lx.org = fs.org*lx.org,
                                 x.mx.lx.opt = fs*lx*ages,x.mx.lx.org = fs.org*lx.org*ages,
                                 zx.opt = zx,
                                 mx.zx.opt = fs*zx,mx.zx.org = fs.org*zx.org,
                                 x.mx.zx.opt = fs*zx*ages,x.mx.zx.org = fs.org*zx.org*ages,
                                 cohort=years[y],age = ages,meta.tmp)

          res.year[[y]] <- data.frame(mx.opt = unlist(fec.tmp[y,]),lx.opt = lx.year,
                                 mx.lx.opt = unlist(fec.tmp[y,])*lx.year,
                                 x.mx.lx.opt = unlist(fec.tmp[y,])*lx.year*ages,
                                 zx.opt = zx.year,
                                 mx.zx.opt = unlist(fec.tmp[y,])*lx.year,
                                 x.mx.zx.opt = unlist(fec.tmp[y,])*lx.year*ages,
                                 age = ages,year=years[y],meta.tmp)

    } # end the years loop

  # Do a years loop to get lambda with no fishing for each year, rather than by cohort...
    for(yy in 1:(n.years-1))
    {
      nm <- as.numeric(tst$nm.opt[yy,])
      z <- as.numeric(tst$z.opt[yy,])
      fec <- as.numeric(tst$fecund.opt[yy,])
      nm.init <- as.numeric(tst$nm.org[yy,])
      fec.init <- as.numeric(tst$fecund.org[yy,])
      z.init <- as.numeric(tst$z.org[yy,])
      lam.no.fish <- exp(simple.lotka.r(mort = nm,fecund = fec,ages=ages)$res$r)
      lam.fish <- exp(simple.lotka.r(mort = z,fecund = fec,ages=ages)$res$r)
      lam.org.no.fish <- exp(simple.lotka.r(mort = nm.init,fecund = fec.init,ages=ages)$res$r)
      lam.org.fish <- exp(simple.lotka.r(mort =  z.init,fecund = fec.init,ages=ages)$res$r)

      # Now summarize these
      res.lambda[[yy]] <-  data.frame(lam.no.fish = lam.no.fish,lam.fish = lam.fish,
                                      lam.org.no.fish = lam.org.no.fish,
                                      lam.org.fish = lam.org.fish,
                                      year=years[yy],meta.tmp)
    } # End the yy loop to get the lambdas...
  # unpack it...
  res.fin[[i]] <- do.call('rbind',res.cohort)
  res.year.fin[[i]] <- do.call('rbind',res.year)
  res.lambda.fin[[i]] <-  do.call('rbind',res.lambda)
  }  # end if to make sure we have enough data to get the cohort estimates.

  }# end the stock loop

  res.cohort.final <- do.call('rbind',res.fin)
  res.year.final <- do.call('rbind',res.year.fin)
  res.lambda.final <- do.call('rbind',res.lambda.fin)

  pop.dam <- res.cohort.final |> collapse::fgroup_by(cohort,Stock) |> collapse::fsummarise(gen.len.opt = sum(x.mx.lx.opt)/sum(mx.lx.opt),
                                                                                   gen.len.org = sum(x.mx.lx.org)/sum(mx.lx.org),
                                                                                   R0.opt = sum(mx.lx.opt),R0.org = sum(mx.lx.org))
  pop.dam <- left_join(pop.dam,meta.sub,by="Stock")

  dub.time.summary <- data.frame(dub.time.summary)
  pop.dam <- data.frame(pop.dam)
  # saveRDS(pop.dam,paste0("D:/Github/ICM/Results/R0_and_gen_length_",what.2.tune,".Rds"))
  # saveRDS(dub.time.summary,paste0("D:/Github/ICM/Results/doubling_time_",what.2.tune,".Rds"))
  # saveRDS(for.tunes,paste0("D:/Github/ICM/Results/forward_tunes_for_figures_",what.2.tune,".Rds"))
  save(dub.time.summary,pop.dam,res.year.final,res.cohort.final,for.tunes,dub.time,for.tune.all,res.lambda.final,meta,meta.sub,
       file=paste0(loc,"/Results/all_cleaned_forward_tune_summaries_no_age_corection_",what.2.tune,".Rdata"))

  write.csv(meta.sub, paste0(loc,"/Data/metadata_no_age_corection.csv"))
} # end if load results
```

```{r clean-up-and-stats,include=F,echo=F,message=F,warning=F}
load(file=paste0(loc,"/Results/all_cleaned_forward_tune_summaries_no_age_corection_",what.2.tune,".Rdata"))
for.tune.summary <- readRDS(paste0(loc,"/Results/forwards_fully_tuned_summary_no_age_correction_",what.2.tune,".Rds"))

Stocks <- unique(res.lambda.final$Stock)
res.lambda.fin <- data.frame(res.lambda.final)
six.species <- c('morhua','harengus','aeglefinus','virens','solea','platessa')
pop.dam.6 <- pop.dam |> collapse::fsubset(Species %in% six.species)
for.tunes.6 <- for.tunes |> collapse::fsubset(Species %in% six.species)
dub.time.6 <- dub.time.summary |> collapse::fsubset(Species %in% six.species)
res.lambda.fin$decade <- floor(res.lambda.fin$year/10)*10
res.lambda.fin <- res.lambda.fin |> collapse::fgroup_by(Stock,decade) |> collapse::fmutate(mn.lam = mean(lam.fish,na.rm=T),
                                                                                 med.lam = median(lam.fish,na.rm=T),
                                                                                 sd.lam = sd(lam.fish,na.rm=T))
res.lambda.fin <- res.lambda.fin |> collapse::fgroup_by(Species,decade) |> collapse::fmutate(lm.spec.mn = mean(mn.lam,na.rm=T))              
res.lambda.fin <- res.lambda.fin |> collapse::fgroup_by(Stock) |> 
                                                          collapse::fmutate(
                                                          lam.no.fish.mn = mean(lam.no.fish,na.rm=T),
                                                          lam.fish.mn = mean(lam.fish,na.rm=T),
                                                          lam.org.no.fish.mn = mean(lam.org.no.fish,na.rm=T),
                                                          lam.org.fish.mn = mean(lam.org.fish,na.rm=T),
                                                          lam.no.fish.med = median(lam.no.fish,na.rm=T),
                                                          lam.fish.med = median(lam.fish,na.rm=T),
                                                          lam.org.no.fish.med = median(lam.org.no.fish,na.rm=T),
                                                          lam.org.fish.med = median(lam.org.fish,na.rm=T),
                                                          lam.no.fish.sd = sd(lam.no.fish,na.rm=T),
                                                          lam.fish.sd = sd(lam.fish,na.rm=T),
                                                          lam.org.no.fish.sd = sd(lam.org.no.fish,na.rm=T),
                                                          lam.org.fish.sd = sd(lam.org.fish,na.rm=T)
                                                          ) |> data.frame()
# Get the order right for the stock no species for the boxplots to come later
sns.order <- c("Baltic Sea ", "Baltic Sea 22-24", "Baltic Sea 25-32", "Baltic Sea 28", 
               "North Sea ", "North Sea 4", "North Sea 22-24", "North Sea 4-7 20", "North Sea 6a","North Sea 4 20", "North Sea 7d",     
               "Celtic Sea ","Celtic Sea RA", "Celtic Sea 7a", "Celtic Sea 27.6a", 
               "Irish Sea ","Irish Sea 7a", "Irish Sea 7e","Bay of Biscay ", "NE Arctic ", 
               "Faroes 5a", "Faroes 5b",  "Iceland ", "Iceland 5a",
                "NAFO 2-4 Fa-SGSL", "NAFO 2-4 Sp-SGSL", "NAFO 2-4 SGSL", "NAFO 2-4 NGSL", "NAFO 2-4 NFLD",  
                "NAFO 5-6 ", "NAFO 5-6 GB", "NAFO 5-6 GOM", "NAFO 5-6 EGB",
                "Bering Sea & AI ", "Gulf of Alaska ", "Aleutian Islands ")

res.lambda.fin$Stock.no.species <- factor(res.lambda.fin$Stock.no.species,levels = sns.order,ordered = T)
res.lambda.final.6 <- res.lambda.fin |> collapse::fsubset(Species %in% six.species)
# Sort the data into a nice order...
res.lambda.fin <- res.lambda.fin |> collapse::roworder(Species,Area.broad) |>
                                    collapse::fmutate(Stock.no.species = forcats::fct_inorder(factor(Stock.no.species, ordered=TRUE)))
# Making the Aleutian Islands their own Area long...
res.lambda.fin$Area.long[res.lambda.fin$Area == "AI"] <- "Aleutian Islands"
res.lambda.fin$Area.long <- factor(res.lambda.fin$Area.long,levels = c("Baltic Sea", "North Sea","Celtic Sea",
                                                                       "Irish Sea","Bay of Biscay","NE Arctic",
                                                                       "Faroes","Iceland","NAFO 2-4","NAFO 5-6","Gulf of Alaska","Bering Sea & AI","Aleutian Islands"),ordered = T)

res.lambda.final.6 <- res.lambda.final.6 |> collapse::roworder(Species,Area.broad) |>
                                            collapse::fmutate(Stock.no.species = forcats::fct_inorder(factor(Stock.no.species, ordered=TRUE)))
dub.time <- dub.time |> collapse::roworder(LME) |>
                        collapse::fmutate(Stock.no.species = forcats::fct_inorder(factor(Stock.no.species, ordered=TRUE)))

dub.time.summary <- dub.time.summary |> collapse::roworder(LME,Order,Species,Area.broad) |>
                                        collapse::fmutate(Stock.short = forcats::fct_inorder(factor(Stock.short, ordered=TRUE)))

dub.time.6 <- dub.time.6 |> collapse::roworder(LME) |>
                             collapse::fmutate(Stock.no.species = forcats::fct_inorder(factor(Stock.no.species, ordered=TRUE)))
pop.dam <- pop.dam |> collapse::roworder(Species,LME) |>
                       collapse::fmutate(Stock = forcats::fct_inorder(factor(Stock, ordered=TRUE)))
pop.dam.6 <- pop.dam.6 |> collapse::roworder(LME,Area.broad) |>
                       collapse::fmutate(Stock.no.species = forcats::fct_inorder(factor(Stock.no.species, ordered=TRUE)))
pop.dam$Area.long[pop.dam$Area == "AI"] <- "Aleutian Islands"
pop.dam$Area.long <- factor(pop.dam$Area.long,levels = c("Baltic Sea", "North Sea","Celtic Sea",
                                                         "Irish Sea","Bay of Biscay","NE Arctic",
                                                         "Faroes","Iceland","NAFO 2-4","NAFO 5-6","Gulf of Alaska","Bering Sea & AI","Aleutian Islands"),ordered = T)
# I'd also like a pop.dam including fishing
pop.dam.fish <- res.cohort.final |> collapse::fgroup_by(cohort,Stock) |> collapse::fsummarise(gen.len.opt = sum(x.mx.zx.opt)/sum(mx.zx.opt),
                                                                                   gen.len.org = sum(x.mx.zx.org)/sum(mx.zx.org),
                                                                                   R0.opt = sum(mx.zx.opt),R0.org = sum(mx.zx.org))
pop.dam.fish <- left_join(pop.dam.fish,meta.sub,by="Stock")
pop.dam.fish$Area.long[pop.dam.fish$Area == "AI"] <- "Aleutian Islands"
pop.dam.fish$Area.long <- factor(pop.dam.fish$Area.long,levels = c("Baltic Sea", "North Sea","Celtic Sea",
                                                         "Irish Sea","Bay of Biscay","NE Arctic",
                                                         "Faroes","Iceland","NAFO 2-4","NAFO 5-6","Gulf of Alaska","Bering Sea & AI","Aleutian Islands"),ordered = T)
pop.dam <- data.frame(pop.dam)

pop.dam.fish <- pop.dam.fish |> collapse::roworder(Species,LME) |>
                                collapse::fmutate(Stock = forcats::fct_inorder(factor(Stock, ordered=TRUE)))
pop.dam.fish.6 <- pop.dam.fish |> collapse::fsubset(Species %in% six.species)

pop.dam.fish.6 <- pop.dam.fish.6 |> collapse::roworder(LME,Area.broad) |>
                                    collapse::fmutate(Stock.no.species = forcats::fct_inorder(factor(Stock.no.species, ordered=TRUE)))

# Put the population dynamics stuff together.
pop.dam$group <- "no fish"
pop.dam.fish$group <- 'fish'
pop.dam.all <- rbind(pop.dam,pop.dam.fish)

pop.dam.6$Stock.no.species <- factor(pop.dam.6$Stock.no.species,levels = sns.order,ordered = T)
pop.dam$Stock.no.species <- factor(pop.dam$Stock.no.species,levels = sns.order,ordered = T)
pop.dam.fish.6$Stock.no.species <- factor(pop.dam.fish.6$Stock.no.species,levels = sns.order,ordered = T)
pop.dam.fish$Stock.no.species <- factor(pop.dam.fish$Stock.no.species,levels = sns.order,ordered = T)
pop.dam.all$Stock.no.species <- factor(pop.dam.all$Stock.no.species,levels = sns.order,ordered = T)
# Drop the periods....
pop.dam.all.6 <- pop.dam.all |> collapse::fsubset(!Species %in% six.species)
pop.dam.all$G.Species <- gsub("\\.","",pop.dam.all$G.Species)
pop.dam.all$G.Species.short <- substr(pop.dam.all$G.Species,1,5)
pop.dam.all$G.Species <- gsub("\\.","",pop.dam.all$G.Species)
pop.dam.all.6 <- pop.dam.all |> collapse::fsubset(Species %in% six.species)
pop.dam.all.others <- pop.dam.all |> collapse::fsubset(!Species %in% six.species)

# Reorder the doubling time, either by Species and LME, or the other way around

dub.time.summary$Order.short <- "Gadi"
dub.time.summary$Order.short[dub.time.summary$Order == "Pleuronectiformes"] <- "Plero"
dub.time.summary$Order.short[dub.time.summary$Order == "Clupeiformes"] <- "Clupei"
dub.time.summary$Order.short[dub.time.summary$Order == "Scorpaeniformes "] <- "Scor"
dub.time.summary$Order.short[dub.time.summary$Order == "Scombriformes"] <- "Scomb"
dub.time.summary$Order.short[dub.time.summary$Order == "Perciformes"] <- "Perci"
# First need to get the abundance and lambdas with no fishing into one object.
ft.tmp <- for.tunes |> collapse::fselect(year,Stock,vpa.abund,est.abund,removals)
dd.dat <- left_join(res.lambda.fin,ft.tmp,by=c('year','Stock'))
# Now I need to get a relative abundance for each stock...
dd.dat <- dd.dat |> collapse::fgroup_by(Stock) |> collapse::fmutate(max.abund = max(est.abund,na.rm=T))
dd.dat <- data.frame(dd.dat)
dd.dat$prop.abund <- dd.dat$est.abund/dd.dat$max.abund
# Drop nas....
drops <- -which(is.na(dd.dat$prop.abund))
dd.dat <- dd.dat[drops,]
#lump the data into bins
dd.dat$prop.bin <- "< 10%"
dd.dat$prop.bin[dd.dat$prop.abund > 0.1 & dd.dat$prop.abund <= 0.2] <- "10-20%"
dd.dat$prop.bin[dd.dat$prop.abund > 0.2 & dd.dat$prop.abund <= 0.3] <- "20-30%"
dd.dat$prop.bin[dd.dat$prop.abund > 0.3 & dd.dat$prop.abund <= 0.4] <- "30-40%"
dd.dat$prop.bin[dd.dat$prop.abund > 0.4 & dd.dat$prop.abund <= 0.6] <- "40-60%"
dd.dat$prop.bin[dd.dat$prop.abund > 0.6 & dd.dat$prop.abund <= 0.8] <- "60-80%"
dd.dat$prop.bin[dd.dat$prop.abund > 0.8 & dd.dat$prop.abund <= 1] <- "80-100%"
dd.dat$prop.bin <- factor(dd.dat$prop.bin,levels=c("< 10%","10-20%","20-30%","30-40%","40-60%","60-80%","80-100%"))

# Subset to the big six
dd.dat.six <- dd.dat  |>collapse::fsubset(Species %in% six.species)

# This is for the new lambda figures with fishing and no fishing together

res.lambda.fin.long <- res.lambda.fin |> tidyr::pivot_longer(cols = c(lam.no.fish,lam.fish),names_to = 'group',values_to = 'lambda')
res.lambda.fin.long$Stock.no.species <- factor(res.lambda.fin.long$Stock.no.species,levels = sns.order,ordered = T)
res.lambda.fin.long$group <- substr(res.lambda.fin.long$group,5,20)
res.lambda.fin.long$group <- gsub("\\."," ",res.lambda.fin.long$group)
res.lambda.fin.long.others <- res.lambda.fin.long |> collapse::fsubset(!Species %in% six.species)
res.lambda.fin.long.others$G.Species <- gsub("\\.","",res.lambda.fin.long.others$G.Species)
res.lambda.fin.long.others$G.Species.short <- substr(res.lambda.fin.long.others$G.Species,1,5)
res.lambda.fin.long.6 <- res.lambda.fin.long |> collapse::fsubset(Species %in% six.species)
res.lambda.fin.long.6$G.Species <- gsub("\\.","",res.lambda.fin.long.6$G.Species)

# GAM FOR DENSITY DEPENENCE# GAM FOR DENSITY DEPENENCE# GAM FOR DENSITY DEPENENCE# GAM FOR DENSITY DEPENENCE# GAM FOR DENSITY DEPENENCE
# here is the GAM for no fishing with subsequent processing.
mod.dd.stock.gam.nf <- gam(log(lam.no.fish) ~   s(prop.abund,k=4) + s(prop.abund,as.factor(Stock.short),bs='fs',k=3) ,data=dd.dat)

# This extracts the GAM fit so we can custom plot it on our own figure
tmp <- plot(mod.dd.stock.gam.nf, pages = 0) # Leave a space after this so the code runs (it needs a return as this thing makes a two panel figure)

no.stock.gam.nf.dat <- data.frame(fit = tmp[[1]]$fit + coefficients(mod.dd.stock.gam.nf)[1],prop.abund = tmp[[1]]$x,se = tmp[[1]]$se,Scenario = "F = 0")
no.stock.gam.nf.dat$LCI <- no.stock.gam.nf.dat$fit - 1.96*no.stock.gam.nf.dat$se
no.stock.gam.nf.dat$UCI <- no.stock.gam.nf.dat$fit + 1.96*no.stock.gam.nf.dat$se

#summary(mod.dd.stock.gam.nf)
#Residuals, definitely still issues here, maybe a bit better than some others, maybe...
#plot(residuals(mod.dd.stock.gam.nf) ~ fitted(mod.dd.stock.gam.nf))
# Normality = no
#qqnorm(residuals(mod.dd.stock.gam.nf))
#qqline(residuals(mod.dd.stock.gam.nf))

pred.dd.stock.gam.nf <- expand.grid(prop.abund = seq(0, 1, length.out = 250),
                                         Stock.short=c(unique(dd.dat$Stock.short)))
tmp <- predict(mod.dd.stock.gam.nf, pred.dd.stock.gam.nf, type = 'response', se.fit = TRUE) 
# Another fun trick to extract the overall smooth fit, but didn't end up using this method.
# sms <- smooths(mod.dd.stock.gam.nf)
# excl <- grepl("as.factor",sms)
# rm <- sms[excl]
# tmp.ex <- predict(mod.dd.stock.gam.nf, newdata =pred.dd.stock.gam.nf, type = 'response', se.fit = TRUE,exclude =rm) 
# Get the fit by stock
pred.dd.stock.gam.nf$lam.no.fish   <- exp(tmp$fit)
pred.dd.stock.gam.nf$lower <- exp(tmp$fit - 1.96 * tmp$se.fit)
pred.dd.stock.gam.nf$upper <- exp(tmp$fit + 1.96 * tmp$se.fit)

# Thin the data so I can add points to the figure
tmp <- NULL
for(i in 1:(77)) 
{
  if(i==1) tmp[[i]] <- seq(10,230,length.out=5)
  if(i>1) tmp[[i]] <- seq(10,230,length.out=5)+250*(i-1)
}
thinner.nf <- do.call('c',tmp)
thin.pred.nf <- pred.dd.stock.gam.nf[thinner.nf,]
# Now I want to get rid of extrapolation from this prediction data.frame, I think I need to loop it...
# 
short.stocks <- unique(dd.dat$Stock.short)
n.stocks <- length(short.stocks)
pred.out <- NULL
thin.out <- NULL
for(i in short.stocks)
{
  # Get the right stocks
  tmpy <- dd.dat |> collapse::fsubset(Stock.short == i)
  tmp.pred <- pred.dd.stock.gam.nf |> collapse::fsubset(Stock.short == i)
  tmp.thin <- thin.pred.nf |> collapse::fsubset(Stock.short == i)
  # I dont want to predict below the minimum observed abundance
  mins <- min(tmpy$prop.abund,na.rm=T)
  pred.out[[i]] <- tmp.pred |> collapse::fsubset(prop.abund >= mins)
  thin.out[[i]] <- tmp.thin |> collapse::fsubset(prop.abund >= mins)
}

pred.dd.stock.gam.ss.nf <- do.call('rbind',pred.out)
thin.pred.nf <- do.call('rbind',thin.out)
# Combine pred stuff with the meta data for fun plots
pred.dd.stock.gam.ss.nf <- left_join(pred.dd.stock.gam.ss.nf,meta.sub,by="Stock.short")
pred.dd.stock.gam.ss.nf <- data.frame(pred.dd.stock.gam.ss.nf)
thin.pred.nf <- left_join(thin.pred.nf,meta.sub,by="Stock.short")
# Here are the fits (sans SE cause messy + didn't put the overall relationship on here)
# By region
cols.gam <- rep(c("black","blue","orange",'firebrick2','lightgreen'),6)
pty.gam <- c(rep(0,5),rep(1,5),rep(2,5),rep(3,5),rep(4,5),rep(5,5))

# Now do the same thing with the data that includes the effect of fishing.

#mod.dd.stock.gam.fish <- gam(log(lam.fish) ~  as.factor(Stock.short) + s(prop.abund,by=as.factor(Stock.short),k=3),data=dd.dat)
mod.dd.stock.gam.fish <- gam(log(lam.fish) ~  s(prop.abund,k=4) + s(prop.abund,as.factor(Stock.short),bs='fs',k=3),data=dd.dat)

tmp <- plot(mod.dd.stock.gam.fish, pages = 0) # Leave a space after this so the code runs (it needs a return as this thing makes a two panel figure)

no.stock.gam.fish.dat <- data.frame(fit = tmp[[1]]$fit + coefficients(mod.dd.stock.gam.fish)[1],prop.abund = tmp[[1]]$x,se = tmp[[1]]$se,Scenario = "F > 0")
no.stock.gam.fish.dat$LCI <- no.stock.gam.fish.dat$fit - 1.96*no.stock.gam.fish.dat$se
no.stock.gam.fish.dat$UCI <- no.stock.gam.fish.dat$fit + 1.96*no.stock.gam.fish.dat$se

#summary(mod.dd.stock.gam.fish)
#Residuals, definetly still issues here, maybe a bit better than some others, maybe...
#plot(residuals(mod.dd.stock.gam.fish) ~ fitted(mod.dd.stock.gam.fish))
# Normality = no
#qqnorm(residuals(mod.dd.stock.gam.fish))
#qqline(residuals(mod.dd.stock.gam.fish))

pred.dd.stock.gam.fish <- expand.grid(prop.abund = seq(0, 1, length.out = 250),
                                         Stock.short=c(unique(dd.dat$Stock.short)))
tmp <- predict(mod.dd.stock.gam.fish, pred.dd.stock.gam.fish, type = 'response', se.fit = TRUE) 
pred.dd.stock.gam.fish$lam.fish   <- exp(tmp$fit)
pred.dd.stock.gam.fish$lower <- exp(tmp$fit - 1.96 * tmp$se.fit)
pred.dd.stock.gam.fish$upper <- exp(tmp$fit + 1.96 * tmp$se.fit)

# Thin the data so I can add points to the figure
tmp <- NULL
for(i in 1:(77)) 
{
  if(i==1) tmp[[i]] <- seq(10,230,length.out=5)
  if(i>1) tmp[[i]] <- seq(10,230,length.out=5)+250*(i-1)
}
thinner.fish <- do.call('c',tmp)
thin.pred.fish <- pred.dd.stock.gam.fish[thinner.fish,]
# Now I want to get rid of extrapolation from this prediction data.frame, I think I need to loop it...
# 
short.stocks <- unique(dd.dat$Stock.short)
pred.out <- NULL
thin.out <- NULL
for(i in short.stocks)
{
  # Get the right stocks
  tmpy <- dd.dat |> collapse::fsubset(Stock.short == i)
  tmp.pred <- pred.dd.stock.gam.fish |> collapse::fsubset(Stock.short == i)
  tmp.thin <- thin.pred.fish |> collapse::fsubset(Stock.short == i)
  # I dont want to predict below the minimum observed abundance
  mins <- min(tmpy$prop.abund,na.rm=T)
  pred.out[[i]] <- tmp.pred |> collapse::fsubset(prop.abund >= mins)
  thin.out[[i]] <- tmp.thin |> collapse::fsubset(prop.abund >= mins)
}

pred.dd.stock.gam.ss.fish <- do.call('rbind',pred.out)
thin.pred.fish <- do.call('rbind',thin.out)
# Combine pred stuff with the meta data for fun plots
pred.dd.stock.gam.ss.fish <- left_join(pred.dd.stock.gam.ss.fish,meta.sub,by="Stock.short")
pred.dd.stock.gam.ss.fish <- data.frame(pred.dd.stock.gam.ss.fish)

thin.pred.fish <- left_join(thin.pred.fish,meta.sub,by="Stock.short")

# Here are the fits (sans SE cause messy + didn't put the overall relationship on here)
# By region

# What proportion of the stocks had the lowest 'GAM' estimate happen when the stock was at it's maximum size.
low.gam <- data.frame(low.pa.fish = NA,diff.low.pa.fish = NA,low.pa.nf = NA,diff.low.pa.nf = NA,Stock = NA,Species = NA)
count <- 0
for(i in short.stocks) 
{
  count = count +1
  tmp <- pred.dd.stock.gam.ss.fish |> collapse::fsubset(Stock.short == i)
  tmp.nf <- pred.dd.stock.gam.ss.nf |> collapse::fsubset(Stock.short == i)
  low.pa <- data.frame(low.pa.fish = tmp$prop.abund[tmp$lam.fish == min(tmp$lam.fish)],
                       diff.low.pa.fish = tmp$lam.fish[tmp$lam.fish == min(tmp$lam.fish)] - tmp$lam.fish[tmp$prop.abund == max(tmp$prop.abund)],
                       low.pa.nf = tmp.nf$prop.abund[tmp.nf$lam.no.fish == min(tmp.nf$lam.no.fish)],
                       diff.low.pa.nf = tmp.nf$lam.no.fish[tmp.nf$lam.no.fish == min(tmp.nf$lam.no.fish)] - tmp.nf$lam.no.fish[tmp.nf$prop.abund == max(tmp.nf$prop.abund)],
                       Stock = i,Species = tmp$G.Species[1])
  low.gam[count,] <- low.pa
}

# Let's compare lambda at minimum, 0.4, and maximum and make a Table
comp.lam.diff <- NULL
comp.lam <- NULL
comp.all.nf <- NULL
comp.all.fish <- NULL
for(i in short.stocks) 
{
  tmp <- pred.dd.stock.gam.ss.fish |> collapse::fsubset(Stock.short == i)
  which.40 <- max(which(tmp$prop.abund <=0.4)) # Some stocks might not have a which 40.
  which.max <- which(tmp$prop.abund ==1)
  which.min <- which(tmp$prop.abund == min(tmp$prop.abund))
  tmp.sub <- tmp[c(which.min,which.40,which.max),]
  comp.all.fish[[i]] <- tmp.sub
  tmp.nf <- pred.dd.stock.gam.ss.nf |> collapse::fsubset(Stock.short == i)
  # I think I'm overkilling as this should be the same was the above witches....
  which.40.nf <- max(which(tmp.nf$prop.abund <=0.4))
  which.max.nf <- which(tmp.nf$prop.abund ==1)
  which.min.nf <- which(tmp.nf$prop.abund ==min(tmp.nf$prop.abund))
  tmp.nf.sub <- tmp.nf[c(which.min.nf,which.40.nf,which.max.nf),]
  comp.all.nf[[i]] <- tmp.nf.sub
  # Now we just want some summaries here I think...
  comp.lam.diff[[i]] <- data.frame(min.abund = min(tmp.sub$prop.abund,na.rm=T),
                              lam.fish.min.40 = tmp$lam.fish[which.min] - tmp$lam.fish[which.40],
                              lam.fish.min.max = tmp$lam.fish[which.min] - tmp$lam.fish[which.max],
                              lam.fish.40.max = tmp$lam.fish[which.40] - tmp$lam.fish[which.max],
                              lam.nf.min.40 = tmp.nf$lam.no.fish[which.min.nf] - tmp.nf$lam.no.fish[which.40.nf],
                              lam.nf.min.max = tmp.nf$lam.no.fish[which.min.nf] - tmp.nf$lam.no.fish[which.max.nf],
                              lam.nf.40.max = tmp.nf$lam.no.fish[which.40.nf] - tmp.nf$lam.no.fish[which.max.nf],
                              Stock = i,Species = tmp$G.Species[1])

    comp.lam[[i]] <- data.frame(min.abund = min(tmp.sub$prop.abund,na.rm=T),
                                lam.fish.min = tmp$lam.fish[which.min] ,
                                lam.fish.40 = tmp$lam.fish[which.40],
                                lam.fish.max = tmp$lam.fish[which.max],
                                lam.nf.min = tmp.nf$lam.no.fish[which.min.nf],
                                lam.nf.40 = tmp.nf$lam.no.fish[which.40.nf],
                                lam.nf.max = tmp.nf$lam.no.fish[which.max.nf],
                                Stock = i,Species = tmp$G.Species[1])
}

# A table for the estimates at 'interesting' values
comp.lam.res <- do.call('rbind',comp.lam)
# I want to turn this into a table....
comp.lam.res.tab <- comp.lam.res
# Round the numbers
comp.lam.res.tab$min.abund <- round(comp.lam.res.tab$min.abund,digits=2)
rounders <- c("lam.fish.min","lam.fish.40","lam.fish.max",
              "lam.nf.min","lam.nf.40","lam.nf.max")
comp.lam.res.tab[,rounders]      <- round(comp.lam.res.tab[,rounders] ,digits=3)
# Now give it pretty names
tab.names <- c("Minimum Abundance",
               "$\\lambda_{F>0}^{Min}$",
               "$\\lambda_{F>0}^{40}$",
               "$\\lambda_{F>0}^{Max}$",
               "$\\lambda_{F=0}^{Min}$",
               "$\\lambda_{F=0}^{40}$",
               "$\\lambda_{F=0}^{Max}$",
               "Stock","Species")
names(comp.lam.res.tab) <- tab.names

# Now do the same with the differences
comp.lam.diff.res <- do.call('rbind',comp.lam.diff)
# I want to turn this into a table....
comp.lam.diff.res.tab <- comp.lam.diff.res
# Round the numbers
comp.lam.diff.res.tab$min.abund <- round(comp.lam.diff.res.tab$min.abund,digits=2)
rounders <- c("lam.fish.min.40","lam.fish.min.max","lam.fish.40.max",
                    "lam.nf.min.40","lam.nf.min.max","lam.nf.40.max")
comp.lam.diff.res.tab[,rounders] <- round(comp.lam.diff.res.tab[,rounders] ,digits=3)
# Now give it pretty names
tab.names <- c("Minimum Abundance",
               "$\\lambda_{F>0}^{Min-40}$",
               "$\\lambda_{F>0}^{Min-Max}$",
               "$\\lambda_{F>0}^{40-Max}$",
               "$\\lambda_{F=0}^{Min-40}$",
               "$\\lambda_{F=0}^{Min-Max}$",
               "$\\lambda_{F=0}^{40-Max}$",
               "Stock","Species")
names(comp.lam.diff.res.tab) <- tab.names

# Used for the figure showing the number of stocks that have recovered by year x.
n.dub <- data.frame(years = 1:n.sim.years, count = NA)
for(i in 1:n.sim.years) n.dub$count[i] <- dub.time.summary |> collapse::fsubset(year==i & prop >= 0.75) |> nrow()

# Compare the population recovery with the variance, using 2010 because all stocks have a value at 2010... 
# I'm using the overall lambda sd, so the year doesn't actually matter other than to make sure all stocks are included here
# This is used for the recovery variance figure, is this data good?
# How many of the stocks had doubled by year 20, with 75% being my standard, which is a high standard
dub.p.high <- 0.75
dub.p.low <- 0.5
n.dub.10 <- dub.time.summary |> collapse::fsubset(year==10 & prop >= dub.p.high) |> nrow()
n.dub.20 <- dub.time.summary |> collapse::fsubset(year==20 & prop >= dub.p.high) |> nrow()
# Slow stocks, here we define having a low recovery potential has having < 50% of the simulatoins experience a doubling in size.
slow.20 <- dub.time.summary |> collapse::fsubset(year==20 & prop < dub.p.low) 
prop.rec.20 <- dub.time.summary |> collapse::fsubset(year==20) # the proportion that recovers by year 20
lambda.sds <- res.lambda.fin |> collapse::fsubset(year == 2010)
lam.sd.p.rec <- left_join(prop.rec.20,lambda.sds,by=names(prop.rec.20)[c(1,4:24)]) 
lam.sd.p.rec$prop.cat <- "Moderate"
lam.sd.p.rec$prop.cat[lam.sd.p.rec$prop < dub.p.low] <- "Low"
lam.sd.p.rec$prop.cat[lam.sd.p.rec$prop >= dub.p.high] <- "High"
lam.sd.p.rec$prop.cat <- factor(lam.sd.p.rec$prop.cat,levels=c("Low","Moderate","High"))

```

```{r paper-calcs,include=F,echo=F,message=F,warning=F}

# the lambda calcs
n.stock.dfo <- meta.sub |> collapse::fsubset(Manage.short == "DFO-TRAC") |> nrow()
n.stock.ices <- meta.sub |> collapse::fsubset(Manage.short == "ICES") |> nrow()
n.stock.noaa <- meta.sub |> collapse::fsubset(Manage.short %in% c("AFSC","NEFSC")) |> nrow()
n.age.0 <- meta.sub |> collapse::fsubset(age.min == 0) |> nrow()
n.age.1 <- meta.sub |> collapse::fsubset(age.min == 1) |> nrow()
n.age.2.plus <- meta.sub |> collapse::fsubset(age.min >= 2) |> nrow()
min.year <- min(res.lambda.fin$year)
max.year <- max(res.lambda.fin$year)
stock.ts.len <- res.lambda.fin |> collapse::fgroup_by(Stock.short) |> collapse::fsummarise(n.years = length(year))
mn.ts.len <- round(mean(stock.ts.len$n.years),digits=1)
mn.ts.len <- min(stock.ts.len$n.years)
mx.ts.len <- max(stock.ts.len$n.years)
num.orders <- length(unique(meta.sub$Order))
num.gads <- meta.sub |> collapse::fsubset(Order == "Gadiformes") |> nrow()
num.pleur <- meta.sub |> collapse::fsubset(Order == "Pleuronectiformes") |> nrow()
num.scorp <- meta.sub |> collapse::fsubset(Order == "Scorpaeniformes ") |> nrow()
num.clup <- meta.sub |> collapse::fsubset(Order == "Clupeiformes") |> nrow()
num.scomb <- meta.sub |> collapse::fsubset(Order == "Scombriformes") |> nrow()
num.perci <- meta.sub |> collapse::fsubset(Order == "Perciformes") |> nrow()

n.st.spec <- meta.sub |> collapse::fgroup_by(Species ) |> collapse::fsummarise(n.stocks = length(Stock)) 
n.st.spec <- n.st.spec[rev(order(n.st.spec$n.stocks)),]
num.cod <- meta.sub |> collapse::fsubset(Species == "morhua") |> nrow()
num.her <- meta.sub |> collapse::fsubset(Species == "harengus") |> nrow()
num.had <- meta.sub |> collapse::fsubset(Species == "aeglefinus") |> nrow()
num.sol <- meta.sub |> collapse::fsubset(Species == "solea") |> nrow()
num.pol <- meta.sub |> collapse::fsubset(Species == "virens") |> nrow()
num.plat <- meta.sub |> collapse::fsubset(Species == "platessa") |> nrow()

n.six <- meta.sub |> collapse::fsubset(Species %in% six.species) |> nrow()

med.lambda.F0 <- round(median(res.lambda.fin$lam.no.fish),digits=3)
med.r.F0 <- round(median(log(res.lambda.fin$lam.no.fish)),digits=3)*100
med.lambda.F <- round(median(res.lambda.fin$lam.fish),digits=3)
med.r.F <- round(median(log(res.lambda.fin$lam.fish)),digits=3)*100
mn.lambda.F <- round(mean(res.lambda.fin$lam.fish),digits=2)
mn.r.F <- round(mean(res.lambda.fin$lam.fish),digits=2)
# Mean - median...
res.lambda.fin$med.mn.diff.no.fish <- res.lambda.fin$lam.no.fish.med - res.lambda.fin$lam.no.fish.mn
res.lambda.fin$med.mn.diff.fish <- res.lambda.fin$lam.fish.med - res.lambda.fin$lam.fish.mn
# How many stocks had the mean higher than the median
n.mn.higher.med.no.fish <- length(unique(res.lambda.fin$Stock.short[res.lambda.fin$med.mn.diff.no.fish < 0]))
n.mn.higher.med.fish <- length(unique(res.lambda.fin$Stock.short[res.lambda.fin$med.mn.diff.fish < 0]))
low.mn.lambda.stocks <- unique(res.lambda.fin$Stock.short[res.lambda.fin$med.mn.diff.fish > 0])
n.nw.ice <- length(c(grep("NAFO",low.mn.lambda.stocks),grep("Faroe",low.mn.lambda.stocks)))
n.cod.low.mn <- length(grep("morhua",low.mn.lambda.stocks))
n.cod <- length(grep("morhua",Stocks))

# Density dependence related numbers
n.bl.40 <- length(which(!is.na(comp.lam.res$lam.fish.40)))
n.low.40.nf <- length(which(comp.lam.res$lam.nf.40 < 1))
n.low.40.fish <-length(which(comp.lam.res$lam.fish.40 < 1))
low.lam.40.nf <- -round(min(comp.lam.diff.res$lam.nf.40,na.rm=T),digits=3)
low.4.lam.40.nf <- -round(sort(comp.lam.diff.res$lam.nf.40)[4],digits=3)

# If right isn't this interesting!!! And look, 3 cod stocks from around here all say around 0.4 is the worst.  Most interesting...
n.low.lam.at.max.fish <- length(which(low.gam$low.pa.fish == 1))
n.low.lam.at.max.nf <- length(which(low.gam$low.pa.nf == 1))
low.lam.cods <- low.gam[low.gam$low.pa.nf < 1 & low.gam$Species == "G. morhua",]
n.cod.lam.not.at.max.nf <- nrow(low.lam.cods)
pa.low.cod.min <- round(min(low.lam.cods$low.pa.nf),digits=2)
pa.low.cod.max <- round(max(low.lam.cods$low.pa.nf),digits=2)

# more density dependence numbers
pa.lam.nf.at.one <- 100*round(no.stock.gam.nf.dat$prop.abund[min(which(no.stock.gam.nf.dat$fit <0))],digits=2)
pa.lam.fish.at.one <- 100*round(no.stock.gam.fish.dat$prop.abund[min(which(no.stock.gam.fish.dat$fit <0))],digits=2)
# What is the minimum abundance at which we get a 2 with the fishing lambdas.
pa.big.threshold <- 0.65
pa.low.threshold <- 0.8
lam.low.threshold <- 0.8
pa.lam.fish.low.bigs <- max(dd.dat$lam.fish[dd.dat$prop.abund >= pa.big.threshold])
pa.lam.nf.low.bigs <- round(rev(sort(dd.dat$lam.no.fish[dd.dat$prop.abund >= pa.big.threshold]))[2],digits =2)
# What proportion of the stocks had large declines overall, and then the proportion when the population was below 30%
dd.dat.high <- dd.dat |> collapse::fsubset(prop.abund >= 0.3)
prop.low.high.abund <- round(length(which(dd.dat.high$lam.fish < lam.low.threshold)) / nrow(dd.dat.high),digits=3)
prop.low.all <- round(length(which(dd.dat$lam.fish < lam.low.threshold)) / nrow(dd.dat),digits=3)
dd.dat.low <- dd.dat |> collapse::fsubset(prop.abund < 0.3)
prop.low.low.abund <- round(length(which(dd.dat.low$lam.fish < lam.low.threshold)) / nrow(dd.dat.low),digits=3)

# Numbers for the paper, lifetime reproductive growth
r0.mn <- round(mean(pop.dam$R0.opt,na.rm=T),digits=1)
r0.med <- round(median(pop.dam$R0.opt,na.rm=T),digits=1)
# And fishing...
r0.mn.fish <- round(mean(pop.dam.fish$R0.opt,na.rm=T),digits=1)
r0.med.fish <- round(median(pop.dam.fish$R0.opt,na.rm=T),digits=1)

# Where are the slow ones from
n.not.dub.20 <- nrow(slow.20)
n.alaska <- length(unique(dub.time.summary$Stock.short[dub.time.summary$Manage.short== "AFSC"]))
n.nwa <- length(unique(dub.time.summary$Stock.short[dub.time.summary$Manage.short %in% c("DFO-TRAC","NEFSC")]))
n.nea <- length(unique(dub.time.summary$Stock.short[dub.time.summary$Manage.short %in% c("ICES")]))
n.nafo.24 <- length(unique(dub.time.summary$Stock.short[dub.time.summary$Area.broad %in% "NAFO 2-4"]))
# Now the don't recover ones...
n.alaska.slow <- length(slow.20$Stock.short[slow.20$Manage.short== "AFSC"])
n.nwa.slow <- length(slow.20$Stock.short[slow.20$Manage.short %in% c("DFO-TRAC","NEFSC")])
n.nea.slow <- length(slow.20$Stock.short[slow.20$Manage.short %in% "ICES"])

# Need to grab all the per.diff's, which I seem to have dropped
for.diffs <- do.call('rbind',for.tune.summary)
med.per.diff <- round(median(for.diffs$per.diff,na.rm=T),digits=1)
# what is lambda difference, about 6%, but that means the original was 7% higher, so that means we are reducing lambda when optimizing
# If there was 'missing fishing' we are compensating for with this, this is exactly what we'd expect to happen right?
res.lambda.fin$lam.per.diff <- 100*((res.lambda.fin$lam.org.no.fish - res.lambda.fin$lam.no.fish)/ res.lambda.fin$lam.no.fish)
med.lam.per.diff <- round(median(res.lambda.fin$lam.per.diff,na.rm=T),digits=1)
# By stock, can see most are positive, which is in line with missing fishing mortality...
mod.by.stock.lam.pd <- res.lambda.fin |> collapse::fgroup_by(Stock.short) |> collapse::fsummarise(med = median(lam.per.diff,na.rm=T),
                                                                                                  mn = mean(lam.per.diff,na.rm=T))

```

```{r main-figs,include=F,echo=F,message=F,warning=F}

# New lambda plot with the fishing and no fishing beside each other
# This is likely Figure 1
mat <- matrix(c(2, 2, 2, 2, 2, 2, 2, 2,  2,  2,
                3, 3, 3, 3, 3, 3, 5, 5,  5,  5, 
                1, 1, 1, 1, 1, 1, 1, 1, NA, NA,
                4, 4, 4, 6, 6, 6, 6, NA, NA, NA), 
                4, 10, byrow = TRUE)
new.lam.plt.6 <- ggplot(res.lambda.fin.long.6) + geom_boxplot(aes(y=lambda,interaction(group,Stock.no.species),color=group)) +
                                           #geom_boxplot(aes(y=lam.no.fish,x=G.Species))  + 
                                           scale_x_discrete(NULL, guide = "axis_nested") +
                                           xlab("") + scale_y_log10("Lambda",breaks = log_breaks(n=8),limits=c(0.2,12)) + 
                                           geom_hline(yintercept = 1,linetype='dashed') +
                                           facet_manual(vars(G.Species),design=mat,scales='free_x')+
                                           theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none',
                                                 plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/Lambda_boxplot_6_species_",what.2.tune,".png"),new.lam.plt.6,base_width = 19,base_height = 15)

# The other stocks, perhaps we put this in a supplement...
mat <- matrix(c(1, 2, 2, 2, 2, 3, 4, 5,6,NA,NA,
                7, 7, rep(8,9),
                9, 9, rep(10,7),11,11),
                3, 11, byrow = TRUE)

new.lam.plt.others <- ggplot(res.lambda.fin.long.others) + geom_boxplot(aes(y=lambda,interaction(group,G.Species.short),color=group)) +
                                           #geom_boxplot(aes(y=lam.no.fish,x=G.Species))  + 
                                           scale_x_discrete(NULL, guide = "axis_nested") +
                                           xlab("") + scale_y_log10("Lambda",breaks = log_breaks(n=8),limits=c(0.2,12)) + 
                                           geom_hline(yintercept = 1,linetype='dashed') +
                                           facet_manual(vars(Area.long),design=mat,scales='free_x')+
                                           theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none',
                                                 plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/Lambda_boxplot_other_species_",what.2.tune,".png"),new.lam.plt.6,base_width = 19,base_height = 15)

# Standard deviation figure
colrs.sd <- rep(c(rep('blue',4),rep('orange',4),rep('grey',4),rep('black',4)),20)
pty.sd <- rep(rep(c(21,4,9,8),4),20)

lam.sd.plt <- ggplot(res.lambda.fin ) + 
                                       geom_point(aes(y = lam.no.fish.sd,x=G.Species,group=Area.long,color=Area.long,shape = Area.long,fill=Area.long),size=3) +
                                       scale_y_log10("Standard deviation (Lambda)",log_breaks(n=8)) +
                                       scale_color_manual(values=colrs.sd) + scale_shape_manual(values = pty.sd) + scale_fill_manual(values=colrs.sd)+
                                       theme(axis.text.x = element_text(angle =45,hjust=1),
                                                 plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/Lambda_sd_",what.2.tune,".png"),lam.sd.plt,base_width = 10,base_height = 10)

# Overall density dependence
# MAIN PAPER FIGURE#
dd.nf.overall.plt <-ggplot(no.stock.gam.nf.dat , aes(x=prop.abund,y=exp(fit))) + 
                                geom_point(data = dd.dat,aes(x=prop.abund,y=lam.no.fish),size=2,alpha=0.1) +  
                                geom_ribbon(aes(ymin=exp(LCI),ymax=exp(UCI)),alpha=0.5,linewidth=0,fill=u.colors[1])+
                                geom_line(color=u.colors[1],linewidth=1) + 
                                xlab("") + 
                                scale_y_log10(name = "Lambda(F=0)",breaks = log_breaks(n=10),limits=c(0.2,12)) +
                                geom_hline(yintercept = 1,color='firebrick2',linetype = 'dashed')+
                                theme(legend.text=element_text(size=8),
                                      legend.spacing.y = unit(-0.1, 'cm'),
                                      legend.title = element_blank(),
                                      axis.text.x = element_blank()) + 
                                 guides(color = guide_legend(ncol=1,byrow=T),
                                        shape = guide_legend(ncol=1,byrow=T))

# MAIN PAPER FIGURE#
dd.fish.overall.plt <- ggplot(no.stock.gam.fish.dat, aes(x=prop.abund,y=exp(fit))) + 
                                geom_point(data = dd.dat,aes(x=prop.abund,y=lam.fish),size=2,alpha=0.1) +  
                                geom_line(color=u.colors[2],linewidth=2) + 
                                geom_ribbon(aes(ymin=exp(LCI),ymax=exp(UCI)),alpha=0.5,fill=u.colors[2])+
                                xlab("") + 
                                scale_y_log10(name = "Lambda (F>0)",breaks = log_breaks(n=10),limits=c(0.2,12)) +
                                geom_hline(yintercept = 1,color='firebrick2',linetype = 'dashed',linewidth=1.5)+
                                theme(legend.text=element_text(size=8),
                                      legend.spacing.y = unit(-0.1, 'cm'),
                                      legend.title = element_blank(),
                                      axis.text.x = element_blank()) + 
                                 guides(color = guide_legend(ncol=1,byrow=T),
                                        shape = guide_legend(ncol=1,byrow=T))

# Main Paper Figure Density dependence with and without fishing, effectively a zoomed in version of the above.
colrs <- u.colors
t.smooth <- rbind(no.stock.gam.fish.dat,no.stock.gam.nf.dat)
dd.fish.nf.plt <- ggplot(t.smooth, aes(x=prop.abund,y=exp(fit),group=Scenario,color=Scenario,fill=Scenario)) + 
                                 geom_ribbon(aes(ymin=exp(LCI),ymax=exp(UCI),group=Scenario,fill=Scenario),alpha=0.50,linewidth=0)+
                                 geom_line(linewidth=1.5) +
                                 geom_hline(yintercept = 1,color='firebrick2',linetype = 'dashed')+
                                 scale_fill_manual(values = colrs) + scale_color_manual(values = colrs) +
                                 scale_x_continuous(name = "Proportion of maximum abundance",breaks = seq(0,1,by=0.1)) + 
                                 scale_y_log10(name = "Lambda",breaks = log_breaks(n=8)) +
                                 theme(legend.position = c(0.9,0.75))

dd.overall.combo <- plot_grid(dd.nf.overall.plt,dd.fish.overall.plt,dd.fish.nf.plt,nrow=3)

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/density_dependence_overall_",what.2.tune,".png"),dd.overall.combo,base_width = 10,base_height = 10)

new.lam.plt.others <- ggplot(res.lambda.fin.long.others) + geom_boxplot(aes(y=lambda,interaction(group,G.Species.short),color=group)) +
                                           #geom_boxplot(aes(y=lam.no.fish,x=G.Species))  + 
                                           scale_x_discrete(NULL, guide = "axis_nested") +
                                           xlab("") + scale_y_log10("Lambda",breaks = log_breaks(n=8),limits=c(0.2,12)) + 
                                           geom_hline(yintercept = 1,linetype='dashed') +
                                           facet_manual(vars(Area.long),design=mat,scales='free_x')

# Density dependence by stock 
cols.gam <- rep(c("black","blue","orange",'firebrick2','lightgreen'),6)
pty.gam <- c(rep(0,5),rep(1,5),rep(2,5),rep(3,5),rep(4,5),rep(5,5))
# Doing this manually because the ordered factor wasn't working... annoying...
mat <- matrix(c(2, 2,13,13, 5, 5, 9, 9,
                3, 3,12,12, 6, 6, 8, 8,
               NA,NA,10,10,11,11,NA,NA,
               NA,7, 7, 4, 4, 1, 1, NA),
                4, 8, byrow = TRUE)

dd.stocks.nf.plt <- ggplot(pred.dd.stock.gam.ss.nf,aes(x=prop.abund,y=lam.no.fish,group=Stock.short,color=G.Species,shape=G.Species)) + 
                                geom_line() + 
                                geom_point(data = thin.pred.nf,size=2) +  
                                xlab("Proportion of maximum abundance") + 
                                scale_y_log10(name = "Lambda (F=0)",breaks = log_breaks(n=10)) +
                                geom_ribbon(aes(ymin=lower,ymax=upper,color=G.Species),alpha=0.10,linewidth=0)+
                                facet_manual(vars(Area.long),design=mat,scales='free_x')+
                                #facet_wrap(~Area.long)  +
                                scale_shape_manual(values=pty.gam) +
                                scale_color_manual(values = cols.gam) + 
                                geom_hline(yintercept = 1,color='firebrick2',linetype = 'dashed')+
                                theme(legend.text=element_text(size=8),
                                      legend.spacing.y = unit(-0.1, 'cm'),
                                      legend.title = element_blank(),
                                      axis.text.x = element_text(size=7)) + 
                                 guides(color = guide_legend(ncol=1,byrow=T),
                                        shape = guide_legend(ncol=1,byrow=T))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/density_dependence_nf_stocks_",what.2.tune,".png"),dd.stocks.nf.plt,base_width = 10,base_height = 10)

mat <- matrix(c(2, 2,13,13, 5, 5, 9, 9,
                3, 3,12,12, 6, 6, 8, 8,
               NA,NA,10,10,11,11,NA,NA,
               NA,7, 7, 4, 4, 1, 1, NA),
                4, 8, byrow = TRUE)

# MAIN PAPER FIGURE#
dd.stocks.fish.plt <- ggplot(pred.dd.stock.gam.ss.fish,aes(x=prop.abund,y=lam.fish,group=Stock.short,color=G.Species,shape=G.Species)) + 
                                geom_line() + 
                                geom_point(data = thin.pred.fish,size=2) +  
                                geom_ribbon(aes(ymin=lower,ymax=upper,color=G.Species),alpha=0.10,linewidth=0)+
                                xlab("Proportion of maximum abundance") + 
                                scale_y_log10(name = "Lambda (F>0)",breaks = log_breaks(n=10)) +
                                facet_manual(vars(Area.long),design=mat,scales='free_x')+
                                #facet_wrap(~Area.long)  +
                                scale_shape_manual(values=pty.gam) +
                                scale_color_manual(values = cols.gam) + 
                                geom_hline(yintercept = 1,color='firebrick2',linetype = 'dashed')+
                                theme(legend.text=element_text(size=8),
                                      legend.spacing.y = unit(-0.1, 'cm'),
                                      legend.title = element_blank(),
                                      axis.text.x = element_text(size=7)) + 
                                 guides(color = guide_legend(ncol=1,byrow=T),
                                        shape = guide_legend(ncol=1,byrow=T))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/density_dependence_fish_stocks_",what.2.tune,".png"),dd.stocks.fish.plt,base_width = 10,base_height = 10)

# Lifetime reproductive success 
mat <- matrix(c(2, 2, 2, 2, 2, 2, 2, 2,  2,  2,
                3, 3, 3, 3, 3, 3, 5, 5,  5,  5, 
                1, 1, 1, 1, 1, 1, 1, 1, NA, NA,
                4, 4, 4, 6, 6, 6, 6, NA, NA, NA), 
                4, 10, byrow = TRUE)
r0.boxplt.6 <- ggplot(pop.dam.all.6) + geom_boxplot(aes(y=R0.opt,interaction(group,Stock.no.species),color=group))  + 
                                       scale_x_discrete(NULL, guide = "axis_nested") + scale_y_log10(name = "Lifetime Reproductive Output")+
                                       geom_hline(yintercept = 1,linetype='dashed',color='blue')+
                                       facet_manual(vars(G.Species),design=mat,scales='free_x')+
                                           theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none',
                                                 plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/R0_boxplot_6_species_",what.2.tune,".png"),r0.boxplt.6,base_width = 19,base_height = 15)

mat <- matrix(c(1, 2, 2, 2, 2, 3, 4, 5,6,NA,NA,
                7, 7, rep(8,9),
                9, 9, rep(10,7),11,11),
                3, 11, byrow = TRUE)

r0.boxplot.others <- ggplot(pop.dam.all.others) +   geom_boxplot(aes(y=R0.opt,interaction(group,G.Species.short),color=group))  + 
                                                    scale_x_discrete(NULL, guide = "axis_nested") + 
                                                    scale_y_log10(name = "Lifetime Reproductive Output")+
                                                    geom_hline(yintercept = 1,linetype='dashed',color='blue')+
                                                    facet_manual(vars(Area.long),design=mat,scales='free')+
                                                    theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none',
                                                          plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/R0_boxplot_other_species_",what.2.tune,".png"),r0.boxplot.others,base_width = 15,base_height = 10)

r0.heatmap.ne <- ggplot(pop.dam |> collapse::fsubset(LME == "NE Atlantic"), aes(x=cohort, y=Stock.short, fill= R0.opt)) + 
                                    xlab("") + ylab("")+
                                    facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + 
                                    geom_tile() + 
                                     scale_fill_viridis(name = "Lifetime reproductive success", breaks = c(0.4,1,2,4,10,20,40,100,200,400),
                                                        discrete=FALSE,trans='log') + 
                                    theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.9, "cm"))

r0.heatmap.nw <- ggplot(pop.dam |> collapse::fsubset(LME == "NW Atlantic"), aes(x=cohort, y=Stock.short, fill= R0.opt)) + 
                                    xlab("") + ylab("")+
                                    facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                                    geom_tile() + 
                                     scale_fill_viridis(name = "Lifetime reproductive success", breaks = c(0.4,1,2,4,10,20,40,100,200,400),
                                                        discrete=FALSE,trans='log') + 
                                    theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,1.22, "cm"))

r0.heatmap.pac <- ggplot(pop.dam |> collapse::fsubset(LME == "Pacific (Alaska)"), aes(x=cohort, y=Stock.short, fill= R0.opt)) + 
                                     xlab("") + ylab("")+
                                     facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                                     geom_tile() + 
                                     scale_fill_viridis(name = "Lifetime reproductive success", breaks = c(0.4,1,2,4,10,20,40,100,200,400),
                                                        discrete=FALSE,trans='log') + 
                                     theme(legend.position = 'top',text = element_text(size=12),plot.margin = margin(0,0.5,0,0, "cm"),
                                           legend.key.width = unit(3,'cm'))

r0.heatmap <- cowplot::plot_grid(r0.heatmap.pac,r0.heatmap.nw,r0.heatmap.ne,nrow = 3,rel_heights = c(3,4,10))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/r0_heatmap_nf_",what.2.tune,".png"),r0.heatmap,base_width = 20,base_height = 17)

r0.heatmap.ne.fish <- ggplot(pop.dam.fish |> collapse::fsubset(LME == "NE Atlantic"), aes(x=cohort, y=Stock.short, fill= R0.opt)) + 
                                    xlab("") + ylab("")+
                                    facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + 
                                    geom_tile() + 
                                     scale_fill_viridis(name = "Lifetime reproductive success", breaks = c(0.4,1,2,4,10,20,40,100,200,400),
                                                        discrete=FALSE,trans='log') + 
                                    theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.9, "cm"))

r0.heatmap.nw.fish <- ggplot(pop.dam.fish |> collapse::fsubset(LME == "NW Atlantic"), aes(x=cohort, y=Stock.short, fill= R0.opt)) + 
                                    xlab("") + ylab("")+
                                    facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                                    geom_tile() + 
                                     scale_fill_viridis(name = "Lifetime reproductive success", breaks = c(0.4,1,2,4,10,20,40,100,200,400),
                                                        discrete=FALSE,trans='log') + 
                                    theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,1.22, "cm"))

r0.heatmap.pac.fish <- ggplot(pop.dam.fish |> collapse::fsubset(LME == "Pacific (Alaska)"), aes(x=cohort, y=Stock.short, fill= R0.opt)) + 
                                     xlab("") + ylab("")+
                                     facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                                     geom_tile() + 
                                     scale_fill_viridis(name = "Lifetime reproductive success", breaks = c(0.4,1,2,4,10,20,40,100,200,400),
                                                        discrete=FALSE,trans='log') + 
                                     theme(legend.position = 'top',text = element_text(size=12),plot.margin = margin(0,0.5,0,0, "cm"),
                                           legend.key.width = unit(3,'cm'))

r0.heatmap.fish <- cowplot::plot_grid(r0.heatmap.pac.fish,r0.heatmap.nw.fish,r0.heatmap.ne.fish,nrow = 3,rel_heights = c(3,4,10))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/r0_heatmap_fish_",what.2.tune,".png"),r0.heatmap.fish,base_width = 20,base_height = 17)

# Generation length
mat <- matrix(c(2, 2, 2, 2, 2, 2, 2, 2,  2,  2,
                3, 3, 3, 3, 3, 3, 5, 5,  5,  5, 
                1, 1, 1, 1, 1, 1, 1, 1, NA, NA,
                4, 4, 4, 6, 6, 6, 6, NA, NA, NA), 
                4, 10, byrow = TRUE)
gen.boxplt.6 <- ggplot(pop.dam.all.6) +
                            geom_boxplot(aes(y=gen.len.opt,interaction(group,Stock.no.species),color=group))  + 
                            scale_x_discrete(NULL, guide = "axis_nested") +
                            scale_y_continuous("Generation Length (years)",limits=c(0,13),breaks= seq(0,20,by=2)) +
                             facet_manual(vars(G.Species),design=mat,scales='free_x')+
                                           theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none',
                                                 plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/GL_boxplot_6_species_",what.2.tune,".png"),gen.boxplt.6,base_width = 19,base_height = 15)

# Then the generation length for others.
mat <- matrix(c(1, 2, 2, 2, 2, 3, 4, 5,6,NA,NA,
                7, 7, rep(8,9),
                9, 9, rep(10,7),11,11),
                3, 11, byrow = TRUE)

gen.boxplt.others <- ggplot(pop.dam.all.others) +
                            geom_boxplot(aes(y=gen.len.opt,interaction(group,G.Species.short),color=group))  + 
                            scale_x_discrete(NULL, guide = "axis_nested") +
                            scale_y_continuous("Generation Length (years)",limits=c(0,13),breaks= seq(0,20,by=2)) +
                            facet_manual(vars(Area.long),design=mat,scales='free')+
                                           theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none',
                                                 plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/GL_boxplot_other_species_",what.2.tune,".png"),gen.boxplt.others,base_width = 15,base_height = 10)

gen.heatmap.ne <- ggplot(pop.dam |> collapse::fsubset(LME == "NE Atlantic"), aes(x=cohort, y=Stock.short, fill= gen.len.opt)) + 
                                              xlab("") + ylab("")+  
                                              facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + 
                                              geom_tile() + scale_fill_viridis(name = "Generation length (years)",discrete=FALSE) + 
                                              theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.9, "cm"))

gen.heatmap.nw <- ggplot(pop.dam |> collapse::fsubset(LME == "NW Atlantic"), aes(x=cohort, y=Stock.short, fill= gen.len.opt)) + xlab("") + ylab("")+                                    facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                            geom_tile() + scale_fill_viridis(name = "Generation length (years)",discrete=FALSE) + 
                            theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,1.22, "cm"))

gen.heatmap.pac <- ggplot(pop.dam |> collapse::fsubset(LME == "Pacific (Alaska)"), aes(x=cohort, y=Stock.short, fill= gen.len.opt)) + xlab("") + ylab("")+
                            facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                            geom_tile() + scale_fill_viridis(name = "Generation length (years)",discrete=FALSE) + 
                            theme(legend.position = 'top',text = element_text(size=12),plot.margin = margin(0,0.5,0,0, "cm"),
                                  legend.key.width = unit(3,'cm'))

gen.heatmap <- cowplot::plot_grid(gen.heatmap.pac,gen.heatmap.nw,gen.heatmap.ne,nrow = 3,rel_heights = c(3,4,10))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/GL_heatmap_nf_",what.2.tune,".png"),gen.heatmap,base_width = 15,base_height = 10)

gen.heatmap.ne.fish <- ggplot(pop.dam.fish |> collapse::fsubset(LME == "NE Atlantic"), aes(x=cohort, y=Stock.short, fill= gen.len.opt)) + 
                                              xlab("") + ylab("")+  
                                              facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + 
                                              geom_tile() + scale_fill_viridis(name = "Generation length (years)",discrete=FALSE) + 
                                              theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.9, "cm"))

gen.heatmap.nw.fish <- ggplot(pop.dam.fish |> collapse::fsubset(LME == "NW Atlantic"), aes(x=cohort, y=Stock.short, fill= gen.len.opt)) + xlab("") + ylab("")+                                    facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                            geom_tile() + scale_fill_viridis(name = "Generation length (years)",discrete=FALSE) + 
                            theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,1.22, "cm"))

gen.heatmap.pac.fish <- ggplot(pop.dam.fish |> collapse::fsubset(LME == "Pacific (Alaska)"), aes(x=cohort, y=Stock.short, fill= gen.len.opt)) + xlab("") + ylab("")+
                            facet_wrap(~LME,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                            geom_tile() + scale_fill_viridis(name = "Generation length (years)",discrete=FALSE) + 
                            theme(legend.position = 'top',text = element_text(size=12),plot.margin = margin(0,0.5,0,0, "cm"),
                                  legend.key.width = unit(3,'cm'))

gen.heatmap.fish <- cowplot::plot_grid(gen.heatmap.pac.fish,gen.heatmap.nw.fish,gen.heatmap.ne.fish,nrow = 3,rel_heights = c(3,4,10))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/GL_heatmap_fish_",what.2.tune,".png"),gen.heatmap.fish,base_width = 15,base_height = 10)

dub.time.summary <- dub.time.summary |> collapse::fsubset(year %in% 1:50) |>
                            collapse::roworder(G.Species,LME,Area.broad) |>
                            collapse::fmutate(Stock.short = forcats::fct_inorder(factor(Stock.short, ordered=TRUE)))

dub.heatmap.gad <- ggplot(dub.time.summary |> collapse::fsubset(Order == "Gadiformes"), aes(x=year, y=Stock.short, fill= prop)) + 
                            xlab("") + ylab("")+                                    
                            facet_wrap(~Order.short,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                            geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
                            theme(legend.position = 'top',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.45, "cm"),
                                  legend.key.width = unit(3,'cm'))

dub.heatmap.pl <- ggplot(dub.time.summary |> collapse::fsubset(Order == "Pleuronectiformes"), aes(x=year, y=Stock.short, fill= prop)) + 
                            xlab("") + ylab("")+                                    
                            facet_wrap(~Order.short,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                            geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
                            theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0, "cm"))

dub.heatmap.cl <- ggplot(dub.time.summary |> collapse::fsubset(Order == "Clupeiformes"), aes(x=year, y=Stock.short, fill= prop)) + 
                            xlab("") + ylab("")+
                            facet_wrap(~Order.short,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                            geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
                            theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,1.35, "cm"),
                                  legend.key.width = unit(3,'cm'))

dub.heatmap.per <- ggplot(dub.time.summary |> collapse::fsubset(Order == "Perciformes"), aes(x=year, y=Stock.short, fill= prop)) + 
                            xlab("") + ylab("")+
                            facet_wrap(~Order.short,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                            geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
                            theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,2, "cm"),
                                  legend.key.width = unit(3,'cm'))

dub.heatmap.scom <- ggplot(dub.time.summary |> collapse::fsubset(Order == "Scombriformes"), aes(x=year, y=Stock.short, fill= prop)) + 
                            xlab("") + ylab("")+
                            facet_wrap(~Order.short,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
                            geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
                            theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,2.1, "cm"),
                                  legend.key.width = unit(3,'cm'))

dub.heatmap.scorp <- ggplot(dub.time.summary |> collapse::fsubset(Order %in% "Scorpaeniformes "), aes(x=year, y=Stock.short, fill= prop)) + 
                            xlab("") + ylab("")+
                            facet_wrap(~Order.short,nrow=3,scales='free_y',strip.position = "right") + 
                            geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
                            theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.5, "cm"),
                                  legend.key.width = unit(3,'cm'))

dub.heatmap <- cowplot::plot_grid(dub.heatmap.gad,dub.heatmap.pl,dub.heatmap.cl,dub.heatmap.per,dub.heatmap.scom,dub.heatmap.scorp,
                                  nrow = 6,rel_heights = c(39,20,11,4,3,3))

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/Doubling_time_heatmap_by_order_",what.2.tune,".png"),dub.heatmap,base_width = 20,base_height = 20)

# Doubling Time Figure 1
dub.ts.plt <- ggplot(n.dub,aes(x=years,y=count)) + 
                    geom_line() + 
                    geom_hline(yintercept = length(Stocks),linetype="dashed") + 
                    xlab("") + ylab("Number of Stocks")

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/Doubling_time_time_series_",what.2.tune,".png"),dub.ts.plt,base_width = 10,base_height = 10)

# Doubling time and variability

# Is there a relationship between median and sd?
colrs.cat <- c("blue","black","orange")
pts.cat <- 21:23
sd.rec.plt <- ggplot(lam.sd.p.rec, aes(y=lam.no.fish.med,x=lam.no.fish.sd,color=prop.cat,group=prop.cat)) + 
                       geom_point(size =3,aes(shape=prop.cat,fill=prop.cat)) + 
                       geom_hline(yintercept=1,linetype="dashed") + 
                       scale_y_continuous(name="median Lambda") + scale_x_continuous(name="standard deviation Lambda") +
                       scale_color_manual(values=colrs.cat) + scale_fill_manual(values=colrs.cat) +  
                       scale_shape_manual(values=pts.cat) + theme(legend.title = element_blank())

save_plot(paste0(loc,"/Figures/tuned/No_age_correction/Doubling_time_sd_effects_",what.2.tune,".png"),sd.rec.plt,base_width = 10,base_height = 10)

```

```{r the-others,include=F,echo=F,message=F,warning=F}
#             
# 
# dub.heatmap.cod <- ggplot(dub.time.6 |> collapse::fsubset(G.Species == "G. morhua"), aes(x=year, y=Stock.no.species, fill= prop)) + xlab("") + ylab("")+                                     facet_wrap(~G.Species,nrow=3,scales='free_y',strip.position = "right") + 
#                             geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + scale_x_continuous(labels=NULL) +
#                             theme(legend.position = 'top',text = element_text(size=12),plot.margin = margin(0,0.5,0,0, "cm"),
#                                   legend.key.width = unit(3,'cm'))
# 
# dub.heatmap.herring<- ggplot(dub.time.6 |> collapse::fsubset(G.Species == "C. harengus"), aes(x=year, y=Stock.no.species, fill= prop)) + 
#                               xlab("") + ylab("")+                                    
#                               facet_wrap(~G.Species,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
#                               geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
#                               theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.4, "cm"))
# 
# dub.heatmap.haddock<- ggplot(dub.time.6 |> collapse::fsubset(G.Species == "M. aeglefinus"), aes(x=year, y=Stock.no.species, fill= prop)) + 
#                               xlab("") + ylab("")+                                    
#                               facet_wrap(~G.Species,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
#                               geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
#                               theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.125, "cm"))
# 
# dub.heatmap.virens<- ggplot(dub.time.6 |> collapse::fsubset(G.Species == "P. virens"), aes(x=year, y=Stock.no.species, fill= prop)) + 
#                               xlab("") + ylab("")+                                    
#                               facet_wrap(~G.Species,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
#                               geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
#                               theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,1.4, "cm"))
# 
# dub.heatmap.solea<- ggplot(dub.time.6 |> collapse::fsubset(G.Species == "S. solea"), aes(x=year, y=Stock.no.species, fill= prop)) + 
#                               xlab("") + ylab("")+                                    
#                               facet_wrap(~G.Species,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous(labels=NULL) +
#                               geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
#                               theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.4, "cm"))
# 
# dub.heatmap.platessa<- ggplot(dub.time.6 |> collapse::fsubset(G.Species == "P. platessa"), aes(x=year, y=Stock.no.species, fill= prop)) + 
#                               xlab("") + ylab("")+                                    
#                               facet_wrap(~G.Species,nrow=3,scales='free_y',strip.position = "right") + scale_x_continuous() +
#                               geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
#                               theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.68, "cm"))
# 
# 
# 
# dub.heatmap.sp6 <- cowplot::plot_grid(dub.heatmap.cod,dub.heatmap.herring,dub.heatmap.haddock,
#                                  dub.heatmap.virens, dub.heatmap.solea,dub.heatmap.platessa,nrow = 6,rel_heights = c(14,9,8,5,5,4))
# dub.heatmap.sp6
# 
# save_plot(paste0(loc,"/Figures/tuned/No_age_correction/Doubling_time_heatmap_by_sp6_",what.2.tune,".png"),dub.heatmap.sp6,base_width = 20,base_height = 17)
# 
# 
# # What about a plot of the sd of lamda by stock
# #colrs.sd <- rep(rep(c("black","blue","orange",'grey'),3),20)
# #pty.sd <- rep(c(rep(21,4),rep(4,4),rep(9,4)),20)
# 
# 
# # Same plot for the not 6 species
# 
# # lam.sd.other.plt <- ggplot(res.lambda.fin |> collapse::fsubset(!Species %in% six.species)) + 
# #                                        geom_point(aes(y = lam.no.fish.sd,x=G.Species,group=Area.long,color=Area.long,shape = Area.long,fill=Area.long),size=3) +
# #                                        scale_y_log10("sd(lambda)",limits=c(0.05,1.5)) +
# #                                        scale_color_manual(values=colrs.sd) + scale_shape_manual(values = pty.sd) + scale_fill_manual(values=colrs.sd)+
# #                                        theme(axis.text.x = element_text(angle =45,hjust=1),
# #                                                  plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))
# # lam.sd.other.plt
# 
# 
# 
# 
# # Doubling Times
# 
# 
# # Generation length
# 
# # Main text figure!!
# new.lam.plt.6 <- ggplot(res.lambda.fin.long.6) + geom_boxplot(aes(y=lambda,interaction(group,Area.long),color=group)) +
#                                            #geom_boxplot(aes(y=lam.no.fish,x=G.Species))  + 
#                                            scale_x_discrete(NULL, guide = "axis_nested") +
#                                            xlab("") + scale_y_log10("lambda",breaks = c(0.1,0.4,0.6,0.8,1,1.25,1.75,3,5,10),limits=c(0.4,10)) + 
#                                            geom_hline(yintercept = 1,linetype='dashed') +
#                                            facet_wrap(~G.Species,scales='free')+
#                                            theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none',
#                                                  plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))
# 
# # The other stocks, plotting these by area
# # Main Text FIGURE!
# # More data
# 
# 
# ################### Supplemental Figures################### Supplemental Figures################### Supplemental Figures################### Supplemental 
# 
# 
# 
# # Time series of lambdas....
# col.try <- rep(rep(c("black","blue","orange",'firebrick2','lightgreen'),3),20)
# #col.try <- rep(c(rep("black",3),rep("blue",3),rep('orange',3),rep('firebrick2',3),rep('lightgreen',3)),20)
# pty <- rep(c(rep(21,5),rep(22,5),rep(23,5)),20)
# #pty <- rep(rep(c(21:23),5),20)
# 
# lambda.ts.plt <- ggplot(res.lambda.final.6,aes(y=lam.fish,x=year,group=Stock.short,color=Stock.short,shape = Stock.short,fill = Stock.short)) +
#                             geom_point(size=4) +
#                             geom_smooth(method='gam',formula = y ~ s(x, bs = "cs", fx = TRUE, k = 5),se=F) + 
#                             #scale_y_log10(name = "lambda") +
#                             facet_wrap(~G.Species,scales='free_y') +
#                             scale_color_manual(name = "Stock",values = col.try) + scale_shape_manual(name = "Stock",values = pty) +
#                             scale_fill_manual(name = "Stock",values = col.try)+
#                             guides(colour = guide_legend(ncol = 1))+
#                             theme(axis.text.x = element_text(angle =45,hjust=1),legend.key.width = unit(5,'cm'))
# #lambda.ts.plt
# save_plot(paste0(loc,"/Figures/tuned/No_age_correction/lambda_timeseries_6_",what.2.tune,".png"),lambda.ts.plt,base_width = 12,base_height = 10)
# 
# 
# lambda.dec.plt <- ggplot(res.lambda.final.6,aes(y=med.lam,x=as.factor(decade))) +
#                             geom_boxplot() + scale_y_log10()+
#                             xlab("Decade") + ylab("Lambda") + 
#                             geom_point(aes(x=as.factor(decade),y=lm.spec.mn),color='blue',shape=8,fill='blue')+
#                             #scale_x_continuous(breaks=seq(1940,2020,by=10))+
#                             geom_hline(yintercept = 1,linetype='dashed') +
#                             facet_wrap(~G.Species,scales='free_y') 
# save_plot(paste0(loc,"/Figures/tuned/No_age_correction/lambda_by_decade_6_",what.2.tune,".png"),lambda.dec.plt,base_width = 12,base_height = 12)
# 
#

# ggplot(dd.dat) + geom_point(aes(x=prop.abund,y=lam.no.fish-1)) + xlab("Proportion of Maximum Abundance") + ylab("r (no fishing)")
# 
# # Density dependence by Species
# ggplot(dd.dat,aes(x=prop.abund,y=lam.no.fish,color=Species,group=Species)) + geom_point() + 
#                                                    geom_smooth(method="gam") + 
#                                                    geom_hline(yintercept = 1,linewidth=2,linetype='dashed',color='blue') +
#                                                    scale_x_continuous(name = "Proportion of Maximum Abundance", breaks = seq(0,1,by=0.1))+
#                                                    scale_y_log10(name = "lambda (no fishing)", breaks = c(0.5,0.75,1,1.5,2.5,3.5,5,7.5,10,12.5)) #+
#                                                    #scale_color_manual(values = colrs)
# # Density dependence by stock
# ggplot(dd.dat,aes(x=prop.abund,y=lam.no.fish,color=Stock.short,group=Stock.short)) + geom_point() + 
#                                                    geom_smooth(method="lm",se = F) + 
#                                                    geom_hline(yintercept = 1,linewidth=2,linetype='dashed',color='blue') +
#                                                    scale_x_continuous(name = "Proportion of Maximum Abundance", breaks = seq(0,1,by=0.1))+
#                                                    scale_y_log10(name = "lambda (no fishing)", limits=c(0.4,2),
#                                                                  breaks = seq(0.4,2,by=0.1)) +
#                                                    theme(legend.position = 'none')
# 
# 
# 
# 
# ggplot(dd.dat.no.bigs,aes(x=prop.abund,y=lam.no.fish)) + geom_point() + 
#                                                     geom_smooth(method="gam") + 
#                                                    geom_hline(yintercept = 1,linewidth=2,linetype='dashed',color='blue') +
#                                                    scale_x_continuous(name = "Proportion of Maximum Abundance", breaks = seq(0,1,by=0.1))+
#                                                    scale_y_continuous(name = "lambda (no fishing)")
# 
# ggplot(dd.dat.no.bigs,aes(x=prop.abund,y=lam.no.fish,group=Stock.short,color=Stock.short)) + geom_point() + 
#                                                     geom_smooth(method="lm",se=F) + 
#                                                    geom_hline(yintercept = 1,linewidth=2,linetype='dashed',color='blue') +
#                                                    scale_x_continuous(name = "Proportion of Maximum Abundance", breaks = seq(0,2,by=0.1))+
#                                                    scale_y_log10(name = "lambda (no fishing)")
# 
# 
# # how about our big 6
# colrs <- c("blue","black","firebrick2","orange",'grey50','grey90')
# ggplot(dd.dat.six,aes(x=prop.abund,y=lam.no.fish,color=Species,group=Species)) + geom_point() + 
#                                                     geom_smooth(method="gam") + 
#                                                    geom_hline(yintercept = 1,linewidth=2,linetype='dashed',color='blue') +
#                                                    scale_x_continuous(name = "Proportion of Maximum Abundance", breaks = seq(0,1,by=0.1))+
#                                                    scale_y_log10(name = "lambda (no fishing)", breaks = c(0.5,0.75,1,1.5,2.5,3.5,5,7.5,10,12.5)) +
#                                                    scale_color_manual(values = colrs)
# 
# 
# ggplot(dd.dat.six.no.bigs,aes(x=prop.abund,y=lam.no.fish-1,color=Species,group=Species)) + geom_point() + 
#                                                     geom_smooth(method="gam") + 
#                                                    geom_hline(yintercept = 0,linewidth=2,linetype='dashed',color='blue') +
#                                                    scale_x_continuous(name = "Proportion of Maximum Abundance", breaks = seq(0,1,by=0.1))+
#                                                    scale_y_continuous(name = "r (no fishing)", breaks = seq(-1,1,by=0.2)) +
#                                                    scale_color_manual(values = colrs)

# Other things I did just randomly scattered here

# 
# #hmm <- merge(tst,res.lambda.final.6,by="Stock")
# plot(res.lambda.fin$lam.fish.med ~ res.lambda.fin$lam.fish.mn)
# # So this is lambda with fishing included
# lambda.plt <- ggplot(res.lambda.final.6) + geom_boxplot(aes(y=lam.no.fish,x=Stock.no.species))  + 
#                                            geom_point(aes(y=lam.no.fish.mn,x=Stock.no.species),color='blue',shape=8,fill='blue') +
#                                            xlab("") + scale_y_log10("lambda",breaks = c(0.1,0.4,0.6,0.8,1,1.25,1.75,3,5,10),limits=c(0.4,10)) + 
#                                            geom_hline(yintercept = 1,linetype='dashed') +
#                                            facet_wrap(~G.Species,scales='free')+
#                                            theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none',
#                                                  plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))
# 
# save_plot(paste0(loc,"/Figures/tuned/No_age_correction/lambda_no_fishing_6_",what.2.tune,".png"),lambda.plt,base_width = 12,base_height = 10)
# 
# # With fishing
# lambda.plt.fish <- ggplot(res.lambda.final.6) + geom_boxplot(aes(y=lam.fish,x=Stock.no.species))  + 
#                                            geom_point(aes(y=lam.fish.mn,x=Stock.no.species),color='blue',shape=8,fill='blue') +
#                                            xlab("") + scale_y_log10("lambda",breaks = c(0.4,0.6,0.8,1,1.25,1.75,3,5,10),limits=c(0.4,10)) + 
#                                            geom_hline(yintercept = 1,linetype='dashed') +
#                                            facet_wrap(~G.Species,scales='free')+
#                                            theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none',
#                                                  plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))
# 
# 
# save_plot(paste0(loc,"/Figures/tuned/No_age_correction/lambda_with_fishing_6_",what.2.tune,".png"),lambda.plt.fish,base_width = 12,base_height = 10)
# 
# 
# 
# new.lam.plt <- ggplot(res.lambda.fin.long) + geom_boxplot(aes(y=lambda,interaction(group,Order),color=group)) +
#                                            #geom_boxplot(aes(y=lam.no.fish,x=G.Species))  + 
#                                            scale_x_discrete(NULL, guide = "axis_nested") +
#                                            #xlab("") + scale_y_log10("lambda",breaks = c(0.1,0.4,0.6,0.8,1,1.25,1.75,3,5,10),limits=c(0.4,10)) + 
#                                            #geom_hline(yintercept = 1,linetype='dashed') +
#                                            facet_wrap(~Area.long,scales='free')+
#                                            theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none',
#                                                  plot.margin = margin(0,0.5,0,0.5, "cm"),axis.title.y = element_text(margin = margin(r = 30)))
# 
# 
# # The final option I have is to do what I did before and use the bin props
# We can then do a much more ANOVA kinda analysis, drat this doesn't converge..., estimate the 
#mod.dd.bins.stock <- glmer(lam.no.fish ~ prop.bin + (1+prop.bin|Stock),data=dd.dat,family=gaussian(link='log'))
# so go to lmer and log transform... and it is a singular fit
#mod.dd.bins.stock <- lmer(lam.no.fish ~ prop.bin + (1+prop.bin|Stock),data=dd.dat)
# So is an intercept model of any real use?
# mod.dd.bins.stock <- lmer(log(lam.no.fish) ~ prop.bin + (1+prop.bin|Stock.short),data=dd.dat)
# # A non-random effect model does work with the interaction... I'd much rather lmer....
# #mod.dd.bins.stock <- lm(lam.no.fish ~ prop.bin*Stock,data=dd.dat)
# summary(mod.dd.bins.stock)
# #Residuals, nah, still suck, just some crazy outliers...
# plot(mod.dd.bins.stock)
# # Normality = terrible...
# qqnorm(residuals(mod.dd.bins.stock))
# qqline(residuals(mod.dd.bins.stock))
# # So we can predict this ANOVA thing...
# pred.dd.stock.bins <- expand.grid(prop.bin = unique(dd.dat$prop.bin),
#                                          #Species=unique(dd.dat$Species),
#                                          Stock.short=unique(dd.dat$Stock.short))
# tmp <- predict(mod.dd.bins.stock, pred.dd.stock.bins, type = 'response', se.fit = TRUE) 
# pred.dd.stock.bins$lam.no.fish   <- exp(tmp$fit)
# pred.dd.stock.bins$lower <- exp(tmp$fit - 1.96 * tmp$se.fit)
# pred.dd.stock.bins$upper <- exp(tmp$fit + 1.96 * tmp$se.fit)
# # Now to do fun things I need the meta data on the pred dataframe
# pred.dd.stock.bins <- left_join(pred.dd.stock.bins,meta.sub,by="Stock.short")
# # Now a problem here is that this includes estimates for bins in which we don't have data.
# # So we do want to get rid of those
# real.bins <- unique(paste(dd.dat$Stock.short,"_",dd.dat$prop.bin))
# # All the bins in the predict object with same names same name...
# pred.dd.stock.bins$bin.combos <- paste(pred.dd.stock.bins$Stock.short,"_",pred.dd.stock.bins$prop.bin)
# # Now toss all the predicted bins without data.
# pred.dd.stock.bins <- pred.dd.stock.bins |> collapse::fsubset(bin.combos %in% real.bins)
# 
# # By region
# cols.1 <- rep(c("black","blue","orange",'firebrick2','lightgreen'),6)
# #col.try <- rep(c(rep("black",3),rep("blue",3),rep('orange',3),rep('firebrick2',3),rep('lightgreen',3)),20)
# pty.1 <- c(rep(0,5),rep(1,5),rep(2,5),rep(3,5),rep(4,5),rep(5,5))
# ggplot(pred.dd.stock.bins,aes(x=prop.bin,y=lam.no.fish,group=Stock.short,color=G.Species)) + 
#                                 geom_point(aes(shape=G.Species),size=2) + geom_line()  + xlab("Proportion of maximum abundance") + 
#                                 scale_y_log10(name = "Lambda",breaks = log_breaks(n=8)) +
#                                 facet_wrap(~Area.long)  +
#                                 scale_shape_manual(values=pty.1) +
#                                 scale_color_manual(values = cols.1) + 
#                                 geom_hline(yintercept = 1,color='grey',linetype = 'dashed')+
#                                 theme(legend.text=element_text(size=8),
#                                       legend.spacing.y = unit(-0.1, 'cm'),
#                                       legend.title = element_blank(),
#                                       axis.text.x = element_text(size=7)) + 
#                                  guides(color = guide_legend(ncol=1,byrow=T),
#                                         shape = guide_legend(ncol=1,byrow=T))
# 
# # By Species
# ggplot(pred.dd.stock.bins,aes(x=prop.bin,y=lam.no.fish,group=Stock.short,color=Area.long)) + 
#                                 geom_point(aes(shape=Area.long)) + geom_line()  + xlab("Proportion of maximum abundance") + 
#                                 scale_y_log10(name = "Lambda",breaks = log_breaks(n=8)) +
#                                 facet_wrap(~Order)  +
#                                 scale_shape_manual(values=pty.1) +
#                                 scale_color_manual(values = cols.1) + 
#                                 geom_hline(yintercept = 1,color='grey',linetype = 'dashed')+
#                                 theme(legend.text=element_text(size=8),
#                                       legend.spacing.y = unit(-0.1, 'cm'),
#                                       legend.title = element_blank(),
#                                       axis.text.x = element_text(size=7)) + 
#                                  guides(color = guide_legend(ncol=1,byrow=T),
#                                         shape = guide_legend(ncol=1,byrow=T))
# 
# # No fish vs fish lambda, note that DD is weaker when we account for fishing, that seems a fun result to talk about!
# 
# mod.dd.overall.gam.nf <- lm(log(lam.no.fish) ~ (prop.abund),data=dd.dat)
# mod.dd.overall.gam.fish <- lm(log(lam.fish) ~ (prop.abund),data=dd.dat)
# 
# 
# pred.dd.overall.bins.nf <- data.frame(prop.abund = seq(0,1,length.out=500),group = "F = 0")
# pred.dd.overall.bins.fish <- data.frame(prop.abund = seq(0,1,length.out=500),group = "F > 0")
# tmp <- predict(mod.dd.overall.gam.nf, pred.dd.overall.bins, type = 'response', se.fit = TRUE) 
# tmp.fish <- predict(mod.dd.overall.gam.fish, pred.dd.overall.bins, type = 'response', se.fit = TRUE) 
# pred.dd.overall.bins.nf$lam   <- exp(tmp$fit)
# pred.dd.overall.bins.nf$lower <- exp(tmp$fit - 1.96 * tmp$se.fit)
# pred.dd.overall.bins.nf$upper <- exp(tmp$fit + 1.96 * tmp$se.fit)
# 
# pred.dd.overall.bins.fish$lam   <- exp(tmp.fish$fit)
# pred.dd.overall.bins.fish$lower <- exp(tmp.fish$fit - 1.96 * tmp.fish$se.fit)
# pred.dd.overall.bins.fish$upper <- exp(tmp.fish$fit + 1.96 * tmp.fish$se.fit)
# 
# pred.dd.overall <- rbind(pred.dd.overall.bins.fish,pred.dd.overall.bins.nf)
# 
# ggplot(pred.dd.overall,aes(y=lam,x=prop.abund,group=group,color=group)) + 
#                                                   geom_line(linewidth=2) + 
#                                                   geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.2) + 
#                                                   geom_hline(yintercept=1,linetype='dashed') + 
#                                                   xlab("Proportion of Maximum Abundance") + ylab("Lambda")
# 
# # OK, so if we compare the 40-60 bin with the 80-100 bin we should see the lambda's in the 40-60 bin are higher...
# 
# res.40.60 <- pred.dd.stock.bins |> collapse::fsubset(prop.bin %in% c("40-60%"))
# names(res.40.60)[names(res.40.60) == 'lam.no.fish'] <- "lam.40.60"
# res.80.100 <- pred.dd.stock.bins |> collapse::fsubset(prop.bin %in% c("80-100%"))
# names(res.80.100)[names(res.80.100) == 'lam.no.fish'] <- "lam.80.100"
# 
# tst <- left_join(res.40.60,res.80.100, by = names(res.40.60)[c(2,6:26)])
# tst$diff <- tst$lam.40.60 - tst$lam.80.100
# 
# hist(tst$diff)
# 
# res.lt.10 <- pred.dd.stock.bins |> collapse::fsubset(prop.bin %in% c("< 10%"))
# names(res.lt.10)[names(res.lt.10) == 'lam.no.fish'] <- "lam.lt.10"
# 
# tst <- left_join(res.40.60,res.lt.10, by = names(res.40.60)[c(2,6:26)])
# tst$diff <- tst$lam.lt.10 - tst$lam.40.60
# 
# hist(tst$diff)
# 
# res.10.20 <- pred.dd.stock.bins |> collapse::fsubset(prop.bin %in% c("10-20%"))
# names(res.10.20)[names(res.10.20) == 'lam.no.fish'] <- "lam.10.20"
# 
# tst <- left_join(res.40.60,res.10.20, by = names(res.40.60)[c(2,6:26)])
# tst$diff <- tst$lam.10.20 - tst$lam.40.60
# 
# hist(tst$diff)
# 
# 
# res.20.30 <- pred.dd.stock.bins |> collapse::fsubset(prop.bin %in% c("20-30%"))
# names(res.20.30)[names(res.20.30) == 'lam.no.fish'] <- "lam.20.30"
# 
# tst <- left_join(res.40.60,res.20.30, by = names(res.40.60)[c(2,6:26)])
# tst$diff <- tst$lam.20.30 - tst$lam.40.60
# 
# hist(tst$diff)
# 
# res.30.40 <- pred.dd.stock.bins |> collapse::fsubset(prop.bin %in% c("30-40%"))
# names(res.30.40)[names(res.30.40) == 'lam.no.fish'] <- "lam.30.40"
# 
# tst <- left_join(res.40.60,res.30.40, by = names(res.40.60)[c(2,6:26)])
# tst$diff <- tst$lam.30.40 - tst$lam.40.60
# 
# hist(tst$diff)

# 
# 
# dd.dat.no.bigs <- dd.dat |> collapse::fsubset(lam.no.fish < 2)
# dd.dat.six.no.bigs <- dd.dat.six |> collapse::fsubset(lam.no.fish < 2)
# just.bigs <- dd.dat  |> collapse::fsubset(lam.no.fish > 2) 
# sum(table(just.bigs$Stock.short)) # 48 of 68 are Haddock or cod. 13 of remaining 20 are small low trophic species
# 
# dd.dat$lam.ratio <- dd.dat$lam.org.no.fish/dd.dat$lam.no.fish
# dd.dat$lam.ratio.z <- dd.dat$lam.org.fish/dd.dat$lam.fish
# So the original lambda tended to be somewhere between 7 and 14% higher, depending on me(di)an choice.
# Could be way off in some years for a few species (Haddock, Cod, Sand lance, horse makerel, pilchard(sardine)... interesting... )
# summary(dd.dat$lam.ratio)
# summary(dd.dat$lam.ratio.z)
# ggplot(dd.dat,aes(y=lam.ratio,x=Species)) + geom_violin() + scale_y_log10() + geom_hline(yintercept =1) + theme_few(base_size=9)
# ggplot(dd.dat,aes(y=lam.ratio.z,x=Species)) + geom_violin() + scale_y_log10() + geom_hline(yintercept =1) + theme_few(base_size=9)

# Playing around to look at explanations for what we see, looked by LME, low and high stocks, but no smoking guns!
# low.stocks <- unique(dd.dat$Stock.short[dd.dat$prop.abund <= 0.1])
# dd.lows <- dd.dat |> collapse::fsubset(Stock.short %in% low.stocks)
# dd.highs <- dd.dat |> collapse::fsubset(!Stock.short %in% low.stocks)
# mod.dd.stock.gam.lows <- gam(log(lam.no.fish) ~ s(prop.abund,k=4) + s(prop.abund,by=as.factor(Stock.short),k=3),data=dd.lows )
# mod.dd.stock.gam.highs <- gam(log(lam.no.fish) ~ s(prop.abund,k=4) + s(prop.abund,by=as.factor(Stock.short),k=3),data=dd.highs )
# 
# pred.dd.stock.gam.tst <- expand.grid(prop.abund = seq(0, 1, length.out = 50),
#                                          Stock.short = c("Overall",unique(dd.highs$Stock.short)))
# 
# tmp <- predict(mod.dd.stock.gam.highs, pred.dd.stock.gam.tst, type = 'response', se.fit = TRUE) 
# pred.dd.stock.gam.tst$lam.no.fish   <- exp(tmp$fit)
# pred.dd.stock.gam.tst$lower <- exp(tmp$fit - 1.96 * tmp$se.fit)
# pred.dd.stock.gam.tst$upper <- exp(tmp$fit + 1.96 * tmp$se.fit)
# 
# ggplot(pred.dd.stock.gam.tst |> collapse::fsubset(Stock.short=='Overall'), aes(x=prop.abund,y=lam.no.fish)) + 
#                                 geom_point(data = hmm,aes(x=prop.abund,y=lam.no.fish),size=2,alpha=0.1) +  
#                                 geom_line(color='blue',linewidth=2) + 
#                                 xlab("Proportion of maximum abundance") + 
#                                 scale_y_log10(name = "Lambda",breaks = log_breaks(n=8)) +
#                                 geom_hline(yintercept = 1,color='grey',linetype = 'dashed')+
#                                 theme(legend.text=element_text(size=8),
#                                       legend.spacing.y = unit(-0.1, 'cm'),
#                                       legend.title = element_blank(),
#                                       axis.text.x = element_text(size=7)) + 
#                                  guides(color = guide_legend(ncol=1,byrow=T),
#                                         shape = guide_legend(ncol=1,byrow=T))
# 
# 
# ggplot(pred.dd.stock.gam.tst,aes(x=prop.abund,y=lam.no.fish,group=Stock.short,color=Stock.short,shape=Stock.short)) + 
#                                 geom_line() + 
#                                 #geom_point(data = thin.pred,size=2) +  
#                                 xlab("Proportion of maximum abundance") + 
#                                 scale_y_log10(name = "Lambda",breaks = log_breaks(n=8)) +
#                                 #facet_wrap(~Area.long)  +
#                                 #scale_shape_manual(values=pty.gam) +
#                                 #scale_color_manual(values = cols.gam) + 
#                                 geom_hline(yintercept = 1,color='grey',linetype = 'dashed')+
#                                 theme(legend.text=element_text(size=8),
#                                       legend.spacing.y = unit(-0.1, 'cm'),
#                                       legend.title = element_blank(),
#                                       axis.text.x = element_text(size=7)) + 
#                                  guides(color = guide_legend(ncol=1,byrow=T),
#                                         shape = guide_legend(ncol=1,byrow=T))

# Is there any density dependence in fishing mortality??
# dd.dat$mu <- dd.dat$removals/dd.dat$vpa.abund
# dd.dat$fm <- -log(1-dd.dat$mu)
# 
# ggplot(dd.dat , aes(x=prop.abund,y=fm)) + 
#                                 geom_point() + 
#                                 geom_smooth(method = 'gam') + 
#                                 scale_x_continuous(name="Proportion of Maximum Abundance") +
#                                 scale_y__continuous(name = "Fishing Mortality (instantaneous)")
# 
# 
# qqnorm(log(dd.dat$mu))
# qqline(log(dd.dat$mu))
# hist(log(dd.dat$mu))
# 
# fm.mod <- lm(log(fm)~prop.abund,data=dd.dat)
# par(mfrow=c(2,2))
# plot(fm.mod)
# par(mfrow=c(1,1))

# windows(11,11)
# col.try <- rep(rep(c("black","blue","orange",'firebrick2','lightgreen'),3),20)
# pty <- rep(c(rep(21,5),rep(22,5),rep(23,5)),20)

# 
# r0.ts.plt <- ggplot(pop.dam.6,aes(y=R0.opt,x=cohort,group=Stock.short,color=Stock.short,shape = Stock.short,fill = Stock.short)) + 
#                             geom_line(size=2)  + geom_point(size=4) +
#                             scale_y_log10(name = "Lifetime Reproductive Output") +
#                             facet_wrap(~G.Species) +
#                             scale_color_manual(name = "Stock",values = col.try) + scale_shape_manual(name = "Stock",values = pty) +
#                             scale_fill_manual(name = "Stock",values = col.try)+
#                             guides(colour = guide_legend(ncol = 1))+
#                             theme(axis.text.x = element_text(angle =45,hjust=1),legend.key.width = unit(5,'cm')) 
# r0.ts.plt
# 
# save_plot("D:/Github/ICM/Figures/tuned/R0_timeseries_6.png",r0.ts.plt,base_width = 30,base_height = 12)
# 
# r0.boxplt <- ggplot(pop.dam) + geom_boxplot(aes(y=R0.opt,x=Stock.no.species))  + 
#                             scale_y_log10(name = "Lifetime Reproductive Output")+#, breaks = c(0.001,0.01,0.1,1,5,10,50,100),minor_breaks = NULL) + 
#                             facet_wrap(~G.Species,scales = 'free')+
#                             theme(axis.text.x = element_text(angle =45,hjust=1),legend.position = 'none')
# r0.boxplt

# 
# windows(11,11)
# ggplot(pop.dam) + geom_boxplot(aes(y=gen.len.opt,x=Species)) + facet_wrap(~LME,scales='free_x') +
#                   ylab("Generation Length (years)") + theme(axis.text.x = element_text(angle =45,hjust=1))
# 
# windows(11,11)
# col.try <- rep(rep(c("black","blue","orange",'firebrick2','lightgreen'),3),20)
# lty <- rep(c(rep(1,5),rep(2,5),rep(4,5)),20)
# 
# gen.ts.plt <- ggplot(pop.dam.6,aes(y=gen.len.opt,x=cohort,group=Stock.short,color=Stock.short,linetype = Stock.short)) + 
#                             geom_line(size=2)  + 
#                             scale_y_log10(name = "Generation Length") +
#                             facet_wrap(~G.Species) +
#                             scale_color_manual(name = "Stock",values = col.try) + scale_linetype_manual(name = "Stock",values = lty) +
#                             guides(colour = guide_legend(ncol = 1))+
#                             theme(axis.text.x = element_text(angle =45,hjust=1),legend.key.width = unit(5,'cm')) 
# gen.ts.plt
# save_plot("D:/Github/ICM/Figures/tuned/Gen_time_series_6.png",gen.ts.plt,base_width = 20,base_height = 12)

# 
# ggplot(for.tunes) + geom_jitter(aes(y=lambda,x=Species,group=Stock.short,color=Stock.short)) + 
#                     scale_y_continuous(name = 'rate of growth (lambda)',breaks = c(-1,-0.5,0,0.5,1:10)) +
#                             theme(axis.text.x = element_text(angle =45,hjust=1))
# 
# ggplot(for.tunes) + geom_jitter(aes(y=lambda.vpa.init,x=Species)) + scale_y_continuous(name = 'rate of growth (lambda)',breaks = c(-1,-0.5,0,0.5,1:10),limits = c(-1,3))
# 
# unique(for.tunes$Species)
# windows(11,11)
# 
# ggplot(for.tunes) + geom_density(aes(lambda,group=Species,color=Species,fill=Species),alpha = 0.2) + scale_x_continuous(breaks = c(0,0.5,1,2,5,10))
# 
# 
# #options(scipen=999)
# r0.plt <- ggplot(pop.dam) + geom_boxplot(aes(y=R0.opt,x=Area))  + 
#                             scale_y_log10(name = "Lifetime Reproductive Output")+#, breaks = c(0.001,0.01,0.05,0.1,0.5,1,5,10,50,100),minor_breaks = NULL) + 
#                             facet_wrap(~Species,scales = 'free')+
#                             theme(axis.text.x = element_text(angle =45,hjust=1))
# save_plot("D:/Github/ICM/Figures/tuned/R0_ignoring_fishing.png",r0.plt,base_width = 8,base_height = 8)
# 
# 
# 
# vickis.cols <- sort(rep(c( "#172582",  "#F8DC91", "#FBB891", "#FB6A01", "#B80000"),3))
# linetypes <- rep(c(1,2,4),5)
# 
# windows(11,11)
# cod.r0.ts.plt <- ggplot(pop.dam |> collapse::fsubset(Species == 'morhua')) + geom_line(aes(y=R0.opt,x=cohort,group=Area,color=Area,linetype=Area),size=1.5)  + 
#                             scale_y_log10(name = "Lifetime Reproductive Output", breaks = c(0.001,0.01,0.1,1,5,10,50,100),minor_breaks = NULL) + 
#                             scale_color_manual(values = vickis.cols) +
#                             scale_linetype_manual(values = linetypes) +
#                             theme(legend.key.width  = unit(2, "cm"))
# cod.r0.ts.plt
# #save_plot("D:/Github/ICM/Figures/tuned/R0_timeseries_ignoring_fishing.png",r0.plt,base_width = 8,base_height = 8)
# 
# 
# windows(11,11)
# ggplot(pop.dam |> collapse::fsubset(Species =="morhua")) + geom_boxplot(aes(y=R0.opt,x=Area))  + 
#                                                         scale_y_log10(name = "Lifetime Reproductive Output",breaks = c(0.001,0.01,0.1,1,5,10,50,100),minor_breaks = NULL) +
#                                                         theme(axis.text.x = element_text(angle =45,hjust=1))
# 
# # Missing catch questions
# # first did we do much to the Fecundity term to get here
# quantile(for.tunes$mean.fec/for.tunes$mean.vpa.fec,probs=seq(0.01,1,0.01),na.rm=T)
# 
# quantile((for.tunes$removals-for.tunes$removals.init)/for.tunes$removals.init,probs=seq(0.01,1,0.01),na.rm=T)
# quantile((for.tunes$mean.nm-for.tunes$mean.vpa.nm)/for.tunes$mean.vpa.nm,probs=seq(0.01,1,0.01),na.rm=T)
# 
# for.tunes$excess.catch <- for.tunes$removals-for.tunes$removals.init
# for.tunes$per.excess.catch <- 100*((for.tunes$removals-for.tunes$removals.init)/for.tunes$removals.init)
# 
# windows(11,11)
# bycatch.plt <- ggplot(for.tunes) + geom_boxplot(aes(y=per.excess.catch,x=Species)) + 
#                                    ylab("Adjustment to catch (%)") +
#                                    theme(axis.text.x = element_text(angle=45,hjust = 1))
#                                    
# save_plot("D:/Github/ICM/Figures/tuned/Percentage_change_to_bycatch_by_species.png",bycatch.plt,base_width = 8,base_height = 8)
#                                    
# bycatch.plt.cod <- ggplot(for.tunes |> collapse::fsubset(Species == 'morhua')) + geom_boxplot(aes(y=per.excess.catch,x=Area)) + 
#                                                                               ylab("Adjustment to catch (%)") +
#                                                                               theme(axis.text.x = element_text(angle=45,hjust = 1))
#                                                                               
# save_plot("D:/Github/ICM/Figures/tuned/Percentage_change_to_bycatch_by_cod_stock.png",bycatch.plt.cod,base_width = 8,base_height = 8)
# 
# # bycatch.plt <- ggplot(for.tunes) + geom_boxplot(aes(y=per.excess.catch,x=Area)) + facet_wrap(~Species,scales = 'free_x')
# #                                    ylab("Adjustment to catch (%)") +
# #                                    theme(axis.text.x = element_text(angle=45,hjust = 1))
# # bycatch.plt                                   
#

# Doubling time for the rest of the stocks...

# dub.heatmap.no.6 <- ggplot(dub.time.summary |> collapse::fsubset(!Species %in% six.species), aes(x=year, y=Stock.short, fill= prop)) + 
#                               xlab("") + ylab("")+                                    
#                               #facet_wrap(~G.Species,nrow=3,scales='free_y',strip.position = "right") + 
#                               scale_x_continuous() +
#                               geom_tile() + scale_fill_viridis(name = "Proportion",discrete=FALSE) + 
#                               theme(legend.position = 'none',text = element_text(size=12),plot.margin = margin(0,0.5,0,0.68, "cm"))
```

<!-- Here are the models that we ran but didn't end up using# -->
<!-- Here are the models that we ran but didn't end up using# -->
<!-- Here are the models that we ran but didn't end up using# -->
<!-- Here are the models that we ran but didn't end up using# -->
<!-- Here are the models that we ran but didn't end up using# -->
<!-- Here are the models that we ran but didn't end up using# -->
<!-- Here are the models that we ran but didn't end up using# -->

```{r failed-models}

# A density dependence model, start simple and build...
# mod.dd.base <- lm(log(lam.no.fish) ~ prop.abund,data=dd.dat)
# summary(mod.dd.base)
# 
# mod.dd.base.fish <- lm(log(lam.fish) ~ prop.abund,data=dd.dat)
# summary(mod.dd.base.fish)
# #ggplot(dd.dat, aes(y = lam.no.fish,x=prop.abund)) + geom_point() + geom_smooth(method = 'lm') + scale_y_log10() + geom_hline(yintercept = 1,color='firebrick2',linetype='dashed',linewidth=2)
# 
# 
# # Residuals are nowhere near normal, the fit isn't a disaster but def have the old trumpet shape happening
# par(mfrow=c(2,2))
# plot(mod.dd.base)
# par(mfrow=c(1,1))
# # This mostly represents the above model...
# ggplot(dd.dat, aes(y = lam.no.fish,x=prop.abund)) + 
#                   geom_point() + 
#                   geom_smooth(method = 'lm') + scale_y_log10() +
#                   geom_hline(yintercept = 1,color='firebrick2',linetype='dashed',linewidth=2)
# # What happens with a gam?
# mod.dd.gam.base <- gam(log(lam.no.fish) ~ s(prop.abund),data=dd.dat)
# summary(mod.dd.gam.base)
# # Still see that the residuals are higher at low values
# plot(residuals(mod.dd.gam.base) ~ dd.dat$prop.abund)
# # not great...
# plot(residuals(mod.dd.gam.base) ~ fitted(mod.dd.gam.base))
# # What's the fit look like?
# pred.dd.gam.base <- data.frame(prop.abund = seq(0, 1, length.out = 1000))
# tmp <- predict(mod.dd.gam.base, dd.gam.base.pred, type = 'response', se.fit = TRUE) 
# 
# pred.dd.gam.base$lam.no.fish   <- exp(tmp$fit)
# pred.dd.gam.base$lower <- exp(tmp$fit - 1.96 * tmp$se.fit)
# pred.dd.gam.base$upper <- exp(tmp$fit + 1.96 * tmp$se.fit)
# 
# # So here's the game fit, I thnk it does do an ok job of showing that no much is happening below maybe 0.3, but then
# # below 0.15 it starts increasing, suggesting general compensation, but we know the really high values matter here and are
# # limited to a subset of the stocks, anywho we know the residuals suck, so moving on...
# ggplot(dd.dat,aes(y=lam.no.fish,x=prop.abund)) + 
#                   geom_point() + scale_y_log10() +
#                   geom_ribbon(data=pred.dd.gam.base,aes(ymin=lower,ymax=upper,x=prop.abund),alpha=0.4) +
#                   geom_line(data = pred.dd.gam.base,aes(y=lam.no.fish, x= prop.abund))
#     
# 
# # What if we did a glm, Gamma with log-link seems a natural thing to try...
# mod.dd.glm.gamma <- glm(lam.no.fish ~ prop.abund,data=dd.dat,family = Gamma(link='log'))
# # What's that done...
# summary(mod.dd.glm.gamma)
# # Still some pretty gross shit going on here...
# par(mfrow=c(2,2))
# plot(mod.dd.glm.gamma)
# par(mfrow=c(1,1))
# # Now plot it...
# pred.dd.glm.base <- data.frame(prop.abund = seq(0, 1, length.out = 1000))
# tmp <- predict(mod.dd.glm.gamma, dd.gam.base.pred, type = 'response', se.fit = TRUE) 
# 
# pred.dd.glm.base$lam.no.fish   <- (tmp$fit)
# pred.dd.glm.base$lower <- (tmp$fit - 1.96 * tmp$se.fit)
# pred.dd.glm.base$upper <- (tmp$fit + 1.96 * tmp$se.fit)
# # Pretty similar, though I think on the linear scale we see the 'rate' may be decreasing as we approach 0
# # But still not a good model given the above
# ggplot(dd.dat,aes(y=lam.no.fish,x=prop.abund)) + 
#                   geom_point() + scale_y_log10() +
#                   geom_ribbon(data=pred.dd.glm.base,aes(ymin=lower,ymax=upper,x=prop.abund),alpha=0.4) +
#                   geom_line(data = pred.dd.glm.base,aes(y=lam.no.fish, x= prop.abund))
#     
# # Now what if we account for species in these and get an estimate for each species. 
# # I think this should be done as a random effect, we'll start with linear mixed models and see what happens...
# 
# mod.dd.species.re <- lmer(log(lam.no.fish) ~ prop.abund + (1| Species),data=dd.dat)
# # So that seems reasonable...
# summary(mod.dd.species.re)
# #Residuals are still pretty cranky
# plot(mod.dd.species.re)
# # Normality still a hot mess
# qqnorm(residuals(mod.dd.species.re))
# qqline(residuals(mod.dd.species.re))
# # draw some lines...
# pred.dd.lmm.species <- expand.grid(prop.abund = seq(0, 1, length.out = 100),Species=unique(dd.dat$Species))
# tmp <- predict(mod.dd.species.re, pred.dd.lmm.species, type = 'response', se.fit = TRUE) 
# pred.dd.lmm.species$lam.no.fish   <- exp(tmp$fit)
# pred.dd.lmm.species$lower <- exp(tmp$fit - 1.96 * tmp$se.fit)
# pred.dd.lmm.species$upper <- exp(tmp$fit + 1.96 * tmp$se.fit)
# # Here are the fits (sans SE cause messy + didn't put the overall relationship on here)
# ggplot(pred.dd.lmm.species,aes(x=prop.abund,y=lam.no.fish,group=Species,color=Species)) + 
#                                 geom_line()  + 
#                                 geom_point(data=dd.dat,aes(x=prop.abund,y=lam.no.fish,group=Species, color=Species)) +
#                                 scale_y_log10()
# 
# # We could also allow the slope to vary by species, might get a bit better, but really won't do the trick...
# mod.dd.species.slope.re <- lmer(log(lam.no.fish) ~ prop.abund + (1+ prop.abund| Species),data=dd.dat)
# summary(mod.dd.species.slope.re)
# #Residuals are still pretty cranky
# plot(mod.dd.species.slope.re)
# # Normality still a hot mess
# qqnorm(residuals(mod.dd.species.slope.re))
# qqline(residuals(mod.dd.species.slope.re))
# # draw some lines...
# pred.dd.lmm.species.slope <- expand.grid(prop.abund = seq(0, 1, length.out = 100),Species=unique(dd.dat$Species))
# tmp <- predict(mod.dd.species.slope.re, pred.dd.lmm.species.slope, type = 'response', se.fit = TRUE) 
# pred.dd.lmm.species.slope$lam.no.fish   <- exp(tmp$fit)
# pred.dd.lmm.species.slope$lower <- exp(tmp$fit - 1.96 * tmp$se.fit)
# pred.dd.lmm.species.slope$upper <- exp(tmp$fit + 1.96 * tmp$se.fit)
# # Here are the fits (sans SE cause messy + didn't put the overall relationship on here)
# ggplot(pred.dd.lmm.species.slope,aes(x=prop.abund,y=lam.no.fish,group=Species,color=Species)) + 
#                                 geom_line()  + 
#                                 geom_point(data=dd.dat,aes(x=prop.abund,y=lam.no.fish,group=Species, color=Species)) +
#                                 scale_y_log10()
# # Curious what species have high slopes.. Dubious and Esmarkii (North Sea sand lance and mackerel)
# ranef(mod.dd.species.slope.re)
# 
# # So so what if we let the slopes vary by stock
# mod.dd.stock.re <- lmer(log(lam.no.fish) ~ prop.abund +  (1+ prop.abund | Stock.short),data=dd.dat)
# summary(mod.dd.stock.re)
# #Residuals are slightly better I think but still really suck
# plot(mod.dd.stock.re)
# # Normality = no
# qqnorm(residuals(mod.dd.stock.re))
# qqline(residuals(mod.dd.stock.re))
# # What's it look like?
# pred.dd.lmm.stock <- expand.grid(prop.abund = seq(0, 1, length.out = 100),
#                                          #Species=unique(dd.dat$Species),
#                                          Stock.short=unique(dd.dat$Stock.short))
# tmp <- predict(mod.dd.stock.re, pred.dd.lmm.stock, type = 'response', se.fit = TRUE) 
# pred.dd.lmm.stock$lam.no.fish   <- exp(tmp$fit)
# pred.dd.lmm.stock$lower <- exp(tmp$fit - 1.96 * tmp$se.fit)
# pred.dd.lmm.stock$upper <- exp(tmp$fit + 1.96 * tmp$se.fit)
# # Here are the fits (sans SE cause messy + didn't put the overall relationship on here)
# ggplot(pred.dd.lmm.stock,aes(x=prop.abund,y=lam.no.fish,group=Stock.short,color=Stock.short)) + 
#                                 geom_line()  + 
#                                 geom_point(data=dd.dat,aes(x=prop.abund,y=lam.no.fish,group=Stock.short, color=Stock.short)) +
#                                 scale_y_log10()
# 
# # Curious what stocks have high slopes.. Dubious and Esmarkii (North Sea sand lance and mackerel), NS Haddock and Irish Sea Haddock 
# ranef(mod.dd.stock.re)
# 
# 
# 
# 
# # OK, so this all still sucks, wonder if a glmer and gamma would get us anywhere... not really.... 
# # Tried many variants on the glmer and they all land in the same lousy spot.
# mod.dd.stock.gamma.re <- glmer(lam.no.fish ~ prop.abund +  (1+ prop.abund | Stock.short),data=dd.dat,family=Gamma(link='log'))
# # OK...
# summary(mod.dd.stock.gamma.re)
# #Residuals, nah, still suck, just some crazy outliers...
# plot(mod.dd.stock.gamma.re)
# # Normality = no
# qqnorm(residuals(mod.dd.stock.gamma.re))
# qqline(residuals(mod.dd.stock.gamma.re))
# # And the plot...
# pred.dd.lmm.stock.gamma <- expand.grid(prop.abund = seq(0, 1, length.out = 100),
#                                          #Species=unique(dd.dat$Species),
#                                          Stock.short=unique(dd.dat$Stock.short))
# tmp <- predict(mod.dd.stock.gamma.re, pred.dd.lmm.stock.gamma, type = 'response', se.fit = TRUE) 
# pred.dd.lmm.stock.gamma$lam.no.fish   <- (tmp$fit)
# pred.dd.lmm.stock.gamma$lower <- (tmp$fit - 1.96 * tmp$se.fit)
# pred.dd.lmm.stock.gamma$upper <- (tmp$fit + 1.96 * tmp$se.fit)
# # Here are the fits (sans SE cause messy + didn't put the overall relationship on here)
# ggplot(pred.dd.lmm.stock.gamma,aes(x=prop.abund,y=lam.no.fish,group=Stock.short,color=Stock.short)) + 
#                                 geom_line()  + 
#                                 geom_point(data=dd.dat,aes(x=prop.abund,y=lam.no.fish,group=Stock.short, color=Stock.short)) +
#                                 scale_y_log10()
# # None of the above really work at all....

```

<!--chapter:end:tmp-01-analysis_no_age_correction.Rmd-->

# INTRODUCTION  {#intro}

Intro – life history paragraph, environmental variation paragraph, management ties paragraph, goals.

theoretical expectations related to productivity and life history for various species (e.g. herring vs. Cod) (ties to regional variation in lifetime reproductive rates and doubling time).

Evidence for specific species to make use of environmental conditions – boom in recruitment.

How our understanding of productivity is used in a management context – projections, reference points, risk, (ties directly to density dependence figure).  
Species-environment relationship – is it different from what we expect and how does that influence our capacity for management of exploited fishes.

The under appreciated influence of regional variation in determining recovery potential of exploited teleosts.  - can’t look at a single species and expect it will behave the same way.

Objectives
  - Quantify the life history variability of exploited marine teleosts
  - Explore regional and taxonomic correlates of life history variability
  - Based on the observed life history, quantify the recovery potential of marine teleosts in the absence of fishing

<!--chapter:end:tmp-02-intro.Rmd-->

# Methods {#meth}

## Data

This analysis used age structured fisheries data obtained from stock assessments for stocks in the Atlantic and Pacific Oceans [@charbonneauAgestructuredmarinefishdatabase2022; @charbonneauEffectsFishingMortality2022]. This repository had sufficient data to run the analysis for 77 stocks from NE Atlantic, NW Atlantic, and the North Pacific, this included stocks managed by International Council for the Exploration of the Sea (ICES; `r n.stock.ices`), NOAA (NOAA;`r n.stock.noaa`), and Fisheries and Oceans Canada (DFO; `r n.stock.dfo` stocks). The time series used in this analysis started as early as `r min.year` and the most recent data were from `r max.year`, the average length of the time series for the stocks included in this analysis was `r mn.ts.len` years and the number of years used ranged between `r min.ts.len` and `r max.ts.len` years. The stocks used represented `r num.orders` orders, with the most represented orders being the Gadiformes (`r num.gads` stocks), Pleuronectiformes (`r num.pleur` stocks) and Clupeiformes (`r num.clup` stocks). The other three orders had relatively few stocks included in the analysis, with `r num.perci` Perciform stocks, `r num.scomb` Scombriform stocks, and just `r num.scorp` Scorpaeniform stock.  The most common species in this analysis were Atlantic cod (*G. morhua*) and the majority of the stocks (`r n.six`) in this analysis were represented by one of six species (*G. morhua* (`r num.cod`), *C. harengus* (`r num.her`), *M. aeglefinus* (`r num.had`),  *P. virens* (`r num.pol`), *S. solea* (`r num.sol`), and *P. platessa* (`r num.plat`)).

These data included age-specific estimates of abundance, weight, biomass, maturity, natural mortality, and fishery removals (Table \@ref(tab:tab-meta)), these data were all required to calculate survival-at-age and fecundity-at-age and to estimate the annual productivity of these stocks. For the age structured analysis, the last age class was considered the 'terminal' age in the analysis, thus we implicitly assumed there was no contribution to productivity from older aged individuals. In addition, for the majority of the stocks the first age class was age 1 (`r n.age.1` stocks) while another `r n.age.2.plus` of the stocks have a first age of 2 or greater. The young age classes not included for these stocks (e.g. Age 0) consisted of immature individuals that do not contribute to the productivity of the stock; however it has been noted that the exclusion of immature age classes in these types of analyses may lead to a slight overestimate of productivity [@pardoMaximumIntrinsicRate2016].

## Life Table Reconstruction

The discrete time Euler-Lotka  equation was first proposed by Euler in 1760 (subsequently published by @moserGesetzeLebensdauer1839) and is a special case of the continuous time demographic model first proposed by Lotka in 1907 [@lotkaRelationsBirthRates1907; @sharpeProblemAgedistribution1911; @lotkaRelationBirthRate1918]:

$$
1 = \sum_{x=\alpha}^{\omega} \lambda^{-x} s_x f_x dx
$$

Where $x$ is the age, $\alpha$ is the age at first reproduction, $\omega$ is the maximum age, $\lambda$ is the discrete rate of population growth ($\lambda = e^r$), $s_x$ is the survivorship to age $x$, and $f_x$ is the fecundity of individuals of age $x$. Note that the discrete time Euler-Lotka equation can be shown to be the characteristic polynomial of the Leslie Matrix, where $f_0$ is the fecundity of the first age class with mature individuals [@leslieUseMatricesCertain1945, @caswellMatrixPopulationModels2001]. The available stock assessment data are used to calculate survival-at-age and fecundity-at-age to estimate $\lambda$, assuming that each stock represents a discrete population with individual dynamics.

<!-- \begin{align} -->
<!--   \begin{bmatrix} -->
<!--     f_0  & f_1 & f_2  & f_3    & ...          & f_{\omega-1} \\ -->
<!--     s_0  & 0   & 0    & 0      & ...          & 0 \\ -->
<!--     0    & s_1 & 0    & 0      & ...          & 0 \\ -->
<!--     0    & 0   & s_2  & 0      & ...          & 0 \\ -->
<!--     0    & 0   & 0    & \ddots & ...          & 0 \\ -->
<!--     0    & 0   & 0    & ...    & s_{\omega-2} & 0 \\ -->
<!--   \end{bmatrix} -->
<!-- \end{align} -->

### Fecundity-at-age

The number of spawners in year $y$ and age class $a$ ($SSN_{y,a}$) is the proportion of mature individuals ($Mat_{y,a}$) multiplied by the number of individuals in each age

$$
SSN_{y,a} = Mat_{y,a} N_{y,a}
$$
The Euler-Lotka equation requires fecundity to be the number of offspring produced by each female in each age class. Given the uncertainty of both fertilization and egg survival, using the number of eggs at a given length or size is inappropriate.  However, these data do have the number of recruits produced each year, and this can be combined with the spawning stock biomass to obtain an estimate of the number of recruits produced per kilogram of spawning biomass $RPS_{y}$, this is calculated as

$$
RPS_{y} = R_{y} SSB_{tot,(y-ra)} 
$$

Where $SSB_{yot,y-ra}$ is the total spawning stock biomass in year $y$. $R_{y}$ is the number of individuals in the youngest age class available for a specific stock (hereafter referred to as recruits) offset temporally by the age of the recruits. For example, if the age of the recruits ($ra$) is 2 years , we assume that the recruits in 2010 were produced by the spawning stock biomass in 2008 ($SSB_{tot,y-ra}$).  The number of recruits produced by each age class ($RPA_{y,a}$) can then be calculated by multiplying the spawner biomass in each age class $SSB_{y-ra,a}$ by the number of recruits produced per kilogram of SSB

$$
RPA_{y,a} =  SSB_{(y-ra),a} RPS_{y}
$$
This method does not account for any changes in reproductive success by age, each kilogram of spawning biomass contributes the same to the total total reproductive output, irrespective of their age. Finally, the age-specific fecundity ($Fec_{y,a}$) is calculated as number of offspring produced by each individual in the population. This is offset by one year, as the recruits observed in year $y+1$ enter into the Euler-Lotka formulation in year $y$.

$$
Fec_{y,a} =  \frac{RPA_{y+1,a}}{SSN_{(y-ra+1),a}}
$$
### Suvival-at-age

The number of animals surviving to each age depends on age-specific mortality rates. Commercially-exploited fishes experience fishing mortality (F) once they grow large enough to become vulnerable to the gear (reference). Fishing mortality combines with natural mortality (M) to give total mortality (Z) affecting each age class in each year. Varying assumptions can be made about natural mortality during stock assessments, where it may be an assumed value for each age (EXAMPLE), allowed to vary over time and by age (EXAMPLE), or input as a constant across ages (EXAMPLE).  Due to challenges with separability in typical stock assessment models, natural mortality is often assumed for each age class, in some assessments it varies by age and over time, in other assessments it does not.  Irrespective of this, the stock assessment used in this analysis all provide an estimate of natural mortality by age and year, this value is typically an instantaneous rate and is convert to a proportional natural mortality for the Euler-Lotka model.

The data also contains the number of fish harvested by the fishery each year and for each age. This can be combined with the assessed numbers at age to get an estimate of fishing mortality in each year for each age.

$$
FM_{y,a} =  \frac{Catch_{y,a}}{N_{y,a}}
$$
This is combined with the natural mortality to get a total mortality estimate ($Z_{y,a}$) for each age in each year

$$
Z_{y,a} =  FM_{y,a} + NM_{y,a}
$$
### Population Growth Rates $\lambda$

The fecundity, natural mortality, and fishing mortality estimates were then used to obtain an estimate of $\lambda$ for a given year from the Euler-Lotka equation and used to predict the abundance in the following year.

$$
N_{y+1} = \lambda_{y} N_{y}
$$
The Euler-Lotka model was fit using an iterative minimization process, where predicted abundance in the following year was compared to the abundance estimate from the stock assessment for the corresponding year ($N_{y+1}$). Where necessary, the annual age-specific fecundity and natural mortality estimates were adjusted until the $\lambda$ estimate from the Euler-Lotka equation resulted in an abundance time-series estimate ($N_{y+1}$) that was within 5% of the abundance estimates from the stock assessment (median difference was `r med.per.diff`). This tolerance ensured that our analyses reflected the same understanding of stock productivity, natural mortality and fecundity as in the current assessments; effectively this methodology treats the abundance estimates from the stock assessments as the 'truth'.

The life table reconstruction resulted in stock-specific $\lambda$, fecundity, and natural mortality estimates for each year in which data were available accounting for the age structure and life history characteristics of each stock. In other words, we estimated the realized population growth rate using the changes in the population size of the stocks from year to year (realized population growth rates), based on the amount of mortality experienced and annual reproductive success. The realized $\lambda$
included fishing mortality  ($\lambda_{F>0}$), although this method enabled the theoretical population growth rate in the absence of fishing  ($\lambda_{F=0}$) could also be calculated directly fomr the annual the natural mortality and fecundity estimates.

<!-- For consistency in the Euler-lotka estimates across all the stocks, the first age used in the analysis was 'age 0' (which is simply an estimate of abundance of fish in the year they were born). For some stocks this information is not available (see Table X, **possibly in an Appendix where we have meta-data**). For the stocks in which age 0 information is unavailable, an estimate of the abundance of the younger age classes were back calculated using average natural mortality provided for the youngest age class available. -->

## Density Dependence

Life table analyses make no assumptions about population regulation and thus do not incorporate a density-dependent relationship between spawners and recruits (Reference?). However, the data itself could be explored for evidence of density-dependence by modelling the estimated annual population growth rate against the proportion of maximum abundance observed that year for each stock. This was done twice, once using  ($\lambda_{F=0}$) and once using ($\lambda_{F>0}$).  Negative density dependence is observed when $\lambda$ increases as abundance declines, positive density dependence occurs when $\lambda$ declines with increasing abundance, while density independence is observed when $\lambda$ is not impacted by changes in abundance.

Generalized Additive Mixed Models were used to allow for some non-linearity in the relationships (CITE MCGV and WOOD), the overall effect and the effect for each stock was modelled as

\begin{align}
\lambda_{i,j} \sim N(0,\sigma^2_{i,j}) \\
E(\lambda_{i,j}) = \mu_{i,j} \\
log(\mu_{i,j}) =  Prop_{i} + Stock_{j}  \times Prop_{i} \\
Stocks_{i,j} \sim N(0,\sigma^2_{Stocks})
\end{align}

The interaction was modeled using the *fs* basis using the mgcv package in R, this treats the smooth curves as random effects and is used when a factor has a large number of levels (CITE!).  Various alternative modelling methods were explored (various model formulations, alternative error distributions, etc.) but all models had similar challenges overcoming either the underlying non-normality or the heteroskedasticity of the data. As such, the uncertainty surrounding the model output should be treated with caution, while we discuss the general trends this model provides we attempted to avoid discussion of the statistical significance of the trends.

## Demographic Parameters

There are numerous demographic parameters that can be derived from the outputs of the life table reconstruction. Here we focus on two that are strongly linked to recovery potential: lifetime reproductive success and generation time. Lifetime reproductive success describes the average number of spawners produced by one female spawner throughout her lifetime

## Lifetime Reproductive Success

The lifetime reproductive success ($R_0$)  was calculated for each cohort in which fecundity and natural mortality were available for all age classes. It was calculated using only natural mortality (F=0) and with the combined effect of natural mortality and fishing mortality (F>0).  This was calculated for each cohort as

$$
R_0 = \sum_{x=\alpha}^{\omega} s_x f_x
$$

*Here we are treating the plus group as effectively 'terminal', i.e., it assumes that after the last age class the individuals make no contribution, which means these values are probably a bit low, but at this age they really aren't making much of a contribution on average*.

## Generaion Time

The generation time was defined as the mean age of the spawners of the observed recruits in a cohort over its lifetime ($T_G$), here we calculated this for each cohort in which fecundity and natural mortality were available for all age classes. It was calculated twice, using only natural mortality (F=0) and with the combined effect of natural mortality and fishing mortality (F>0).

$$
T_G = \frac{\sum_{x=\alpha}^{\omega} x s_x f_x}{\sum_{x=\alpha}^{\omega}  s_x f_x}
$$

## Recovery Potential

The annual fecundity and natural mortality values  varied considerably for each stock. To account for this variability when evaluating recovery potential, we used a simulated population projection. For `r n.dt.sims` simulations, we took a random sample of the annual values for fecundity and natural mortality, used the life table reconstruction to estimate $\lambda$, and calculated abundance in the following year. Abundance of each stock was projected over `n.sim.years` years, and the percentage of the simulations in which the population size had doubled was recorded each year. For simplicity, we did not allow for density-dependent effects (e.g. potential population decline as abundance exceeds carrying capacity) because of uncertainty surrounding potential maximum population size for each species. Once a population had doubled, it was considered to have remained at that abundance or greater for the remainder of the projection.

<!--chapter:end:tmp-03-methods.Rmd-->

# Results {#reses-peses}

<!-- Throughout the results we will focus primarily the six most commonly found species in the analysis (*Gadus morhua, Melogrammus aeglefinus, Clupie harengus, Plueronecties platessa, Polluck virens* and *Solea Solea*).  The complete results for all of the stocks included in the analysis can be found in the Supplemental material. -->

## Population growth rates ($\lambda$)

The overall median population growth rate when excluding the effects of fishing ($\lambda_{F=0}$) was `r med.lambda.F0`, indicating that there was a latent potential for population growth for these stocks of approximately `r med.r.F0`%. When the effect of fishing was accounted for, the median population growth rate ($\lambda_{F>0}$) declined to `r med.lambda.F`, which suggested that the populations in this analysis may have been over-fished by approximately `r -med.r.F`% annually. However, for many of the stocks included in this analysis large recruitment events resulted in large estimates of $\lambda_{F>0}$, in some years, this led to a skewed distribution of the lambda estimates, with these large events being an integral component of the stocks population dynamics (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-bx-other-fig)). The mean $\lambda_{F>0}$ for the stocks in this analysis was `r mn.lambda.F` which would suggest there remains scope for population growth even after fishing. The mean$ \lambda_{F>0}$ was lower than the median for just `r n.stocks-n.mn.higher.med.fish` of the `r n.stocks` in this analysis. The majority of these stocks (`r n.nw.ice`) were in the central or NW Atlantic, and this included `r n.cod.low.mn`  of the `r n.cod` Atlantic cod stocks in this analysis.

There were regional and species specific trends in both $\lambda_{F=0}$ and $\lambda_{F}$. Haddock (*M. aeglefinus*) and Atlantic cod (*G. morhua*) stocks had some of highest variability in lambda, both at the stock level and between stocks (Figure \@ref(fig:lambda-sd-fig)). For these two species, the variability of lambda within a stock declined as you moved from the Eastern Atlantic to the Western Atlantic (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-sd-fig))). The pattern was more pronounced for Atlantic cod, with stocks in the eastern Atlantic tending to have relatively low within stock variability. The two other species with stocks found in the NW and NE Atlantic (*C.harengus* and *P. virens*) the variability of lambda between the stocks was much lower than for Haddock and Atlantic cod and there were no clear differences between regions (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-sd-fig))). This was largely due to the within stock lambda variability being lower than the within stock variability of Haddock and Atlantic cod stocks in the NE Atlantic (Figure \@ref(fig:lambda-bx-6-fig)). Two other stocks (*A. dubious* and *T. esmarkii*) in the North Sea had highly variably lambda estimates, though notably the lambda was less variable than the estimates for Haddock and Atlantic cod in this area (Figure \@ref(fig:lambda-bx-other-fig)). For the remainder of other stocks in this analysis, years in which the value of $\lambda_{F=0}$ exceeded 1.5 were more rare. Exceptions to this included *M. merlangus* in the Celtic Sea, *P. dentatus* and *S. Chrysops* in the NW Atlantic, and the Gulf of Alaska *T. chalcogramma* stock in the Pacific Ocean (Figure \@ref(fig:lambda-bx-other-fig)).

## Density Dependence

Overall, the evidence suggested that negative density dependence was common as both $\lambda_{F=0}$ and $\lambda_{F>0}$ increased as abundance declined (Figure \@ref(fig:dd-overall-plt)). The strength of density dependence appeared to be more pronounced at low and high abundance levels (i.e., changes in abundance had larger impacts on $\lambda$ when abundance was below $\approx$ 0.4 or above $\approx$ 0.7). $\lambda_{F=0}$ only declined below 1 when the stocks approached their maximum observed abundance ($\approx$ `r pa.lam.nf.at.one`%), indicating that in the absence of fishing there was a latent capacity for population growth in these stocks.  However, when accounting for fishing mortality, $\lambda_{F>0}$ was below 1 until the stocks were below approximately `r pa.lam.fish.at.one`% of their maximum abundance. At lower abundances, the rate of increase in $\lambda_{F>0}$ was higher than $\lambda_{F=0}$, likely due to reductions in fishing mortality at low abundance, but overall $\lambda_{F>0}$ was lower than $\lambda_{F=0}$, suggesting that some fishing mortality was common even at very low abundance (Figure \@ref(fig:dd-overall-plt)).

Generally, there was an increase in the variability of the $\lambda$ estimates as abundance declines, largely driven by an increase in the magnitude of the maximum value of $\lambda$ and a small increase in the frequency of low $\lambda$ values (Figure \@ref(fig:dd-overall-plt)). For example, values of $\lambda_{F>0}$ > `r pa.lam.fish.low.bigs` were not observed at abundances > `100*pa.big.threshold`% of the maximum (similar trends were observed for $\lambda_{F=0}$, with only one point being above `r pa.lam.nf.low.bigs` when the abundance was > `r 100*pa.big.threshold`% of the maximum). Similarly, the percentage of $\lambda_{F>0}$ values below `r lam.low.threshold` was `r 100*prop.low.high.abund`% when the proportion of the abundance was >= 0.3, but this increased to `r 100*prop.low.low.abund`% of the data when the proportion of the abundance was below 0.3.

However, these general patterns were highly variable between stocks. All of the stocks included in this analysis had $\lambda_{F=0}$ estimates > 1 at their lowest observed abundances (Table \@ref(tab:tab-lambdas) and Figure \@ref(fig:dd-nf-stock-plt)). Stocks in the Northwest Atlantic (particularly Atlantic cod) and Alaska had some of the lowest $\lambda_{F=0}$ estimates at low abundance, while several stocks in the North Sea had among the highest $\lambda_{F=0}$ estimates at low abundance (Table \@ref(tab:tab-lambdas)).  Of the `r n.bl.40` stocks that had declined below 40% of the maximum abundance, only `r `n.low.40.nf` of the stocks had a $\lambda_{F=0}$ estimates < 1 at 40% of their maximum abundance; when the effects of fishing were included `r `n.low.40.nf` of the stocks had $\lambda_{F>0}$ estimates below 1 (Table \@ref(tab:tab-lambdas)).

For the vast majority of stocks the lowest values of both $\lambda_{F=0}$ (`r n.low.lam.at.max.nf` of `r n.stocks`) and $\lambda_{F>0}$ (`r n.low.lam.at.max.fish` of `r n.stocks`) were found near the maximum abundance in the time series (Figures \@ref(fig:dd-nf-stock-plt) and \@ref(fig:dd-fish-stock-plt)). Atlantic cod in the Northwest Atlantic comprised `r n.cod.lam.not.at.max.nf` of the `r n.stocks - n.low.lam.at.max.nf` stocks in which the lowest $\lambda$ was observed below the maximum abundance, the low $\lambda$ occurred at population abundances that were remarkably similar for these Atlantic cod stocks (ranging between `r pa.low.cod.min` and `r pa.low.cod.max`). For two of the `r n.stocks - n.low.lam.at.max.nf` stocks, the lowest $\lambda_{F=0}$ estimates were less than 0.005 smaller than the estimate at the maximum abundance, for the other four stocks the decline in $\lambda_{F=0}$ from the highest abundance observed was somewhat pronounced, ranging from `r low.4.lam.40.nf` to `r low.lam.40.nf` (Tables \@ref(tab:tab-lambdas) and \@ref(tab:tab-lamda-diffs) and Figures \@ref(fig:dd-nf-stock-plt) and \@ref(fig:dd-fish-stock-plt) ).

## Demographic Parameters

The lifetime reproductive output and generation length were estimated for each cohort in which data were available. When excluding fishing mortality, the overall median lifetime reproductive output, (i.e., the number of females produced by one female that survive to reproduce) was `r r0.med` (mean = `r r0.mn`; Figures \@ref(fig:r0-6-plt) and \@ref(fig:r0-other-plt)). When the realized fishing mortality was accounted for, the median reproductive output dropped to `r r0.med.fish` (mean = `r r0.mn.fish`), this again highlighted the importance of large reproductive events for survival of these stocks and the latent capacity for population growth (Figures \@ref(fig:r0-6-plt) and \@ref(fig:r0-other-plt)). For stocks which were found across the north Atlantic, the lifetime reproductive output when excluding fishing mortality tended to decline as your move west, this again was most evident for the Atlantic cod stocks.  While the reproductive potential of cohorts within a stock could be highly variable, there were no clear temporal trends in the $R_0$ (Figures \@ref(fig:r0-nf-ts-plt) and \@ref(fig:r0-fish-ts-plt)).

The generation lengths tended to become more similar between stocks both intra- and inter-specifically when accounting fishing mortality (Figures \@ref(fig:gl-6-plt) and \@ref(fig:gl-other-plt)). In addition, the generation length of *C. harengus*, generally considered a short lived species, in the Central and NW Atlantic tended to be higher than the generation length of Atlantic cod and haddock, especially for the stocks in the NE Atlantic (Figure \@ref(fig:gl-6-plt)). There were no obvious patterns in the temporal trends in the generation length (Figures \@ref(fig:gl-nf-ts-plt) and \@ref(fig:gl-fish-ts-plt)).

## Recovery Potential

The recovery potential in the absence of fishing was relatively high for the majority of species (Figure \@ref(fig:dub-time-map)). Of the `r n.stocks` stocks in this analysis, `r n.dub.10` had a high recovery potential (we define 'high' as at least `r 100*dub.p.high`% of the simulations having experienced a doubling in size) within 10 years, by year 20 this had increased to `r n.dub.20` stocks (Figure \@ref(fig:dub-ts)).

The stocks in which recovery potential was low (we define low as < `r 100*dub.p.low`% of the simulations experiencing a doubling in size) exhibited clear regional trends. Of the `r n.not.dub.20` stocks in this analysis in which the recovery potential was low after 20 years, `r n.alaska.slow` of the stocks were in Alaska (`r n.alaska` stocks in total), and `r n.nwa.slow` stocks in NAFO Regions 2-4 in the NW Atlantic (`r n.nafo.24` stocks were in NAFO Regions 2-4). This contrasts to only `r n.nea.slow` of the `r n.nea` stocks in the NE Atlantic with a low recovery potential within 20 years in the absence of fishing (Figure \@ref(fig:dub-time-map)).

The recovery potential was influenced both by $\lambda_{F=0}$ and the variability in $\lambda_{F=0}$. Stocks with a median $\lambda_{F=0} < 1$  had moderate (defined as stocks in which 50-75% of the simulations doubled within 20 years) to high (> 75% doubled in size within 20 years) recovery potential whenever the standard deviation in lambda was greater than approximately 0.4 (Figure \@ref(fig:dub-vs-sd)). Stocks with a median $\lambda_{F=0}$ near 1 and with low variability were the stocks in which the recovery potential tended to be categorized as low (< 50% of the simulations doubled within 20 years).

<!--chapter:end:tmp-04-results.Rmd-->

# DISCUSSION {#disin}

I think we disagree with @kindsvaterTenPrinciplesEvolutionary2016 because we can't generalize based on life-history very easily (cod haddock herring).

Looks like density dependence is real. Dave's pet theory is there is solid evidence of density dependence for many stocks above around 0.4, below 0.4 the evidence suggest that density dependence weakens for some stocks, BUT, some stocks have an ability to very occasionally very spectacularly take advantage of the environment.  But not all stocks and species have this ability and the strength and frequency seems to be influenced by where you live and what species you are. Little evidence for high lambda for most stocks in the NW Atlantic, while many stocks assessed in the North Sea had anomalously high lambdas in 1 or more years. Cod show this  perfectly....

Looks like the median lambda is around 0.2 at low abundance (mean is much higher because outliers) for many stocks. So potential growth at low abundance is around 20%, tends to trend towards 0 as you approach the maximum observed abundance. This is very interesting b/c on average is suggests that the maximum observed abundance may be a reasonable proxy for the carrying capacity of a stock. Of course there is a great deal of variability around this, which may be related to the timing of the start of the time series in relation to historical fishing activity, fishing pressure, environmental variability, etc... but still it is kinda interesting.

The recovery potential for some species is tied to the variability in their growth rate... Based on historically observed lambdas, all the Haddock stock in this analysis were likely to double in size within a decade of fishing cessation. The worst situation is low lambda + low variability... you can have a low lamda 'most of the time', but if highly variable the stock can take off (this in many ways is what fisheries science is based on)... Also of note these large events do tend to be far more likely when abundance is < 40% of maximum (of course this would be in part because it's easier to go from having \$10 to \$20 than it is to go from \$1,000 to \$2,000).

Idea- realized productivity of cod/haddock has been highly variable. But from life history theory you would have expected that to be herring/smelt. Episodic recruitment events. (boom and bust?) not population boom but more like a baby boom.

Likelihood of stock collapse related to life history characteristics, maybe reframe relative to large productivity events of cohorts – link to idea of cod/haddock above. No expectation that intersection between life history traits and environmental conditions to result in such marked differences among populations within a species in realized productivity.

Center around the idea of equilibrium theory? That management might be fairly good for stocks that aren’t very variable, but when you make a decision based on averages you can make really bad ones.

By ignoring the density dependence patterns observed in this analysis, the doubling time analysis would overestimate the doubling time for stocks that were at low abundance with classic negative density dependence dynamics, while it would underestimate it for stocks with positive density dependence. Here perhaps talk about infrequent big events and how using a theoretical rmax within a logistic model isn't a good idea because this will way over-estimate recovery potential at low abundance.

The optimization procedure assumes that both the estimates of abundance and the fishery removals from the stock assessment are correct. *A couple sentences about why fisheries removals may be underestimated*. This would result in the estimates of the $\lambda$ excluding fishing mortality in this analysis being lower than reality, as this 'missing' mortality was accounted for by reductions in either fecundity or an increase in natural mortality. The optimization routine resulted in a median decrease in the no-fishing $\lambda$ estimates of approximately `r med.lam.per.diff`%, the overall directionality of this change is consistent with fishing mortality being generally underestimated.

Fisheries appear to slightly weaken the effects of density dependence overall. Maybe due to fisheries having positive density dependence? Seemingly below around 0.5 fishing leads to a positive lambda, below 0.5 it is a negative lambda... that's kinda kooky cool!

The stocks which reached very low abundance (< 10% of maximum) tended to have high values of lambda, suggesting strong compensation in stocks when they are at extremely low abundance. Oddly, the path to this isn't linear, and positive density dependence (or at least density independence was evident between 10 and 40% of maximum abundance).

Probably something worth citing in here @thorsonPredictingRecruitmentDensity2020

<!--chapter:end:tmp-05-discussion.Rmd-->

<!-- Do not edit this file! -->

\clearpage

# REFERENCES {-}

<!--chapter:end:tmp-06-references.Rmd-->

---
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---
\clearpage
\newpage

# Tables {-}

```{r, process}
# This will be the meta data really...
# remove columns we don't need.

rm.meta <- which(names(meta.sub) %in% c("Meeting_or_reference","G.short","G.Species","Manage.short","LME","Area.broad","Area.long","Area.narrow","Stock.short","Stock","Model","Case","Notes","Area"))
meta.4.tab <- meta.sub[,-rm.meta]
re.ord <- c(8,2:7,1)
meta.4.tab <- meta.4.tab[,re.ord]
names(meta.4.tab) <- c("Stock", "Order", "Family", "Genus", "Species", "Minimum Age", "Maximum Age", "Management Body")

```

```{r,tab-lambdas}
# The table with the lambda values
csas_table(comp.lam.res.tab,caption ="A comparison of the $\\lambda_{F=0}$ and $\\lambda_{F>0} estimates for stock at the minimum and maximum abundance, along with the estimate at 40\\% of the maximum abundance. values greater than 1 indicate that the population would be expected to increase in size, while values less than 1 indicate the population is likely to decline in size")
```

\newpage

```{r,tab-lamda-diffs}
# Now make the table for the differences
csas_table(comp.lam.diff.res.tab,caption ="A comparison of the difference in $\\lambda_{F=0}$ and $\\lambda_{F>0} for each stock at the minimum and maximum abundance, along with the change at 40\\% of the maximum abundance. Positive values indicate that the $\\lambda$ estimae has increased as abundance declines (consistent with negative density dependence), while negative values indicate the $\\lambda$ estimate has declined (considtent with positive density dependence)")
```

```{r,tab-meta}
# Now make the table for the differences
csas_table(meta.4.tab,caption ="Summary of the stocks included in this analysis.  The minimun and maximum ages are what was used in the stock assessment.")
```

<!--chapter:end:tmp-07-tables.Rmd-->

---
output:
  pdf_document:
    fig_caption: yes
    extra_dependencies: float
linkcolor: blue
header-includes: \usepackage{float}
editor_options:
  chunk_output_type: console
---

```{r, include=F, echo=F}
knitr::opts_knit$set(eval.after = "fig.cap")
knitr::opts_chunk$set(fig.pos = 'H')
options(knitr.kable.NA = '')
require(tidyverse)
options(tinytex.verbose = TRUE)

```

# Figures {-}

```{r lambda-bx-6-fig, echo=F, fig.cap="$\\lambda$ estimates for each stock of G. morhua, M. aeglefinus, P. virens, C. harengus, P. platessa, and S. solea."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/Lambda_boxplot_6_species_",what.2.tune,".png"))

```

\newpage
```{r lambda-bx-other-fig, echo=F, fig.cap="$\\lambda$ estimates for the remainded of the stocks in this analysis, plotted by the location of the stock. The abreviated species name is provided on the x axis."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/Lambda_boxplot_other_species_",what.2.tune,".png"))

```

\newpage
```{r lambda-sd-fig, echo=F, fig.cap="Standard deviation of the $\\lambda$ estimates for each stock, ordered by species, in this analysis. "}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/Lambda_sd_",what.2.tune,".png"))
```

\newpage
```{r dd-overall-plt, echo=F, fig.cap="Density dependence of $\\lambda$ The top panel shows $\\lambda_{F=0}$ against the proportion of maximum abundance, with a GAM smooth and the GAM estimated 95\\% confidence interval. The middle panel shows $\\lambda_{F>0}$ against the proportion of maximum abundance, with a GAM smooth and the GAM estimated 95% confidence interval. The bottom panel compares the GAM smooths for $\\lambda_{F=0}$ and $\\lambda_{F>0}$."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/density_dependence_overall_",what.2.tune,".png"))
```

\newpage
```{r dd-nf-stock-plt, echo=F, fig.cap="Density dependence of $\\lambda_{F=0}$ for each stock in the analysis plotted regionally."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/density_dependence_nf_stocks_",what.2.tune,".png"))
```

\newpage
```{r dd-fish-stock-plt, echo=F, fig.cap="Density dependence of $\\lambda_{F>0}$ for each stock in the analysis plotted regionally."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/density_dependence_fish_stocks_",what.2.tune,".png"))
```

\newpage
```{r r0-6-plt, echo=F, fig.cap="$R_0$ estimates for each stock of G. morhua, M. aeglefinus, P. virens, C. harengus, P. platessa, and S. solea."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/R0_boxplot_6_species_",what.2.tune,".png"))
```

\newpage
```{r r0-other-plt, echo=F, fig.cap="$R_0$ estimates for each stock of G. morhua, M. aeglefinus, P. virens, C. harengus, P. platessa, and S. solea."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/R0_boxplot_other_species_",what.2.tune,".png"))
```

\newpage
```{r r0-nf-ts-plt, echo=F, fig.cap="$R_0$ estimates for every cohort for each species excluding the effects of fishing."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/r0_heatmap_nf_",what.2.tune,".png"))
```

\newpage
```{r r0-fish-ts-plt, echo=F, fig.cap="$R_0$ estimates for every cohort for each species including the effects of fishing."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/r0_heatmap_fish_",what.2.tune,".png"))
```

\newpage
```{r gl-6-plt, echo=F, fig.cap="Generation length estimates for each stock of G. morhua, M. aeglefinus, P. virens, C. harengus, P. platessa, and S. solea."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/GL_boxplot_6_species_",what.2.tune,".png"))
```

\newpage
```{r gl-other-plt, echo=F, fig.cap="Generation length estimates for each stock of G. morhua, M. aeglefinus, P. virens, C. harengus, P. platessa, and S. solea."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/GL_boxplot_other_species_",what.2.tune,".png"))
```

\newpage
```{r gl-nf-ts-plt, echo=F, fig.cap="Generation length estimates for every cohort for each species excluding the effects of fishing."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/GL_heatmap_nf_",what.2.tune,".png"))
```

\newpage
```{r gl-fish-ts-plt, echo=F, fig.cap="Generation length estimates for every cohort for each species including the effects of fishing."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/GL_heatmap_fish_",what.2.tune,".png"))
```

<!-- Now for the Doubling time -->

\newpage
```{r dub-time-map, echo=F, fig.cap="Doubling time estimate for each stock with no fishing mortality."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/Doubling_time_heatmap_by_order_",what.2.tune,".png"))
```

\newpage
```{r dub-ts, echo=F, fig.cap="The number of stocks which doubled in size by a given year with no fishing mortality  The dashed line represents the 77 stocks in the analysis."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/Doubling_time_time_series_",what.2.tune,".png"))
```

\newpage
```{r dub-vs-sd, echo=F, fig.cap="The recovery potential (low, moderate, or high) of each stock after 20 years compared to the median $\\lambda_{F=0}$ and the standard deviation of $\\lambda_{F=0}$."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/Doubling_time_sd_effects_",what.2.tune,".png"))
```

# Supplemental Figures {-}

<!--chapter:end:tmp-08-figures.Rmd-->