### 2024-05-21 Drafts/Paper_1/tmp-04-results.Rmd

# Results {#reses-peses}

<!-- Throughout the results we will focus primarily the six most commonly found species in the analysis (*Gadus morhua, Melogrammus aeglefinus, Clupie harengus, Plueronecties platessa, Polluck virens* and *Solea Solea*).  The complete results for all of the stocks included in the analysis can be found in the Supplemental material. -->

## Population growth rates ($\lambda$)

The overall median population growth rate when excluding the effects of fishing ($\lambda_{F=0}$) was `r med.lambda.F0`, indicating that there was a latent potential for population growth for these stocks of approximately `r med.r.F0`%. When the effect of fishing was accounted for, the median population growth rate ($\lambda_{F>0}$) declined to `r med.lambda.F`, which suggested that the populations in this analysis may have been over-fished by approximately `r -med.r.F`% annually. However, for many of the stocks included in this analysis large recruitment events resulted in large estimates of $\lambda_{F>0}$, in some years, this led to a skewed distribution of the lambda estimates, with these large events being an integral component of the stocks population dynamics (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-bx-other-fig)). The mean $\lambda_{F>0}$ for the stocks in this analysis was `r mn.lambda.F` which would suggest there remains scope for population growth even after fishing. The mean$ \lambda_{F>0}$ was lower than the median for just `r n.stocks-n.mn.higher.med.fish` of the `r n.stocks` in this analysis. The majority of these stocks (`r n.nw.ice`) were in the central or NW Atlantic, and this included `r n.cod.low.mn`  of the `r n.cod` Atlantic cod stocks in this analysis.

There were regional and species specific trends in both $\lambda_{F=0}$ and $\lambda_{F}$. Haddock (*M. aeglefinus*) and Atlantic cod (*G. morhua*) stocks had some of highest variability in lambda, both at the stock level and between stocks (Figure \@ref(fig:lambda-sd-fig)). For these two species, the variability of lambda within a stock declined as you moved from the Eastern Atlantic to the Western Atlantic (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-sd-fig))). The pattern was more pronounced for Atlantic cod, with stocks in the eastern Atlantic tending to have relatively low within stock variability. The two other species with stocks found in the NW and NE Atlantic (*C.harengus* and *P. virens*) the variability of lambda between the stocks was much lower than for Haddock and Atlantic cod and there were no clear differences between regions (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-sd-fig))). This was largely due to the within stock lambda variability being lower than the within stock variability of Haddock and Atlantic cod stocks in the NE Atlantic (Figure \@ref(fig:lambda-bx-6-fig)). Two other stocks (*A. dubious* and *T. esmarkii*) in the North Sea had highly variably lambda estimates, though notably the lambda was less variable than the estimates for Haddock and Atlantic cod in this area (Figure \@ref(fig:lambda-bx-other-fig)). For the remainder of other stocks in this analysis, years in which the value of $\lambda_{F=0}$ exceeded 1.5 were more rare. Exceptions to this included *M. merlangus* in the Celtic Sea, *P. dentatus* and *S. Chrysops* in the NW Atlantic, and the Gulf of Alaska *T. chalcogramma* stock in the Pacific Ocean (Figure \@ref(fig:lambda-bx-other-fig)).

## Density Dependence

Overall, the evidence suggested that negative density dependence was common as both $\lambda_{F=0}$ and $\lambda_{F>0}$ increased as abundance declined (Figure \@ref(fig:dd-overall-plt)). The strength of density dependence appeared to be more pronounced at low and high abundance levels (i.e., changes in abundance had larger impacts on $\lambda$ when abundance was below $\approx$ 0.4 or above $\approx$ 0.7). $\lambda_{F=0}$ only declined below 1 when the stocks approached their maximum observed abundance ($\approx$ `r pa.lam.nf.at.one`%), indicating that in the absence of fishing there was a latent capacity for population growth in these stocks.  However, when accounting for fishing mortality, $\lambda_{F>0}$ was below 1 until the stocks were below approximately `r pa.lam.fish.at.one`% of their maximum abundance. At lower abundances, the rate of increase in $\lambda_{F>0}$ was higher than $\lambda_{F=0}$, likely due to reductions in fishing mortality at low abundance, but overall $\lambda_{F>0}$ was lower than $\lambda_{F=0}$, suggesting that some fishing mortality was common even at very low abundance (Figure \@ref(fig:dd-overall-plt)).

Generally, there was an increase in the variability of the $\lambda$ estimates as abundance declines, largely driven by an increase in the magnitude of the maximum value of $\lambda$ and a small increase in the frequency of low $\lambda$ values (Figure \@ref(fig:dd-overall-plt)). For example, values of $\lambda_{F>0}$ > `r pa.lam.fish.low.bigs` were not observed at abundances > `100*pa.big.threshold`% of the maximum (similar trends were observed for $\lambda_{F=0}$, with only one point being above `r pa.lam.nf.low.bigs` when the abundance was > `r 100*pa.big.threshold`% of the maximum). Similarly, the percentage of $\lambda_{F>0}$ values below `r lam.low.threshold` was `r 100*prop.low.high.abund`% when the proportion of the abundance was >= 0.3, but this increased to `r 100*prop.low.low.abund`% of the data when the proportion of the abundance was below 0.3.

However, these general patterns were highly variable between stocks. All of the stocks included in this analysis had $\lambda_{F=0}$ estimates > 1 at their lowest observed abundances (Table \@ref(tab:tab-lambdas) and Figure \@ref(fig:dd-nf-stock-plt)). Stocks in the Northwest Atlantic (particularly Atlantic cod) and Alaska had some of the lowest $\lambda_{F=0}$ estimates at low abundance, while several stocks in the North Sea had among the highest $\lambda_{F=0}$ estimates at low abundance (Table \@ref(tab:tab-lambdas)).  Of the `r n.bl.40` stocks that had declined below 40% of the maximum abundance, only `r `n.low.40.nf` of the stocks had a $\lambda_{F=0}$ estimates < 1 at 40% of their maximum abundance; when the effects of fishing were included `r `n.low.40.nf` of the stocks had $\lambda_{F>0}$ estimates below 1 (Table \@ref(tab:tab-lambdas)).

For the vast majority of stocks the lowest values of both $\lambda_{F=0}$ (`r n.low.lam.at.max.nf` of `r n.stocks`) and $\lambda_{F>0}$ (`r n.low.lam.at.max.fish` of `r n.stocks`) were found near the maximum abundance in the time series (Figures \@ref(fig:dd-nf-stock-plt) and \@ref(fig:dd-fish-stock-plt)). Atlantic cod in the Northwest Atlantic comprised `r n.cod.lam.not.at.max.nf` of the `r n.stocks - n.low.lam.at.max.nf` stocks in which the lowest $\lambda$ was observed below the maximum abundance, the low $\lambda$ occurred at population abundances that were remarkably similar for these Atlantic cod stocks (ranging between `r pa.low.cod.min` and `r pa.low.cod.max`). For two of the `r n.stocks - n.low.lam.at.max.nf` stocks, the lowest $\lambda_{F=0}$ estimates were less than 0.005 smaller than the estimate at the maximum abundance, for the other four stocks the decline in $\lambda_{F=0}$ from the highest abundance observed was somewhat pronounced, ranging from `r low.4.lam.40.nf` to `r low.lam.40.nf` (Tables \@ref(tab:tab-lambdas) and \@ref(tab:tab-lamda-diffs) and Figures \@ref(fig:dd-nf-stock-plt) and \@ref(fig:dd-fish-stock-plt) ).

## Demographic Parameters

The lifetime reproductive output and generation length were estimated for each cohort in which data were available. When excluding fishing mortality, the overall median lifetime reproductive output, (i.e., the number of females produced by one female that survive to reproduce) was `r r0.med` (mean = `r r0.mn`; Figures \@ref(fig:r0-6-plt) and \@ref(fig:r0-other-plt)). When the realized fishing mortality was accounted for, the median reproductive output dropped to `r r0.med.fish` (mean = `r r0.mn.fish`), this again highlighted the importance of large reproductive events for survival of these stocks and the latent capacity for population growth (Figures \@ref(fig:r0-6-plt) and \@ref(fig:r0-other-plt)). For stocks which were found across the north Atlantic, the lifetime reproductive output when excluding fishing mortality tended to decline as your move west, this again was most evident for the Atlantic cod stocks.  While the reproductive potential of cohorts within a stock could be highly variable, there were no clear temporal trends in the $R_0$ (Figures \@ref(fig:r0-nf-ts-plt) and \@ref(fig:r0-fish-ts-plt)).

The generation lengths tended to become more similar between stocks both intra- and inter-specifically when accounting fishing mortality (Figures \@ref(fig:gl-6-plt) and \@ref(fig:gl-other-plt)). In addition, the generation length of *C. harengus*, generally considered a short lived species, in the Central and NW Atlantic tended to be higher than the generation length of Atlantic cod and haddock, especially for the stocks in the NE Atlantic (Figure \@ref(fig:gl-6-plt)). There were no obvious patterns in the temporal trends in the generation length (Figures \@ref(fig:gl-nf-ts-plt) and \@ref(fig:gl-fish-ts-plt)).

## Recovery Potential

The recovery potential in the absence of fishing was relatively high for the majority of species (Figure \@ref(fig:dub-time-map)). Of the `r n.stocks` stocks in this analysis, `r n.dub.10` had a high recovery potential (we define 'high' as at least `r 100*dub.p.high`% of the simulations having experienced a doubling in size) within 10 years, by year 20 this had increased to `r n.dub.20` stocks (Figure \@ref(fig:dub-ts)).

The stocks in which recovery potential was low (we define low as < `r 100*dub.p.low`% of the simulations experiencing a doubling in size) exhibited clear regional trends. Of the `r n.not.dub.20` stocks in this analysis in which the recovery potential was low after 20 years, `r n.alaska.slow` of the stocks were in Alaska (`r n.alaska` stocks in total), and `r n.nwa.slow` stocks in NAFO Regions 2-4 in the NW Atlantic (`r n.nafo.24` stocks were in NAFO Regions 2-4). This contrasts to only `r n.nea.slow` of the `r n.nea` stocks in the NE Atlantic with a low recovery potential within 20 years in the absence of fishing (Figure \@ref(fig:dub-time-map)).

The recovery potential was influenced both by $\lambda_{F=0}$ and the variability in $\lambda_{F=0}$. Stocks with a median $\lambda_{F=0} < 1$  had moderate (defined as stocks in which 50-75% of the simulations doubled within 20 years) to high (> 75% doubled in size within 20 years) recovery potential whenever the standard deviation in lambda was greater than approximately 0.4 (Figure \@ref(fig:dub-vs-sd)). Stocks with a median $\lambda_{F=0}$ near 1 and with low variability were the stocks in which the recovery potential tended to be categorized as low (< 50% of the simulations doubled within 20 years).