### 2024-05-21 Drafts/Paper_1/tmp-07-tables.Rmd

---
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---
\clearpage
\newpage

# Tables {-}

```{r, process}
# This will be the meta data really...
# remove columns we don't need.

rm.meta <- which(names(meta.sub) %in% c("Meeting_or_reference","G.short","G.Species","Manage.short","LME","Area.broad","Area.long","Area.narrow","Stock.short","Stock","Model","Case","Notes","Area"))
meta.4.tab <- meta.sub[,-rm.meta]
re.ord <- c(8,2:7,1)
meta.4.tab <- meta.4.tab[,re.ord]
names(meta.4.tab) <- c("Stock", "Order", "Family", "Genus", "Species", "Minimum Age", "Maximum Age", "Management Body")

```

```{r,tab-lambdas}
# The table with the lambda values
csas_table(comp.lam.res.tab,caption ="A comparison of the $\\lambda_{F=0}$ and $\\lambda_{F>0} estimates for stock at the minimum and maximum abundance, along with the estimate at 40\\% of the maximum abundance. values greater than 1 indicate that the population would be expected to increase in size, while values less than 1 indicate the population is likely to decline in size")
```

\newpage

```{r,tab-lamda-diffs}
# Now make the table for the differences
csas_table(comp.lam.diff.res.tab,caption ="A comparison of the difference in $\\lambda_{F=0}$ and $\\lambda_{F>0} for each stock at the minimum and maximum abundance, along with the change at 40\\% of the maximum abundance. Positive values indicate that the $\\lambda$ estimae has increased as abundance declines (consistent with negative density dependence), while negative values indicate the $\\lambda$ estimate has declined (considtent with positive density dependence)")
```

```{r,tab-meta}
# Now make the table for the differences
csas_table(meta.4.tab,caption ="Summary of the stocks included in this analysis.  The minimun and maximum ages are what was used in the stock assessment.")
```