### 2024-05-24 Drafts/Paper_1/04-results.Rmd

# Results {#reses-peses}

<!-- Throughout the results we will focus primarily the six most commonly found species in the analysis (*Gadus morhua, Melogrammus aeglefinus, Clupie harengus, Plueronecties platessa, Polluck virens* and *Solea Solea*).  The complete results for all of the populations included in the analysis can be found in the Supplemental material. -->

## Population growth rates ($\lambda$)

The overall median population growth rate when excluding the effects of fishing ($\lambda_{FM=0}$) was `r med.lambda.F0`, indicating that there was a latent potential for population growth for these populations of approximately `r med.r.F0`%. When the effect of fishing was accounted for, the median population growth rate ($\lambda_{FM>0}$) declined to `r med.lambda.F`, which suggested that the populations in this analysis may have been over-fished by approximately `r -med.r.F`% annually. Indeed, $\lambda_{FM>0}$ was below one `r prop.lam.fish.lt.1`% of the time. However, for many of the populations included in this analysis large recruitment events resulted in large estimates of $\lambda_{FM>0}$, in some years, this led to a skewed distribution of the $\lambda$ estimates, with these large events being an integral component of the populations population dynamics (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-bx-other-fig)). The mean $\lambda_{FM>0}$ for the populations in this analysis was `r mn.lambda.F` which would suggest there remains scope for population growth even after fishing for most populations, although the mean $\lambda_{FM>0}$ was less than one for `r num.mn.lambda.fish.lt.1` of the `r n.stocks` populations. Finally, the mean $\lambda_{FM>0}$ was lower than the median for just `r n.stocks-n.mn.higher.med.fish` of the `r n.stocks` in this analysis. The majority of these populations (`r n.nw.ice`) were in the central or NW Atlantic, and this included `r n.cod.low.mn`  of the `r n.cod` Atlantic cod (*G. morhua*) populations in this analysis.

There were regional and species specific trends in both $\lambda_{FM=0}$ and $\lambda_{FM>0}$. Haddock (*M. aeglefinus*) and Atlantic cod populations had some of highest variability in lambda, both at the population level and between populations (Figure \@ref(fig:lambda-sd-fig)). For these two species, the variability of $\lambda$ within a population declined as you moved from the Eastern Atlantic to the Western Atlantic (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-sd-fig))). The pattern was more pronounced for Atlantic cod, with populations in the eastern Atlantic tending to have relatively low within population variability. The two other species with populations found in the NW and NE Atlantic (*C.harengus* and *P. virens*) the variability of $\lambda$ between the populations was much lower than for Haddock and Atlantic cod and there were no clear differences between regions for these two species (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-sd-fig))). This was largely due to the within population $\lambda$ variability being lower than the within population variability of Haddock and Atlantic cod populations in the NE Atlantic (Figure \@ref(fig:lambda-bx-6-fig)). Two other populations (*A. tobianus* and *T. esmarkii*) in the North Sea had highly variably $\lambda's$, though notably they were less variable than the Haddock and Atlantic cod $\lambda's$ in this area (Figure \@ref(fig:lambda-bx-other-fig)). For the remainder of other populations in this analysis, years in which the value of $\lambda_{FM=0}$ exceeded 1.5 were more rare. Exceptions to this included *M. merlangus* in the Celtic Sea, *P. dentatus* and *S. Chrysops* in the NW Atlantic, and the Gulf of Alaska *G. chalcogrammus* population in the Pacific Ocean (Figure \@ref(fig:lambda-bx-other-fig)).

## Density Dependence

Overall, the evidence suggested that negative density dependence was common as both $\lambda_{FM=0}$ and $\lambda_{FM>0}$ increased as abundance declined (Figure \@ref(fig:dd-overall-plt)). The strength of density dependence appeared to be more pronounced at low and high abundance levels (i.e., changes in abundance had larger impacts on $\lambda$ when abundance was below $\approx$ 0.4 or above $\approx$ 0.7). $\lambda_{FM=0}$ only declined below 1 when the populations approached their maximum observed abundance ($\approx$ `r pa.lam.nf.at.one`%), indicating that there was a latent capacity for population growth in these populations.  However, when accounting for fishing mortality, $\lambda_{FM>0}$ was below one until the populations were below approximately `r pa.lam.fish.at.one`% of their maximum abundance. At lower abundances, the rate of increase in $\lambda_{FM>0}$ was higher than $\lambda_{FM=0}$, likely due to reductions in fishing mortality at low abundance, but overall $\lambda_{FM>0}$ was lower than $\lambda_{FM=0}$, suggesting that some fishing mortality was common even at very low abundance (Figure \@ref(fig:dd-overall-plt)).

Generally, there was an increase in the variability of the $\lambda$ estimates as abundance declines, largely driven by an increase in the magnitude of the maximum value of $\lambda$ and a small increase in the frequency of low $\lambda$ values (Figure \@ref(fig:dd-overall-plt)). For example, values of $\lambda_{FM>0}$ > `r pa.lam.fish.low.bigs` were not observed at abundances > `100*pa.big.threshold`% of the maximum (similar trends were observed for $\lambda_{FM=0}$, with only one point being above `r pa.lam.nf.low.bigs` when the abundance was > `r 100*pa.big.threshold`% of the maximum). Similarly, the percentage of $\lambda_{FM>0}$ values below `r lam.low.threshold` was `r 100*prop.low.high.abund`% when the proportion of the abundance was >= 0.3, but this increased to `r 100*prop.low.low.abund`% of the data when the proportion of the abundance was below 0.3.

However, these general patterns were highly variable between populations. All of the populations included in this analysis had $\lambda_{FM=0}$ estimates > 1 at their lowest observed abundances (Table \@ref(tab:tab-lambdas) and Figure \@ref(fig:dd-nf-stock-plt)). Populations in the Northwest Atlantic (particularly Atlantic cod) and Alaska had some of the lowest $\lambda_{FM=0}$ estimates at low abundance, while several populations in the North Sea had among the highest $\lambda_{FM=0}$ estimates at low abundance (Table \@ref(tab:tab-lambdas)).  Of the `r n.bl.40` populations that had declined below 40% of the maximum abundance, only `r `n.low.40.nf` of the populations had a $\lambda_{FM=0}$ estimates < 1 at 40% of their maximum abundance; when the effects of fishing were included `r `n.low.40.nf` of the populations had $\lambda_{FM>0}$ estimates below 1 (Table \@ref(tab:tab-lambdas)).

For the vast majority of populations the lowest values of both $\lambda_{FM=0}$ (`r n.low.lam.at.max.nf` of `r n.stocks`) and $\lambda_{FM>0}$ (`r n.low.lam.at.max.fish` of `r n.stocks`) were found near the maximum abundance in the time series (Figures \@ref(fig:dd-nf-stock-plt) and \@ref(fig:dd-fish-stock-plt)). Atlantic cod in the Northwest Atlantic comprised `r n.cod.lam.not.at.max.nf` of the `r n.stocks - n.low.lam.at.max.nf` populations in which the lowest $\lambda$ was observed below the maximum abundance, the low $\lambda$ occurred at population abundances that were remarkably similar for these Atlantic cod populations (ranging between `r pa.low.cod.min` and `r pa.low.cod.max`). For two of the `r n.stocks - n.low.lam.at.max.nf` populations, the lowest $\lambda_{FM=0}$ estimates were less than 0.005 smaller than the estimate at the maximum abundance, for the other four populations the decline in $\lambda_{FM=0}$ from the highest abundance observed was somewhat pronounced, ranging from `r low.4.lam.40.nf` to `r low.lam.40.nf` (Tables \@ref(tab:tab-lambdas) and \@ref(tab:tab-lamda-diffs) and Figures \@ref(fig:dd-nf-stock-plt) and \@ref(fig:dd-fish-stock-plt) ).

## Demographic Parameters

The lifetime reproductive output and generation length were estimated for each cohort in which data were available. When excluding fishing mortality, the overall median lifetime reproductive output, (i.e., the number of females produced by one female that survive to reproduce) was `r r0.med` (mean = `r r0.mn`; Figures \@ref(fig:r0-6-plt) and \@ref(fig:r0-other-plt)). When the realized fishing mortality was accounted for, the median reproductive output dropped to `r r0.med.fish` (mean = `r r0.mn.fish`), this again highlighted the importance of large reproductive events for survival of these populations and the latent capacity for population growth (Figures \@ref(fig:r0-6-plt) and \@ref(fig:r0-other-plt)). For populations which were found across the north Atlantic, the lifetime reproductive output when excluding fishing mortality tended to decline as your move west, this again was most evident for the Atlantic cod populations.  While the reproductive potential of cohorts within a population could be highly variable, there were no clear temporal trends in the $R_0$ (Figures \@ref(fig:r0-nf-ts-plt) and \@ref(fig:r0-fish-ts-plt)).

The generation lengths tended to become more similar between populations both intra- and inter-specifically when accounting fishing mortality (Figures \@ref(fig:gl-6-plt) and \@ref(fig:gl-other-plt)). In addition, the generation length of *C. harengus*, generally considered a short lived species, in the Central and NW Atlantic tended to be higher than the generation length of Atlantic cod and haddock, especially for the populations in the NE Atlantic (Figure \@ref(fig:gl-6-plt)). There were no obvious patterns in the temporal trends in the generation length (Figures \@ref(fig:gl-nf-ts-plt) and \@ref(fig:gl-fish-ts-plt)).

## Recovery Potential

The recovery potential in the absence of fishing was relatively high for the majority of species (Figure \@ref(fig:dub-time-map)). Of the `r n.stocks` populations in this analysis, `r n.dub.10` had a high recovery potential (we define 'high' as at least `r 100*dub.p.high`% of the simulations having experienced a doubling in size) within 10 years, by year 20 this had increased to `r n.dub.20` populations (Figure \@ref(fig:dub-ts)).

The populations in which recovery potential was low (we define low as < `r 100*dub.p.low`% of the simulations experiencing a doubling in size) exhibited clear regional trends. Of the `r n.not.dub.20` populations in this analysis in which the recovery potential was low after 20 years, `r n.alaska.slow` of the populations were in Alaska (`r n.alaska` populations in total), and `r n.nwa.slow` populations in NAFO Regions 2-4 in the NW Atlantic (`r n.nafo.24` populations were in NAFO Regions 2-4). This contrasts to only `r n.nea.slow` of the `r n.nea` populations in the NE Atlantic with a low recovery potential within 20 years in the absence of fishing (Figure \@ref(fig:dub-time-map)).

The recovery potential was influenced both by $\lambda_{FM=0}$ and the variability in $\lambda_{FM=0}$. Populations with a median $\lambda_{FM=0} < 1$  had moderate (defined as populations in which 50-75% of the simulations doubled within 20 years) to high (> 75% doubled in size within 20 years) recovery potential whenever the standard deviation in $\lambda$ was greater than approximately 0.4 (Figure \@ref(fig:dub-vs-sd)). Populations with a median $\lambda_{FM=0}$ near 1 and with low variability were the populations in which the recovery potential tended to be categorized as low (< 50% of the simulations doubled within 20 years).