### 2024-05-28 Drafts/Paper_1/05-discussion.Rmd
Full draft!

# DISCUSSION {#disin}

Here we deconstructed the ecological demographics of numerous harvested marine fish populations . In general, these populations have latent capacity for growth and, even when accounting for the influence of fishing, these growth rates are very close to 1 in general.  There is evidence that fisheries stocks are fished in such a way that they will tend to decline until they have declined below 30% of the maximum population size observed, after that most fished populations tend to experience population growth rates above 1. Most populations appear to exhibit evidence for negative density dependence, which is in line with general beliefs within fisheries science and population ecology more generally, however there are some populations where the evidence suggests some level of density independence or weak positive density dependence , including most of the *G. morhua* populations in NW Atlantic.  Both variability and median $\lambda_{F=0}$ influence recovery, stocks with a relatively high variability in $\lambda_{F=0}$ had a moderate-high recovery potential even when their median $\lambda_{F=0}$ was below one. Effectively recovery of the stocks was likely within 20 years whenever the median $\lambda_{F=0}$ was greater than 1.06 or the standard deviation of lambda was greater than 0.2.

For every population in this analysis the $\lambda_{F=0}$ at the minimum observed abundance was estimated to be greater than one; suggesting all of the populations have the potential for some degree of recovery given enough time when there is no fishing mortality. However, there are seven populations in which $\lambda_{F=0}$ at 40% of the maximum abundance is below 1, five of these populations are found in the NW Atlantic, and five of these are in the Gadus genus (*G. morhua* (3) and *G. chalcogrammus* (2)). For both of these Gadus species, there are examples of populations in which the latent capacity for growth is far more robust.  For example, *G. chalcogrammus* is one of the largest fisheries in the world and the vast majority of the landings are from the Eastern Bering Sea population which has a high latent capacity for growth. However the $\lambda_{F=0}$ estimates for two neighboring populations in Alaska are much lower, and it is likely that these populations experience much lower fishing pressure because of their low latent capacity for growth.  
In contrast, in the NW Atlantic, several *G. morhua* populations supported large distant waters fishery for centuries (citations), however the data used in this analysis indicated the majority of the stocks in the NW Atlantic have minimal recovery potential. The maximum abundance used for these populations was likely a small fraction of the historic maximum abundance of these stocks, primarily due to over fishing by the industrial fishery that developed in the latter half of the 20th century (Cite Rose book maybe?, others). Here it is likely that the fishery induced changes to the population, which combined with changes in the environment (Cite cold water and seals) has resulted in these populations having little latent capacity for growth or recovery. Indeed, it has been suggested that the SGSL cod population will go extinct unless natural mortality declines significantly (Doug’s predator pit stuff); the results of this analysis largely align with these suggestions given the low recovery potential of this stock and it's extremely low latent capacity for growth. However, as one moves towards the eastern Atlantic, the low recovery potential of *G. morhua* populations was no longer observed, with stocks around mainland Europe all having a high recovery potential. As with *G. chalcogrammus*, this indicates that regional variability is a factor in the ability of populations to withstand fishing pressure. While some Gadus populations have an ability to withstand large fisheries due to their latent capacity for growth, the life-history at a species level is insufficient to predict how a population will respond to fishing. Moreover, this analysis is relevant for only the period under study, the recovery potential assumes that the underlying population demographic rates in the future will be similar to those observed in the past. However, as is likely the case with the *G. morhua* stocks in the NW Atlantic, the response for a population may change over time due to environmental, ecological, and anthropocentric factors (CITE!).

More generally, we observed that populations with high variability in $\lambda_{F=0}$ have a high recovery potential, however, this variability does not necessarily follow on from the expectations of conventional life-history theory (Cite Kingsvader). For example, *C. harengus* are a small, relatively short lived, low trophic level pelagic species (Cite!), based on life-history theory *C. harengus* should exhibit highly variable population growth potential in which they ‘exploit’ opportunities within the environment (CITE). However, the variability of $\lambda_{F=0}$ for the *C. harengus* populations was much lower than the *G. morhua* and *M. aeglefinus* populations found throughout the NE Atlantic. Other small pelagic populations, such at *A. tobianus* and *T. esmarkii* do have elevated  $\lambda_{F=0}$ as life-history theory would expect for small low trophic level species. However, caution would be advised in suggesting that other populations of these species would have similar population dynamics, especially given these two populations are found in the North Sea, which appears to be an environment in which high variability in $\lambda_{F=0}$ is commonplace. In addition, while stocks with a high variability in $\lambda_{F=0}$ do have years with very high population growth, these events happen at irregular intervals, even when populations are at low abundances. As such, for populations in which recovery plans and conservation targets are being developed, the developers must account for *both* the magnitude and the frequency of these events. Populations do not *always* increase at their maximum intrinsic rate of growth when at low abundance, but they may have the ability to do so occasionally.

The optimization procedure used for the life-table reconstruction herein assumes that both the estimates of abundance and the fishery removals from the stock assessment are correct. Given the underlying assessment models estimate biomass and abundance from fishery independent surveys that have undergone multiple years of peer review, we are confident that the underlying trends captured by the assessments provide a reasonable snapshot of the population dynamics of these population. However, it is likely that the fisheries removals used in the stock assessments do not capture all the mortality due to fishing, unreported fishing, bycatch and incidental mortality could all result in higher fishery removals than estimated in the assessments (CITE). This would result in the estimates of  $\lambda_{F=0}$ being lower than reality; ‘missing’ mortality would have been accounted for by reductions in fecundity and/or an increase in natural mortality. The optimization routine resulted in a median decrease in the no-fishing $\lambda_{F=0}$ estimates of approximately 6.8% when compared to the non-optimized estimates; the directionality of this change is consistent with fishing mortality being underestimated. Thus, $\lambda_{F=0}$ is likely an underestimate of the latent capacity for growth of these populations, this magnitude of this bias in any year would be dependent on the level of unaccounted for fishing mortality a population experienced.

Here we provide an overview of the population growth and recovery potential of numerous harvested marine fish populations. There is a latent capacity for growth for most populations and the majority of the populations appear to have the ability to recover relatively quickly if fishing mortality is reduced to low levels. However, there can be significant intra-species variability which may be related to environmental, ecological, and anthropocentric factors. This makes it difficult to use species level life-history traits to predict the population dynamics of specific populations.  The life-table reconstruction enables the underlying demographic rates to be estimated for populations with sufficient age-structured data, this provides an opportunity of understanding the dynamics of these populations using concepts developed in conservation biology and ecology.

## Ignore this, unless something is in here you want to pull into discussion

I think we disagree with @kindsvaterTenPrinciplesEvolutionary2016 because we can't generalize based on life-history very easily (cod haddock herring).

Looks like density dependence is real. Dave's pet theory is there is solid evidence of density dependence for many stocks above around 0.4, below 0.4 the evidence suggest that density dependence weakens for some stocks, BUT, some stocks have an ability to very occasionally very spectacularly take advantage of the environment.  But not all stocks and species have this ability and the strength and frequency seems to be influenced by where you live and what species you are. Little evidence for high lambda for most stocks in the NW Atlantic, while many stocks assessed in the North Sea had anomalously high lambdas in 1 or more years. Cod show this  perfectly....

Looks like the median lambda is around 0.2 at low abundance (mean is much higher because outliers) for many stocks. So potential growth at low abundance is around 20%, tends to trend towards 0 as you approach the maximum observed abundance. This is very interesting b/c on average is suggests that the maximum observed abundance may be a reasonable proxy for the carrying capacity of a stock. Of course there is a great deal of variability around this, which may be related to the timing of the start of the time series in relation to historical fishing activity, fishing pressure, environmental variability, etc... but still it is kinda interesting.

The recovery potential for some species is tied to the variability in their growth rate... Based on historically observed lambdas, all the Haddock stock in this analysis were likely to double in size within a decade of fishing cessation. The worst situation is low lambda + low variability... you can have a low lamda 'most of the time', but if highly variable the stock can take off (this in many ways is what fisheries science is based on)... Also of note these large events do tend to be far more likely when abundance is < 40% of maximum (of course this would be in part because it's easier to go from having \$10 to \$20 than it is to go from \$1,000 to \$2,000).

Idea- realized productivity of cod/haddock has been highly variable. But from life history theory you would have expected that to be herring/smelt. Episodic recruitment events. (boom and bust?) not population boom but more like a baby boom.

Likelihood of stock collapse related to life history characteristics, maybe reframe relative to large productivity events of cohorts – link to idea of cod/haddock above. No expectation that intersection between life history traits and environmental conditions to result in such marked differences among populations within a species in realized productivity.

Center around the idea of equilibrium theory? That management might be fairly good for stocks that aren’t very variable, but when you make a decision based on averages you can make really bad ones.

By ignoring the density dependence patterns observed in this analysis, the doubling time analysis would overestimate the doubling time for stocks that were at low abundance with classic negative density dependence dynamics, while it would underestimate it for stocks with positive density dependence. Here perhaps talk about infrequent big events and how using a theoretical rmax within a logistic model isn't a good idea because this will way over-estimate recovery potential at low abundance.

The optimization procedure assumes that both the estimates of abundance and the fishery removals from the stock assessment are correct. *A couple sentences about why fisheries removals may be underestimated*. This would result in the estimates of the $\lambda$ excluding fishing mortality in this analysis being lower than reality, as this 'missing' mortality was accounted for by reductions in either fecundity or an increase in natural mortality. The optimization routine resulted in a median decrease in the no-fishing $\lambda$ estimates of approximately `r med.lam.per.diff`%, the overall directionality of this change is consistent with fishing mortality being generally underestimated.

Fisheries appear to slightly weaken the effects of density dependence overall. Maybe due to fisheries having positive density dependence? Seemingly below around 0.5 fishing leads to a positive lambda, below 0.5 it is a negative lambda... that's kinda kooky cool!

The stocks which reached very low abundance (< 10% of maximum) tended to have high values of lambda, suggesting strong compensation in stocks when they are at extremely low abundance. Oddly, the path to this isn't linear, and positive density dependence (or at least density independence was evident between 10 and 40% of maximum abundance).

Probably something worth citing in here @thorsonPredictingRecruitmentDensity2020

While alternative frameworks have been suggested and are being implemented globally, (e.g., Ecosystem Based management) the uptake of these methods has been relatively slow due to a lack of mechanistic understanding, a lack of data, and the more complicated nature of the science advice alternative methods provide.