### 2024-07-02 Drafts/Paper_1/02-intro.Rmd
Updated Intro with HB edits

# INTRODUCTION  {#intro}

Marine fishes were once believed to be essentially inexhaustible and thus extremely resilient to the effects of fishing [@mcintoshFisheriesInterNationalCouncil1919a]. However, high profile collapses of large fisheries and a subsequent lack of rebuilding (cite, cite) have fundamentally challenged this perception. The assumption of resilience originated from early theoretical population dynamics modeling [e.g., @verhulstNoticeLoiQue1838] that culminated in the development of modern fisheries science, largely built around the concept of maximum sustainable yield [MSY, @rickerStockRecruitment1954; @schaeferAspectsDynamicsPopulations1954; @bevertonDynamicsExploitedFish1957]. From its inception, MSY was controversial [@finleyAllFishSea2019; @larkinEpitaphConceptMaximum197], yet the theoretical underpinning in equilibrium concepts and the associated idea that annual reproductive output and/or juvenile survival rates increase as abundance declines (implying harvesting activity will increase the per capita productivity of a population) persisted. The vast majority of modern assessments incorporate a stock-recruit function (cite, cite) making compensatory dynamics a foundational assumption about the life history of a species. Such fisheries assessments implicitly assume that lifetime reproductive rates in the absence of fishing would be expected to be greater than one (Myers cite, cite) and that even heavily-exploited fish populations retain a latent capacity for increase (cite, cite). Although MSY and equilibrium concepts are relatively simple and intuitive, and provide a means of developing tractable mechanistic models, questions remain regarding the applicability of these concepts and whether they adequately reflect the reality of the system being modeled [[@wormRebuildingGlobalFisheries2009; cite cite], particularly for species with markedly different life history characteristics (cite, cite).

Life history characteristics are now considered indicative of a species’ compensatory capacity (cite, cite) and are increasingly used within data-limited assessment approaches to provide management advice (cite). Theory suggests that small fishes, characterized by numerous offspring, rapid growth, early maturity and short lifespans, should have greater compensatory capacity, greater resilience to environmental variability and consequently, be better able to withstand fishing pressure [@kindsvaterTenPrinciplesEvolutionary2016; cite, cite]. Conversely, life histories characterized by longer lifespans and later maturity, or those with lower per-capita reproductive output are considered less resilient (cite). Furthermore, closely related species are anticipated to be more similar than distantly related species owing to a shared evolutionary history (Thorson fishlife), implying less heterogeneity among populations of a single species than among closely-related species. There is a general impression that forage fishes or populations at lower trophic levels are better able to take advantage of favorable environmental conditions and grow more quickly, particularly from low abundance (cite, cite). This resilience should manifest as a high latent capacity for population growth in the absence of fishing, as well as substantial variability in realized population growth rates among years. Dramatic increases in eulecan (FIND OTHER EXAMPLES) following years of particularly low abundance have provided empirical evidence of high compensatory capacity in forage fishes (cites). However, it remains unclear if correlated and compounding ecological and environmental processes (e.g. climate change effects, predatory prey dynamics) can largely obscure the importance of life history on realized productivity, making evolutionary history and life history characteristics a weak foundation for data-poor assessment in light of our changing oceans.

Stock assessment data from harvested marine telosts provide a unique opportunity to explore demographic variability and realized compensatory responses from a conservation biology or population ecology perspective. The long assessment history of these populations means both that their life history characteristics are well-known and that it is possible to quantify realized variability in survival and reproductive output, while accounting for the influence of fisheries. We based our analyses on a database containing age specific abundance information for 77 populations in the north Atlantic and the northeast Pacific Oceans [@charbonneauAgestructuredmarinefishdatabase2022]. By constructing life tables from abundance at age and catch at age data, we calculated common demographic metrics, including  annual population growth rates (cite), doubling time (cite), lifetime reproductive success (cite), and generation time (cite). Comparing the strength and magnitude of variability in realized productivity among species with different life histories as well as among populations of the same species yielded unexpected insights. Our results have implications for sustainable management of these exploited marine teleosts and could be used inform future management actions for populations that have collapsed. More broadly, our results challenged longstanding ideas on the applicability of equilibrium theory, life history theory, and compensatory dynamics to explain expected productivity in fishes.