### 2025-02-03 Drafts/Paper_1/02-intro.Rmd

# INTRODUCTION  {#intro}

Marine fishes were once believed to be essentially inexhaustible and thus extremely resilient to the effects of fishing [@mcintoshFisheriesInterNationalCouncil1919a; @nielsenEvolutionFisheriesManagement1976]. However, high profile collapses of large fisheries and a subsequent lack of rebuilding [@hutchingsCollapseRecoveryMarine2000; @frankAlleeEffectsCompensatory2000; @myersMaximumPopulationGrowth1997] have fundamentally challenged this perception. The assumption of resilience originated from early theoretical population dynamics modeling [e.g., @verhulstNoticeLoiQue1838] that culminated in the development of modern fisheries science, largely built around the concept of maximum sustainable yield [MSY, @rickerStockRecruitment1954; @schaeferAspectsDynamicsPopulations1954; @bevertonDynamicsExploitedFish1957]. Despite long standing concerns regarding the applicability of MSY [@finleyAllFishSea2019; @larkinEpitaphConceptMaximum1977], its theoretical underpinning in equilibrium concepts and the associated idea that annual reproductive output and/or juvenile survival rates increase as abundance declines persisted along with the implication that harvesting activity will increase the per capita productivity of a stock [@schaeferAspectsDynamicsPopulations1954]. Most modern assessments incorporate a stock-recruit function [@rickerStockRecruitment1954; @bevertonDynamicsExploitedFish1957; @hilbornQuantitativeFisheriesStock1992; @subbeyModellingForecastingStock2014] making compensatory dynamics a foundational assumption about the life history of a species. Such fisheries assessments implicitly assume that lifetime reproductive rates in the absence of fishing would be expected to be greater than one [@hilbornQuantitativeFisheriesStock1992; @myersPopulationDynamicsExploited1995] and that even heavily-exploited fish stocks retain a latent capacity for increase [maceRelationshipsCommonBiological1994]. Although MSY and equilibrium concepts are relatively straightforward, intuitive, and provide a means of developing tractable mechanistic models, questions remain regarding their applicability [@finleyAllFishSea2019; @larkinEpitaphConceptMaximum1977]. There are concerns about whether they accurately reflect the the system being modeled [@hilbornQuantitativeFisheriesStock1992; paulyPrimaryProductionRequired1995; @wormRebuildingGlobalFisheries2009], particularly for species with markedly different life history characteristics [hilbornFisheriesInducedChangesGrowth2008].

Life history characteristics are now considered indicative of a species’ compensatory capacity [@myersStockRecruitmentGeneralizations2001; @roseCompensatoryDensityDependence2001; @winemillerLifeHistoryStrategies2005] and are increasingly used within data-limited assessment approaches to provide management advice [@martellSimpleMethodEstimating2013; @pedersenStochasticSurplusProduction2017] . Theory suggests that small fishes, characterized by numerous offspring, rapid growth, early maturity and short lifespans, should have greater compensatory capacity, greater resilience to environmental variability and consequently, be better able to withstand fishing pressure [@kindsvaterTenPrinciplesEvolutionary2016; @lawFishingSelectionPhenotypic2000]. Conversely, life histories characterized by longer lifespans and later maturity, or those with lower per-capita reproductive output are considered less resilient [@winemillerLifeHistoryStrategies2005]. Furthermore, the life history traits of closely related species are anticipated to be more similar than those distantly related species owing to a shared evolutionary history [@thorsonPredictingRecruitmentDensity2020]. This implies there should be less heterogeneity in the productivity among stocks of a single species than among closely-related species. There is a general impression that forage fishes or stocks at lower trophic levels are better able to take advantage of favorable environmental conditions and grow more quickly, particularly from low abundance [@essingtonFishingMarineFood2006]. This resilience should manifest as a high latent capacity for population growth in the absence of fishing, as well as substantial variability in realized population growth rates among years. However, it remains unclear if correlated and compounding ecological and environmental processes (e.g. climate change effects, predator-prey dynamics) can largely obscure the importance of life history on realized productivity, making evolutionary history and life history characteristics a weak foundation for data-poor assessment in light of our changing oceans.

Stock assessment data from harvested marine telosts provide a unique opportunity to explore demographic variability and realized compensatory responses. The long assessment history of these stocks means that their life history characteristics are well-known and that it is possible to quantify realized variability in survival and reproductive output, while accounting the impact of fishery removals. We based our analyses on 77 marine fish stocks from a database of age-specific data [@charbonneauAgestructuredmarinefishdatabase2022]. By constructing life tables from abundance-at-age and catch-at-age data, we calculated common demographic metrics, including  annual population growth rates, doubling time, lifetime reproductive success, and generation time [@caswellMatrixPopulationModels2001]. Comparing the strength and magnitude of variability in realized productivity among species with different life histories as well as among stocks of the same species yielded unexpected insights. Our results have implications for sustainable management of these exploited marine teleosts and could be used to inform future management actions for stocks that have collapsed. More broadly, our results challenged longstanding ideas on the applicability of equilibrium theory, life history theory, and compensatory dynamics to explain expected productivity in fishes.